### 2018-02-22 paper.tex
Access Format

\documentclass{ieeeaccess}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{relsize}
\usepackage{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\usepackage{array}
%\usepackage{caption}
%\usepackage[caption=false,font=footnotesize]{subfig}
%\usepackage{fixltx2e}
\usepackage{float}
\usepackage{url}
%\usepackage{becs}
\usepackage{listings}
\lstset{%
  basicstyle=\scriptsize,
  numbers=left,
  numberstyle=\tiny,
  stepnumber=1,
  xleftmargin=2em
}
\usepackage{color}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\let\latexvec=\vec
\let\vec=\latexvec
\interdisplaylinepenalty=2500
\newtheorem{definition}{Definition}
\newtheorem{exampleb}{Example}
\newcommand{\subsubsubsection}[1]{\medskip\par\emph{#1:}}
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
\history{Date of publication xxxx 00, 0000, date of current version xxxx 00, 0000.}
\doi{10.1109/ACCESS.2017.DOI}

\title{An Improved Metric for the Analysis of Swarms}
\author{\uppercase{Neil Eliot}\authorrefmark{1},
\uppercase{David Kendall\authorrefmark{1}, and Michael Brockway\authorrefmark{1}}}
\address[1]{Northumbria University, Department of Computing and Information Sciences, Newcastle upon Tyne, NE1 8ST}

\markboth
{Eliot \headeretal: An Improved Metric for the Analysis of Swarms}
{Eliot \headeretal: An Improved Metric for the Analysis of Swarms}

\corresp{Corresponding author: Neil Eliot (e-mail: neil.eliot@northumbria.ac.uk).}

\begin{abstract}
An important criterion in the assessment of swarm coordination algorithms is
the extent to which they cause unnecessary movement, i.e. movement which
occurs as a side affect of the algorithm rather than the the intended
function. This paper examines the widely used distance metric as a mechanism
to measure the internal movement of agents and introduces a new
\emph{cohesion/repulsion metric}.  Both metrics are based upon the resultant
internal movement of a swarm identified by analysing the changes in the
inter-agent interactions. The two metrics differ in their approach to
identifying the changes. The distance metric uses variations in the
inter-agent distances and the new metric uses the resultant magnitudes from
the agents' \emph{inter-agent vector magnitudes}  which we claim offers a
number of advantages over the distance metric.  Both metrics allow a
comparison of the effects of different swarming algorithms on a swarm's
structure to identify the `level' of change in a swarm's internal structure
which is an indication of the amount of unnecessary movement an algorithm
creates.  In addition the \emph{cohesion/repulsion} metric also identifies a
swarm as expanding or cohesive.  This additional information has practical
benefits for the implementation of algorithms for heterogeneous swarms.
\end{abstract}

\begin{keywords}
Swarming, Swarm Dynamics, Mobile Sensor Networks, Algorithms.
\end{keywords}

\titlepgskip=-15pt
\maketitle

\section{Introduction\label{sec:intro}}

%\section{Introduction\label{methods:SwarmStability}}
\emph{Swarming} in the animal kingdom of ants, bees, fish and birds for instance has long been studied by scientists. From these studies mathematical models and algorithms have evolved. The models and algorithms have in turn captured the interest of computing scientists who are interested in applying them to large groups of autonomous mobile \emph{agents} (`robots'). The cooperative coordination of these agents can take many forms such as following a set path~\cite{HCS:09}, existing in a static space~\cite{EP:10, GP:02, GP:04} or foraging as a colony~\cite{HER:11, GK:07}. One of the attributes of swarms that has captured the interest of scientists is that the models and algorithms used to coordinate them are generally sets of simple rules. These simple rules cause the agents to appear to work cooperatively. These rules can also generate movement that does not contribute to the goal of the swarm this movement is a direct result of the agents attempting to move towards an optimum position within the structure. This internal movement can be identified by analysing the changes in the inter-agent interactions. Measuring this unnecessary movement allows the effectiveness of a swarming algorithm to be evaluated. If the internal movement can be reduced then stability of the swarm is improved and more of the swarms energy is focused on the goal. Reducing the internal movement also provides a more stable platform for the deployment of sensor arrays. Both metrics allow this evaluation to be carried out when all the sensor ranges are the same however the distance metric is not able to identify the internal movement accurately when the sensor ranges of each agent varies producing varying distances even though the interactions at the magnitude level indicates the agents are in fact stable and optimum. This paper discusses the two metrics and how they differ in their approach to identify these changes. The \emph{distance metric} which uses variations in the inter-agent distances, as discussed by Navarro et al.~\cite{NIM:09}. The new metric uses the resultant magnitudes of the agents' \emph{inter-agent vectors} that are induced by an agent interacting with its neighbours which allows a more accurate analysis of a swarm when using non-homogeneous sensor distances. The metric also identifies when a swarm is expanding or contracting.

This paper presents a swarm model and then applies the two metrics to this model to present a comparison of the two metrics and uses the new metric to identify a swarm's `state'.

%% \section{Metric principles\label{section:MagnitudeDynamics}}
%% This section discusses the theory behind swarm movement and the effects that can be measured using both the metrics described above.
\section{Swarm modelling}\label{SwarmModelling}
Currently, much swarm research uses field effects as the method of modelling inter-agent interactions~\cite{BAF:06, BAFVM:06, BM:09, APZDAMC:09, GP:02, GP:04, GP:04a, GP:05, GP:11, MYP:09}. The models usually use two field effects to implement the swarming characteristic. These effects are \emph{cohesion}, to draw agents closer, and \emph{repulsion} to prevent agents colliding. \emph{Field effects} are the ranges around an agent that determine the effect other agents have upon its movement~(Figure~\ref{methods:FieldEffects}). It is usual for the cohesion field to have a radius $C_b$ which is larger than the repulsion radius $R_b$. When an agent ($b'$) moves into the \emph{neighbour field} of an agent ($b$) then $b'$ is said to be a neighbour of $b$ and is subject to cohesion. When an agent $b'$ moves into the repulsion field of $b$ then $b$ also subjected to repulsion. When the repulsion magnitude exceeds the cohesion magnitude the agent will have a tendency to move away from $b'$, i.e. repulsed. When an agent $b$ moves too close to an obstacle, i.e. within the obstacle repulsion range $O_b$, it has repulsion vector applied and the agent will tend towards moving away from the obstacle.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/FieldEffects}
{Agent field effects.\label{methods:FieldEffects}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=5cm]{figures/FieldEffects}
%% \end{center}
%% \caption{Agent field effects\label{methods:FieldEffects}}
%% \end{figure}

\section{Inter-agent vector magnitude effect on internal movement}\label{Section:StabilityMagnitude}
Figure \ref{methods:Stability5} shows the cohesion and repulsion vector contributions to $v_c(b)$ and $v_r(b)$ due to neighbour $b'$, as given in (Equations \ref{eq:FlyToCentre1} and \ref{eq:Repulsion1}). Notice that the vectors are along the line of separation $bb'$. \\

%% Equation (\ref{eq:Neighbours1}) identifies the set of agents that are neighbours of $b$.
%% 
%% \begin{equation}\label{eq:Neighbours1}
%% nbr(b) \buildrel \Delta \over = \{b' \in S | |bb'| <= C_b\}
%% \end{equation}‎

Equation (\ref{eq:FlyToCentre1}) identifies the resultant cohesion effect of all the neighbours of $b$. $nbr(b)$ defines the set of all the agents in the swarm that are a neighbour of $b$; a neighbour is any agent that fails within the cohesion field.

\begin{equation}\label{eq:FlyToCentre1}
v_{c}(b) = \frac{1}{|nbr(b)|}\left({\mathlarger{\sum_{b' \in nbr(b)}}{bb'}}\right)
\end{equation}

Equation (\ref{eq:Repulsion1}) identifies the resultant repulsion effect of all the neighbours of $b$ where $rep(.)$ returns a set of all agents that are within the repulsion field.

\begin{equation}
\label{eq:Repulsion1}
v_{r}(b) = 
\frac{1}{|rep(b)|}
\left(
\mathlarger{\mathlarger{\sum_{b' \in rep(b)}}}
{\left( 1-\frac{|bb'|}{R_b} \right)}
{bb'}
\right)
\end{equation}

Using the cohesion and repulsion vectors generated by the relationship of $b$ to its neighbour $b'$ a resultant vector can be calculated. This vector creates an agent characteristic that can be used as a metric. Summing the vectors creates a resultant vector that affects the agent. Summing the vectors also provides an indication of the direction an agent will move based on the relationship. This is the \emph{inter-agent vector}.

Equation (\ref{eq:InterAgentMovement1}) identifies the \emph{inter-agent vector} for agent $b$ with respect to its neighbours.

\begin{equation}\label{eq:InterAgentMovement1}
v(b) = v_{c}(b) + v_{r}(b)
\end{equation}

If a \emph{weighted model} is used then equation~\ref{eq:InterAgentMovement1} becomes equation~\ref{eq:InterAgentMovement2}.

\begin{equation}\label{eq:InterAgentMovement2}
v(b) = k_cv_{c}(b) + k_rv_{r}(b)
\end{equation}

where $k_c$ and $k_r$ are weightings to change the effect of each vector as shown in Table~\ref{tab:MetricPhysics1}.

%% From here throughout sections \ref{metric:MagnitudeDynamics2} and \ref{metric:StabilityNullVector} imagine agent $b$ has just a single neighbour $b'$ and consider the effect of $b'$ on $v(b)$, the \emph{inter-agent vector} of $b$.
%% The relationship between the agents is that: \emph{the vector for $b$ with respect to $b^{'}$, and the vector for $b^{'}$ with respect to $b$ are on the same line}~(\ref{methods:Stability5}).
%% If we consider an agent ($b$) as having only one neighbour ($b'$) then using the formulae defined in chapter~\ref{chapter:methods} the vectors $v_c(b)$~(\ref{eq:FlyToCentre1}) and $v_r(b)$~(\ref{eq:Repulsion1}) are the magnitudes of the vectors that are generated by the field effects between the two agents. $k_c$ and $k_r$ are the weighting factors for the cohesion and repulsion. The resulting vectors are aligned along the same line of separation between the agents.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=7cm]{figures/Stability5}
{Vectors on line of separation.\label{methods:Stability5}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=7cm]{figures/Stability5}
%% \end{center}
%% \caption{Vectors on line of separation} \label{methods:Stability5}
%% \end{figure}

\section{Modelling movement}\label{sec:Movement1}
Each agent within a swarm calculates its \textit{movement-direction vector} based on its \textit{interaction} and \textit{destination} vectors. The \textit{movement vector}~($b_{pos}$) is calculated using the unit \textit{movement-direction vector} of Equation~\ref{eq:InterAgentMovement2} multiplied by the time elapsed~($t$) in the system and the speed characteristic of the agent~($s_b$).

This process is carried out for every agent in the swarm to generate the swarm's next position.

\begin{center}
\begin{equation}
\label{eq:AgentMovement}
b_{pos}=s_{b}t\big(v(b)\big)\string^ 
\end{equation}
\end{center}

The increment in the location of agent $b$ over time interval $t$ is shown in Equation~\ref{eq:AgentMovement} where $s_b$ is the speed of agent~$b$.

\section{Swarm movement analysis\label{metric:MagnitudeDynamics2}}
The repulsion and cohesion vectors are generated for an agent through the interaction of their field effects~(Figure~\ref{methods:FieldEffects}). There are a limited number of interactions that can occur; These are illustrated in Figures \ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}.

Following repeated experiments in the initial deployment of a swarm it was found that using the same parameters for a swarm's environmental parameters and the cohesion and repulsion ranges that the results where always very similar. The results shown in this paper are therefore an example set typical of the results obtained. The data extracts shown in Tables \ref{tab:SampleReplusion0}, \ref{tab:SampleReplusionPositive}, \ref{tab:SampleCohesionPositive} and  \ref{tab:SampleEquilibrium} are the simulation results that are produced by the parameters listed in Table~\ref{tab:MetricPhysics1}. The simulation consists of 200 agents over a 20 second period. The cohesion of an agent pair is shown as $k_cv_c$ and the repulsion as $k_rv_r$.

\begin{table}[H]
\begin{center}
\begin{tabular}{| p{2.5cm} | c | p{3cm} |}
\hline
\bf Weight \bf Component & \bf Setting & \bf Description \\ \hline
Sample Rate & 100 & ms - Unit sampling interval\\  \hline
$k_c$ & 5 & weight adjuster for cohesion bias\\  \hline
$k_r$ & 15 & weight adjuster for repulsion  bias\\  \hline
$k_d$ & 0 & weight adjuster for directional bias 0 for static baseline 100 from directional\\  \hline
Repulsion Boundary & 70 & units\\  \hline
Neighbour Distance & 80 & units\\  \hline
Speed & 20 & units/s\\  \hline
\end{tabular}\caption{Swarm model parameters} \label{tab:MetricPhysics1}
\end{center}
\end{table}

Figure~\ref{methods:Stability1} shows two agents within each other's cohesion fields but sufficiently distant to be outside of the repulsion fields. In this case $|k_cv_c| > 0$ and $|k_rv_r| = 0$: the result is the agent's resultant magnitude $v(b)$ causes the agent to move towards is neighbour $b'$. Likewise the neighbour's resultant magnitude $v(b')$ will cause it to move towards $b$. Table~\ref{tab:SampleReplusion0} shows the repulsion magnitude with a value of 0. The only influence on the agent pairs are cohesive vectors.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/Stability1}
{Internal movement cohesion (no repulsion).\label{methods:Stability5}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=5cm]{figures/Stability1}
%% \end{center}
%% \caption{Internal movement cohesion (no repulsion)} \label{methods:Stability1}
%% \end{figure}

\begin{table}[H]
\begin{center}
\begin{tabular}{| l | l | l | l | l | l |}
\hline
Log &	Id &	N.Id &	Distance &	{\color{green}Cohesion} &	{\color{red}Repulsion} 	\\ \hline
0 &	1 &	3 	 & 70.50359 &	{\color{green}352.51799} &	{\color{red}0} \\ \hline
0 &	1 &	100 & 71.78005 &	{\color{green}358.90027} &	{\color{red}0} \\ \hline
0 &	1 &	151 & 78.33995 &	{\color{green}391.69979} &	{\color{red}0} \\ \hline
0 &	2 &	99  &	72.04066 &	{\color{green}360.20334} &	{\color{red}0} \\ 
\hline
\end{tabular}\caption{Data extract ($|k_rv_r| = 0$)} \label{tab:SampleReplusion0}
\end{center}
\end{table}

Figure~\ref{methods:Stability2} shows two agents close together with repulsion dominating cohesion: $|k_cv_c|~<~|k_rv_r|$. The resultant vector directs the agents away from each other.

Table~\ref{tab:SampleReplusionPositive} shows the repulsion magnitude with a value greater than the cohesion magnitude.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/Stability2}
{Internal movement repulsion.\label{methods:Stability2}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=5cm]{figures/Stability2}
%% \end{center}
%% \caption{Internal movement repulsion} \label{methods:Stability2}
%% \end{figure}

\begin{table}[H]
\begin{center}
\begin{tabular}{| l | l | l | l | l | l |}
\hline
Log &	Id &	N.Id &	Distance &	{\color{green}Cohesion} &	{\color{red}Repulsion} 	\\ \hline
0 & 1 & 2 & 28.32522 & {\color{green}141.62612} & {\color{red}1544.86014} \\ \hline
0 & 1 & 6 & 41.48517 & {\color{green}207.42586} & {\color{red}721.71736} \\ \hline
0 & 1 & 7 & 35.26412 & {\color{green}176.32064} & {\color{red}1034.27101} \\ \hline
0 & 1 & 8 & 43.54503 & {\color{green}217.72518} & {\color{red}637.90759} \\
\hline
\end{tabular}\caption{Data extract ($|k_cv_c| < |k_rv_r|$)} \label{tab:SampleReplusionPositive}
\end{center}
\end{table}

Figure~\ref{methods:Stability3} shows two agents close together but with cohesion vector magnitudes greater than the repulsion magnitudes $|k_cv_c|~>~|k_rv_r|$. The resultant vector draws the agents together. The magnitude of the resultant cohesion vector reduces due to the cancelling effect of the repulsion vector. Table~\ref{tab:SampleCohesionPositive} shows a data extract with the cohesion magnitude greater than repulsion.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/Stability3}
{Internal movement cohesion and repulsion.\label{methods:Stability3}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=5cm]{figures/Stability3}
%% \end{center}
%% \caption{Internal movement cohesion and repulsion} \label{methods:Stability3}
%% \end{figure}

\begin{table}[H]
\begin{center}
\begin{tabular}{| l | l | l | l | l | l |}
\hline
Log &	Id &	N.Id &	Distance &	{\color{green}Cohesion} & {\color{red}Repulsion} 	\\ \hline
0 & 1 & 5 &	64.17214 &	{\color{green}320.86072} &	{\color{red}95.35676} \\ \hline
0 & 1 & 9 &	63.88049 &	{\color{green}319.40248} &	{\color{red}100.58590} \\ \hline
0 & 1 & 95 & 65.61522 &	{\color{green}328.07613} &	{\color{red}70.16681} \\ \hline
0 & 1 & 152 & 63.10700 & {\color{green}315.5350} & {\color{red}114.68844} \\ 
\hline
\end{tabular}\caption{Data extract ($|k_cv_c| > |k_rv_r|$)} \label{tab:SampleCohesionPositive}
\end{center}
\end{table}

Figure~\ref{methods:Stability4} shows two agents close together with $|k_cv_c|~=~|k_rv_r|$ the resultant vector is a \emph{null vector} and the agents have no overall influence upon each other due to the magnitude of the resultant vector being zero. Table~\ref{tab:SampleEquilibrium} is an extract of data from the simulator. The data shows near equilibrium but due to the dynamic nature of a swarm system no agents meet the condition fully.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/Stability4}
{Internal movement equilibrium.\label{methods:Stability4}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=5cm]{figures/Stability4}
%% \end{center}
%% \caption{Internal movement equilibrium} \label{methods:Stability4}
%% \end{figure}

\begin{table}[H]
\begin{center}
\begin{tabular}{| l | l | l | l | l | l |}
\hline
Log &	Id &	N.Id &	Distance &	{\color{green}Cohesion} &	{\color{red}Repulsion} 	\\ \hline
7 & 76 &	91 & 55.39031 & {\color{green}276.95156} & {\color{red}276.94684} \\ \hline
24 & 75 & 6 & 55.39032 & {\color{green}276.95160} & {\color{red}276.94661} \\ \hline
32 & 72 & 38 &	55.39002 & {\color{green}276.95013} & {\color{red}276.95370} \\ \hline
35 & 63 & 64 & 55.39022 &	{\color{green}276.95113} &	{\color{red}276.94887} \\
\hline
\end{tabular}\caption{Data extract ($|k_cv_c| \approx |k_rv_r|$)} \label{tab:SampleEquilibrium}
\end{center}
\end{table}

\section{Internal movement and the null vector}\label{metric:StabilityNullVector}
When the two vectors (cohesion and repulsion) have magnitudes that are equal and opposite they produce a null vector, indicating that two agents are optimally spaced for a given set of conditions. Although the agents are at an optimum position when resultants are zero it does not mean the swarm is optimally distributed. If a swarm is in a confined space it is possible for an optimum position to be created where the vector magnitude is affected by a compression effect. This phenomenon is used in the identification of the emergent behaviour of \emph{area flooding}.

If we consider the equilibrium state (Figure~\ref{methods:Stability4}) the resultant $v(b)=(0,0)$. A null vector cannot be normalised to produce a directional vector ($\hat{v} = \frac{v}{|v|}$ if $v\neq0$; $0$ if $v=0$) The effect is that the agent will remain stationary. If all agent pairs are in this condition the swarm will stop moving~(Figure~\ref{methods:StabilityNullVector}).

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/StabilityNullVector}
{Equilibrium with null vectors.\label{methods:StabilityNullVector}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=5cm]{figures/StabilityNullVector}
%% \end{center}
%% \caption{Equilibrium with null vectors} \label{methods:StabilityNullVector}
%% \end{figure}

Due to the movement of agents being based only on neighbour interactions this situation is very rare. The residual motion that persists in a swarm is the background `noise' or `jitter' that an algorithm creates.

If a swarm is goal-based, the addition of a \emph{directional vector} will prevent all agents simultaneously producing null vectors~(Figure~\ref{concave:VoidPerimeter2}, Equation~\ref{eq:InterAgentMovementDirectional1}). Here, $k_dv_{d}(b)$ is the calculated directional vector to be applied to an agent.

\begin{equation}\label{eq:InterAgentMovementDirectional1}
v(b) = k_cv_{c}(b) + k_rv_{r}(b) + k_dv_{d}(b)
\end{equation}

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5.5cm]{figures/StabilityNullVector3}
{(t).\label{concave:VoidPerimeter1}}

%% \begin{figure}[H]
%% \begin{center}
%% 	 \includegraphics[width=5.5cm]{figures/StabilityNullVector3}
%% \end{center}
%% \label{concave:VoidPerimeter1}
%% \caption{(t)}
%% \end{figure}

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/StabilityNullVector2}
{(t + 1).\label{concave:VoidPerimeter2}}

%% \begin{figure}[H]
%% \begin{center}
%% 	 \includegraphics[width=6cm]{figures/StabilityNullVector2}
%% \end{center}
%% \caption{(t + 1)}
%% \label{concave:VoidPerimeter2}
%% \end{figure}

\section{Residual internal movement (Jitter)}\label{metric:Jitter}
Due to the dynamic nature of a swarm, maintaining optimum internal movement as in~Figure \ref{methods:StabilityNullVector} is highly unlikely. The agent pairs may fluctuate between the 4 states~(Figures~\ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}); These transitions are \emph{jitter}. The degree to which this variation occurs can be measured using either the change in distance between the agents, or the change in the magnitude $P(b)$ (Equation~\ref{eq:BotStabilityT}) between the agents. Jitter arises as motion to maintain the structure of a swarm. A swarm coordination algorithm that reduces jitter is generally desirable.

\section{Distance metric\label{section:DistanceDynamics}}
The distance metric considers a swarm in terms of how the agents are physically distributed: i.e. only the inter-agent distances and their standard deviation are considered. The standard deviation indicates the extent to which the swarm is out of balance and will move to re-balance itself. If the standard deviation is zero then all the agents are evenly spaced.

%The distance metric does not take into consideration the vector magnitudes between the agents as discussed above. The metric therefore is unable to identify the potential state of the swarm in terms of its cohesive or repulsive state.

Navarro and Fernando describe a \emph{mean distance error} metric that is based on the variations in distances between inter-agent spaces~\cite{NIM:09}. This is the same as the standard deviation of the distance based internal movement metric.

%% \section{Distance metric internal movement}
%% The distance based internal movement is measured by identifying the mean length of the vectors between an agent and its neighbours. As with the \emph{agent resultant magnitude} a coordination algorithm produces `jitter' which is the variations from the mean. In the case of the distance based metric the jitter is identified by the changes in the distances rather than the changes in vector magnitude (\emph{agent resultant magnitude}). The distance metric is the mean and the standard deviation `jitter' of the inter-agent distances.

\section{Calculating distance internal movement}
Equation~\ref{eq:SwarmStabilityDistance1} computes the mean distance of an agent to its neighbours $nbr(b)$. 
%% The relative position vector generated for an agent $b$ to its neighbour $b'$, $bb'$, is shown in~(\ref{eq:FlyToCentre1}). The magnitude of that vector gives the distance between two agents. For an individual agent the average magnitude $\mu_d(b)$ is calculated as \ref{eq:SwarmStabilityDistance1} where $b$ is the agent and $|nbr(b)|$ is the number of neighbours.

\begin{equation}
\label{eq:SwarmStabilityDistance1}
\mu_d(b) = \frac{1}{\mathlarger{|nbr(b)|}}\left({\mathlarger{\sum_{b' \in nbr(b)}}|bb'|}\right)
\end{equation}

The mean distance for a swarm is calculated by Equation~\ref{eq:SwarmStabilityDistance2}. All the inter-agent distances are included for the swarm ($S$).

%% $\sum_{b \in S}|nbr(b)|$ calculates how many inter-agent relationships exist in the swarm and $\sum_{b' \in nbr(b)}|bb'|$ calculates the total distance between each agent and its neighbours. $\sum_{b \in S}$ iterates over all the agents in the swarm~($S$).

\begin{equation}
\label{eq:SwarmStabilityDistance2}
\mu_d(S) = \frac{\mathlarger{\sum_{b \in S}}~\mathlarger{\sum_{b' \in nbr(b)}}|bb'|}{\mathlarger{\sum_{b \in S}|nbr(b)|}}
\end{equation}

\section{Standard deviation in distance metric}\label{Section:VarianceInDistance}
The mechanism above provides an overall indication of the distribution of the agents. The model for a swarm $S$ ($\mu_d(S)$) is not sufficient to give an indication of the internal distribution of the agents. The standard deviation clarifies the distribution within the swarm as shown in~Equation~\ref{eq:SwarmStabilityDistance3}.

%% $(|bb'| - \mu(S))^2$ is the square of the difference in a distance to the mean and $\sum_{b \in S}~\sum_{b' \in nbr(b)}$ calculates the number of inter-agent interactions.

\begin{equation}
\label{eq:SwarmStabilityDistance3}
\sigma_d(S) = \sqrt{\frac{\mathlarger{\sum_{b \in S}}~\mathlarger{\sum_{b' \in nbr(b)}}\Big(|bb'| - \mu_d(S)\Big)^2}{\mathlarger{\sum_{b \in S}|nbr(b)|}}}
\end{equation}

The distance metric for the internal distribution of the agents is the pair consisting of $\mu_d(S)$, $\sigma_d(S)$. This can be written informally as:

\begin{equation}
\label{eq:SwarmPotentialMagnitude}
\psi_d(S) = \mu_d(S)\pm \sigma_d(S)
\end{equation}

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=14cm]{figures/BaselineDistance1}
{Baseline internal movement - distance.\label{coord:BaselineDistance1}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=9cm]{figures/BaselineDistance1}
%% \end{center}
%% \caption{Baseline internal movement - distance\label{coord:BaselineDistance1}}
%% \end{figure}

\section{Cohesion/repulsion metric\label{section:MagnitudeDynamics}}
An agent's \emph{resultant magnitude} is identified by Gazi and Passino~\cite{GP:11} and Barnes et al~\cite{BFV:07} as a `resultant characteristic' of a swarm. The sum of the cohesion and repulsion vectors between an agent $b$ and a neighbour $b'$ produce a resultant vector.

The cohesion/repulsion internal movement (\emph{cohesion/repulsion metric}) is measured by identifying the balance between the cohesion and repulsion vectors between agents as shown in Equation~\ref{eq:CohesionRepulsion}. This is derived from the \emph{inter-agent cohesion/repulsion vector} as shown in Equation~\ref{eq:InterAgentMovementDirectional1}. `Jitter' in the case of the agent resultant \emph{cohesion/repulsion metric} is measured as the deviation created by the inter-agent relationships. The identification of this deviation produces the clarifying component of the agent \emph{cohesion/repulsion metric}.
%% The magnitude based metric identifies the resultant internal magnitude (vector magnitude) through the cohesive and repulsive magnitudes that exist between each agent pair in a swarm. 
An agent's resultant vector magnitude is used as a scalar values to give an overall `size' to the agent relationship. The resultant scalar value is then modified based upon the `size' of the two relationship magnitudes.

This paper will refer to the scalarised resultant magnitude as the \emph{`cohesion/repulsion metric'} of the relationship. The `state' of a swarm is therefore the effect the environmental constraints and algorithms have upon the agent resultant magnitude. The agent's \emph{`cohesion/repulsion metric'} can be considered part of the `quality' measure for a swarm's performance.

The resultant \emph{cohesion/repulsion metric}, which is a scalar, is generated by comparing the length of the cohesion and repulsion vector magnitudes. If the cohesion magnitude is larger than the repulsion vector magnitude the resultant length is considered to be positive. If the repulsion vector magnitude is larger than the cohesion vector magnitude the resultant length is considered to be negative.

The \emph{`cohesion/repulsion metric'} is therefore the scalar value of the \emph{`agent resultant magnitude'} which is altered based upon the difference between the repulsion and cohesion magnitudes~(Equation~\ref{eq:CohesionRepulsion}).

If an agent's {cohesion/repulsion metric} is negative the swarm's bias is to expand as the repulsion is greater than cohesion. This is seen in the disorganised stage of a swarm~(Figure~\ref{coord:BaselineMagnitude1}). The disorganised stage is the initial random deployment of the agents. If the agent \emph{agent resultant magnitude} is positive then the swarm is exhibiting a tendency to contract and this indicates the swarm is a cohesive entity. This could also be described as the swarm being `sticky' as the agents bias is to `pull' towards each other.

The \emph{cohesion/repulsion metric} on its own does not give a complete measure of a swarm's internal state. There needs to be a qualifying component to the metric that identifies the degree of deviation in the resultant magnitude, this is the \emph{jitter}. The smaller the deviation the more uniform the structure of the swarm. These two components identify the degree to which a swarm has progressed towards a stable state.

The \emph{cohesion/repulsion metric} provides a view of the swarm's state through the balance between the repulsive and the cohesive vectors that are being applied to each agent. The deviation component identifies the degree to which the swarm has stabilised. The ideal status for inter-agent interactions would be for the agents to have a (\emph{cohesion/repulsion metric}) of zero or above. This would indicate that the agents are distributed such that they are at their distribution limit (outer most range of the cohesion field) or at a level that causes the agents to `pull' together. The ideal degree of deviation is zero as this indicates an even distribution of agents. These two aspects of a swarm's features are not considered by Gazi and Passino~\cite{GP:11} or Barnes et al~\cite{BFV:07} as a means of quantifying the structure of a swarm in terms of stability.

\section{Cohesion/repulsion internal movement model}\label{Section:StabilityModel}
Section~\ref{Section:StabilityMagnitude}~discussed calculation of resultant cohesion~(Equation \ref{eq:FlyToCentre1}) and resultant repulsion~(Equation \ref{eq:Repulsion1}) vectors for each agent deriving a combined \emph{cohesion/repulsion vector} via equation~\ref{eq:InterAgentMovement2}.

%% This value represents the overall `potential' of an agent. This vector when normalised produces a component of the \emph{movement-destination vector} for a swarm. If the total cohesion/repulsion vector is zero (null vector) then the agent will not move. $M(b)$ is the total \emph{cohesion/repulsion vector} for agent $b$ defined by:
%% 
%% \begin{equation}
%% \label{eq:BotStabilityT}
%% M(b) = k_cv_c(b) + k_rv_r(b)
%% \end{equation}

Now, we define for each agent $b$ a \emph{cohesion/repulsion scalar}, $P(b)$, to measure the influence of the neighbours of $b$ on $b$ as cohesive (positive) or repulsive (negative); this is Equation~\ref{eq:CohesionRepulsion}.

\begin{equation}
\label{eq:CohesionRepulsion}
P(b) = \left\{\begin{array}{lll}
               |v(b)|& if & |k_cv_c(b)| > |k_r v_r(b)|\\
              -|v(b)|& otherwise
              \end{array}\right.
\end{equation}

Although it is possible for $v(b)$ to be null there could still be variation in the constituent components. The variation calculation (standard deviation) is shown below.

%% Equation~\ref{eq:SwarmStabilityMetricT} is the mean of the cohesion/repulsion scalar for an agent.
%% 
%% \begin{equation}
%% \label{eq:SwarmStabilityMetricT}
%% \mu_p(b) = \frac{P(b)}{|nbr(b)|}
%% \end{equation}

%% \ref{eq:SwarmStabilityMetricT} should be considered as being applied at discrete points in time within a simulation. The formulae could also be shown based on time $t$ (\ref{eq:SwarmStabilityMetricT2}) where $b_t$ is an agent at a specific time interval. All formulae for agent resultant magnitude in this thesis will be considered as being applied at a point in time and will therefore be shown without a $t$ subscript.
%% 
%% \begin{equation}
%% \label{eq:SwarmStabilityMetricT2}
%% \mu_p(b_{t}) = \frac{P(b_{t})}{|nbr(b_{t})|}
%% \end{equation}

The mean cohesion/repulsion scalar for the swarm is now given by~Equation~\ref{eq:SwarmStabilityMetricT3}.

\begin{equation}
\label{eq:SwarmStabilityMetricT3}
\mu_p(S) = \frac{\mathlarger{\sum_{b \in S} P(b)}}{\mathlarger{\sum_{b \in S}}|nbr(b)| + \mathlarger{\sum_{b \in S}}|rep(b)|}
\end{equation}

%% \section{Standard deviation in agent cohesion/repulsion metric}\label{Section:VarianceInPotential}
%% To improve the metric clarification in terms of the deviation from the \emph{cohesion/repulsion metric} norm is calculated. The standard deviation of the entire swarm  is shown in Equation \ref{eq:SwarmStabilityMetricT}.

The standard deviation associated with this mean is calculated as Equation~\ref{eq:SwarmStabilityQuotientT}.
%%  where $\sigma_p(S)$ is the standard deviation and $\mu_p(S)$ is the mean at the same point in time.

\begin{equation}
\label{eq:SwarmStabilityQuotientT}
\sigma_p(S) = \sqrt{\frac{\mathlarger{\sum_{b \in S}}\Big(P(b)-\mu_p(S)\Big)^2}{\mathlarger{\sum_{b \in S}}|nbr(b)| + \mathlarger{\sum_{b \in S}}|rep(b)|}}
\end{equation}

The metric for the internal movement is this pair of numbers, the mean and standard deviation of the swarm's internal \emph{cohesion/repulsion}. The pair $\mu_p(S)$, $\sigma_p(S)$ may be written informally as:

\begin{equation}
\label{eq:SwarmPotentialMagnitude}
\psi_p(S) = \mu_p(S)\pm \sigma_p(S)
\end{equation}

Figure~\ref{coord:BaselineMagnitude1} shows the evolution of the metric in time.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=14cm]{figures/BaselineMagnitude1}
{Baseline internal movement - magnitude.\label{coord:BaselineMagnitude1}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=9cm]{figures/BaselineMagnitude1}
%% \end{center}
%% \caption{Baseline internal movement - magnitude\label{coord:BaselineMagnitude1}}
%% \end{figure}

\section{Conclusion - metric comparison\label{metric:MagnitudeDistanceComparison}}
The two metrics (distance and \emph{cohesion/repulsion}) both measure the structure of a swarm in terms of the distribution of the agents and deviations from a mean. The \emph{cohesion/repulsion} metric also allows the state of the swarm to be identified.

The distance metric is based upon the physical \emph{distribution} of the agents and the cohesion/repulsion metric is based upon the logical \emph{interaction} of the agents.

The distance based metric provides an analysis of the distribution of the agents at a point in time and allows the agitation of the swarm to be assessed without considering the possible distribution of agents that the field effects \emph{could} produce.

The \emph{cohesion/repulsion} metric provides an indication of a swarm's potential movement. This is independent of the physical distribution of the agents within the swarm. The lack of dependence on the physical distribution allows the metric to be used in heterogeneous field effect swarms~\ref{additional:fieldsWork} where the physical distribution of agents may vary due to the varied field effect ranges.

In a homogeneous field effect swarm the two metrics provide a deeper evaluation of a swarm's structure and `potential'. Consider the following: the repulsion field is increased but the internal distances do not change as a result the \emph{cohesion/repulsion metric} rises: This indicates `something' is confining the swarm's distribution and the swarm is filling a confined area. This analysis could be used in identifying effective swarm distribution for the coverage of a sensor array as discussed by Ramaithitima et al.~\cite{RWBK:15}

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FieldEffects2}
{Complex swarm interactions.\label{additional:fieldsWork}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=6cm]{figures/FieldEffects2}
%% \end{center}
%% \caption{Complex swarm interactions\label{additional:fieldsWork}}
%% \end{figure}

\section{Future work\label{metric:Future}}
The detection of a swarm's perimeter \cite{MD:09, MJ:08, ZAPS:07, JG:13} identifies a subset of a swarm which can be used for coordination. The use of the perimeter detection allows a swarm to be used in many different ways~\cite{ZFG:13, AKK:08, APZDAMC:09, AZDPS:11}.

A possible adaptation to both of these areas of research would be to use the swarm perimeter detection to identify a subset of agents then reducing the field effect of those agents to increase the agent density at the perimeter. Increasing the density would allow increased coverage of any sensor arrays that the swarm was using at the swarm edge. The issue with introducing heterogeneous field effects is that existing coordination techniques are required. The cohesion/repulsion metric discussed in this paper would still work within this environment as a tool to identify swarm stability. this is due to the metric being based upon inter-agent effects rather than physical distribution.

\bibliographystyle{plain}
\bibliography{thesis}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Images/NEGrey}}]{Neil Eliot}
is a Senior Lecturer at Northumbria University.  Neil was awarded his Ph.D. from Northumbria in 2017, his Pg.C. in 1994, and his B.Sc. in 1989 all in Computing. His research areas include swarm theory and cybersecurity. Neil is a Certified Ethical Hacker (CEH) and coordinator for the EC Council Academic alliance at Northumbria. Prior to this Neil worked in the chemical industry and the NHS. His research interests are in the area of network cybersecurity, smart home technologies and swarm theory. 
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Images/DKGrey}}]{David Kendall}
has been a Senior Lecturer in Computing at Northumbria University since 1989, except for a brief period as a Lecturer in Computer Science at Durham University (2001-02). Previously, he was a Research Associate at Newcastle University (1987-89), following experience as a Senior Software Engineer in industry (1983-86). He has an M.A. degree in Literae Humaniores from Oxford University, where he studied at New College. He received M.Sc. and Ph.D. degrees in Computing Science from Newcastle University. His research interests are in the area of formal modelling and analysis of embedded systems. 
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Images/MBGrey}}]{Michael Brockway}
has been a Senior Lecturer at Northumbria University for 17 years. His first degree is in pure and applied mathematics (1973-4), his Masters in mathematics (mathematical logic, category theory, 1975), and he has a PhD in computer science (formal methods for distributed embedded systems, 2010) from Northumbria University. Before that Michael worked in teaching and the computing industry.
\end{IEEEbiography}

\EOD

\end{document}

### 2018-02-22 paper.tex
Fixed Refs

@@
 The repulsion and cohesion vectors are generated for an agent through the interaction of their field effects~(Figure~\ref{methods:FieldEffects}).
-There are a limited number of interactions that can occur; These are illustrated in Figures \ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}.
+There are a limited number of interactions that can occur; These are illustrated in Figures \ref{methods:Stability2}, \ref{methods:Stability3}, \ref{methods:Stability4} and \ref{methods:StabilityNullVector}.
 

### 2018-02-26 paper.tex
Compliance Changes

@@
 \uppercase{David Kendall\authorrefmark{1}, and Michael Brockway\authorrefmark{1}}}
-\address[1]{Northumbria University, Department of Computing and Information Sciences, Newcastle upon Tyne, NE1 8ST}
+\address[1]{Northumbria University, Department of Computing and Information Sciences, 
+Newcastle upon Tyne, NE1 8ST}
 
@@
 swarm as expanding or cohesive.  This additional information has practical
-benefits for the implementation of algorithms for heterogeneous swarms.
+benefits for the implementation of algorithms in heterogeneous field-effect swarms.
 \end{abstract}
@@
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/FieldEffects}
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FieldEffects}
 {Agent field effects.\label{methods:FieldEffects}}
@@
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Images/NEGrey}}]{Neil Eliot}
-is a Senior Lecturer at Northumbria University.
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Images/eliot}}]{Neil Eliot}
+has been a Senior Lecturer at Northumbria University since 1998.
 Neil was awarded his Ph.D.
@@
 His research areas include swarm theory and cybersecurity.
-Neil is a Certified Ethical Hacker (CEH) and coordinator for the EC Council Academic alliance at Northumbria.
 Prior to this Neil worked in the chemical industry and the NHS.
@@
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Images/DKGrey}}]{David Kendall}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Images/kenda}}]{David Kendall}
 has been a Senior Lecturer in Computing at Northumbria University since 1989, except for a brief period as a Lecturer in Computer Science at Durham University (2001-02).
@@
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Images/MBGrey}}]{Michael Brockway}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Images/brock}}]{Michael Brockway}
 has been a Senior Lecturer at Northumbria University for 17 years.

### 2018-02-27 paper.tex
Future Work Update

@@
 Increasing the density would allow increased coverage of any sensor arrays that the swarm was using at the swarm edge.
-The issue with introducing heterogeneous field effects is that existing coordination techniques are required.
+The issue with introducing heterogeneous field effects is that existing distance based metric would not allow the swarm's structure to be analysed correctly.
 The cohesion/repulsion metric discussed in this paper would still work within this environment as a tool to identify swarm stability.
-this is due to the metric being based upon inter-agent effects rather than physical distribution.
+This is due to the metric being based upon inter-agent effects rather than physical distribution.
 

### 2018-03-09 paper.tex
Minor Updates

@@
 \Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/Stability1}
-{Internal movement cohesion (no repulsion).\label{methods:Stability5}}
+{Internal movement cohesion (no repulsion).\label{methods:Stability1}}
 
@@
 
-\section{Internal movement and the null vector}\label{metric:StabilityNullVector}
+\section{Internal movement and the null vector}\label{section:StabilityNullVector}
 When the two vectors (cohesion and repulsion) have magnitudes that are equal and opposite they produce a null vector, indicating that two agents are optimally spaced for a given set of conditions.
@@
 The agent pairs may fluctuate between the 4 states~(Figures~\ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}); These transitions are \emph{jitter}.
-The degree to which this variation occurs can be measured using either the change in distance between the agents, or the change in the magnitude $P(b)$ (Equation~\ref{eq:BotStabilityT}) between the agents.
+The degree to which this variation occurs can be measured using either the change in distance between the agents, or the change in the magnitude $P(b)$ (Equation~\ref{eq:CohesionRepulsion}) between the agents.
 Jitter arises as motion to maintain the structure of a swarm.
@@
 
-The distance metric for the internal distribution of the agents is the pair consisting of $\mu_d(S)$, $\sigma_d(S)$. This can be written informally as:
+The distance-based metric for the internal distribution of the agents is the pair consisting of $\mu_d(S)$, $\sigma_d(S)$. This can be written informally as:
 
 \begin{equation}
-\label{eq:SwarmPotentialMagnitude}
+\label{eq:SwarmDistanceMetric}
 \psi_d(S) = \mu_d(S)\pm \sigma_d(S)
@@
 
-\section{Conclusion - metric comparison\label{metric:MagnitudeDistanceComparison}}
-The two metrics (distance and \emph{cohesion/repulsion}) both measure the structure of a swarm in terms of the distribution of the agents and deviations from a mean.
+\section{Conclusion\label{metric:MagnitudeDistanceComparison}}
+%% The two metrics (distance and \emph{cohesion/repulsion}) both measure the structure of a swarm in terms of the distribution of the agents and deviations from a mean.
 The \emph{cohesion/repulsion} metric also allows the state of the swarm to be identified.
+%% 
+%% The distance metric is based upon the physical \emph{distribution} of the agents and the cohesion/repulsion metric is based upon the logical \emph{interaction} of the agents.
+The distance-based metric measures swarm structure in terms of position of agents.
 
-The distance metric is based upon the physical \emph{distribution} of the agents and the cohesion/repulsion metric is based upon the logical \emph{interaction} of the agents.
-
-The distance based metric provides an analysis of the distribution of the agents at a point in time and allows the agitation of the swarm to be assessed without considering the possible distribution of agents that the field effects \emph{could} produce.
+The distance-based metric provides an analysis of the distribution of the agents at a point in time and allows the agitation of the swarm to be assessed.
+It does not take into account the possible distribution of agents that the field effects \emph{could} produce.
 
@@
 This is independent of the physical distribution of the agents within the swarm.
-The lack of dependence on the physical distribution allows the metric to be used in heterogeneous field effect swarms~\ref{additional:fieldsWork} where the physical distribution of agents may vary due to the varied field effect ranges.
+The lack of dependence on the physical distribution allows the metric to be used in heterogeneous field effect swarms~(Figure~\ref{additional:fieldsWork}) where the physical distribution of agents may vary due to the varied field effect ranges.
 
-In a homogeneous field effect swarm the two metrics provide a deeper evaluation of a swarm's structure and `potential'.
-Consider the following: the repulsion field is increased but the internal distances do not change as a result the \emph{cohesion/repulsion metric} rises: This indicates `something' is confining the swarm's distribution and the swarm is filling a confined area.
+In a homogeneous field effect swarm the two metrics together provide a deeper evaluation of a swarm's structure and `potential'.
+Consider the following: the repulsion field is increased but the internal distances do not change; as a result the \emph{cohesion/repulsion metric} rises: This indicates `something' is confining the swarm's distribution.
 This analysis could be used in identifying effective swarm distribution for the coverage of a sensor array as discussed by Ramaithitima et al.~\cite{RWBK:15}
@@
 
-A possible adaptation to both of these areas of research would be to use the swarm perimeter detection to identify a subset of agents then reducing the field effect of those agents to increase the agent density at the perimeter.
+A possible adaptation to this area of research is to use the swarm perimeter detection to identify a subset of agents then reduce the field effects (repulsion and cohesion) of those agents to increase the agent density at the perimeter.
 Increasing the density would allow increased coverage of any sensor arrays that the swarm was using at the swarm edge.
-The issue with introducing heterogeneous field effects is that existing distance based metric would not allow the swarm's structure to be analysed correctly.
-The cohesion/repulsion metric discussed in this paper would still work within this environment as a tool to identify swarm stability.
+The issue with introducing heterogeneous field effects is that the existing distance-based metric would not allow the swarm's structure to be analysed correctly.
+The cohesion/repulsion metric discussed in this paper would still work within this environment as a tool to identify a swarm's structure.
 This is due to the metric being based upon inter-agent effects rather than physical distribution.

### 2018-03-09 paper.tex
Minor Updates

@@
 \section{Inter-agent vector magnitude effect on internal movement}\label{Section:StabilityMagnitude}
-Figure \ref{methods:Stability5} shows the cohesion and repulsion vector contributions to $v_c(b)$ and $v_r(b)$ due to neighbour $b'$, as given in (Equations \ref{eq:FlyToCentre1} and \ref{eq:Repulsion1}).
+Figure \ref{methods:Stability5} shows the cohesion and repulsion vector contributions to $v_c(b)$ and $v_r(b)$ due to neighbour $b'$, as given in (\ref{eq:FlyToCentre1}) and (\ref{eq:Repulsion1}).
 Notice that the vectors are along the line of separation $bb'$.
@@
 
-Equation (\ref{eq:FlyToCentre1}) identifies the resultant cohesion effect of all the neighbours of $b$.
+Equation~(\ref{eq:FlyToCentre1}) identifies the resultant cohesion effect of all the neighbours of $b$.
 $nbr(b)$ defines the set of all the agents in the swarm that are a neighbour of $b$; a neighbour is any agent that fails within the cohesion field.
@@
 
-Equation (\ref{eq:Repulsion1}) identifies the resultant repulsion effect of all the neighbours of $b$ where $rep(.)$ returns a set of all agents that are within the repulsion field.
+Equation~(\ref{eq:Repulsion1}) identifies the resultant repulsion effect of all the neighbours of $b$ where $rep(.)$ returns a set of all agents that are within the repulsion field.
 
@@
 
-Equation (\ref{eq:InterAgentMovement1}) identifies the \emph{inter-agent vector} for agent $b$ with respect to its neighbours.
+Equation~(\ref{eq:InterAgentMovement1}) identifies the \emph{inter-agent vector} for agent $b$ with respect to its neighbours.
 
@@
 
-If a \emph{weighted model} is used then equation~\ref{eq:InterAgentMovement1} becomes equation~\ref{eq:InterAgentMovement2}.
+If a \emph{weighted} model is used then~(\ref{eq:InterAgentMovement1}) becomes~(\ref{eq:InterAgentMovement2}).
 
@@
 Each agent within a swarm calculates its \textit{movement-direction vector} based on its \textit{interaction} and \textit{destination} vectors.
-The \textit{movement vector}~($b_{pos}$) is calculated using the unit \textit{movement-direction vector} of Equation~\ref{eq:InterAgentMovement2} multiplied by the time elapsed~($t$) in the system and the speed characteristic of the agent~($s_b$).
+The \textit{movement vector}~($b_{pos}$) is calculated using the unit \textit{movement-direction vector} of (\ref{eq:InterAgentMovement2}) multiplied by the time elapsed~($t$) in the system and the speed characteristic of the agent~($s_b$).
 
@@
 
-The increment in the location of agent $b$ over time interval $t$ is shown in Equation~\ref{eq:AgentMovement} where $s_b$ is the speed of agent~$b$.
+The increment in the location of agent $b$ over time interval $t$ is shown in~(\ref{eq:AgentMovement}) where $s_b$ is the speed of agent~$b$.
 
@@
 
-If a swarm is goal-based, the addition of a \emph{directional vector} will prevent all agents simultaneously producing null vectors~(Figure~\ref{concave:VoidPerimeter2}, Equation~\ref{eq:InterAgentMovementDirectional1}).
+If a swarm is goal-based, the addition of a \emph{directional vector} will prevent all agents simultaneously producing null vectors as shown in~(Figure~\ref{concave:VoidPerimeter2}, Equation~(\ref{eq:InterAgentMovementDirectional1})).
 Here, $k_dv_{d}(b)$ is the calculated directional vector to be applied to an agent.
@@
 The agent pairs may fluctuate between the 4 states~(Figures~\ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}); These transitions are \emph{jitter}.
-The degree to which this variation occurs can be measured using either the change in distance between the agents, or the change in the magnitude $P(b)$ (Equation~\ref{eq:CohesionRepulsion}) between the agents.
+The degree to which this variation occurs can be measured using either the change in distance between the agents, or the change in the magnitude $P(b)$~(\ref{eq:CohesionRepulsion}) between the agents.
 Jitter arises as motion to maintain the structure of a swarm.
@@
 \section{Calculating distance internal movement}
-Equation~\ref{eq:SwarmStabilityDistance1} computes the mean distance of an agent to its neighbours $nbr(b)$. 
+Equation~(\ref{eq:SwarmStabilityDistance1}) computes the mean distance of an agent to its neighbours $nbr(b)$. 
 %% The relative position vector generated for an agent $b$ to its neighbour $b'$, $bb'$, is shown in~(\ref{eq:FlyToCentre1}).
@@
 
-The mean distance for a swarm is calculated by Equation~\ref{eq:SwarmStabilityDistance2}. All the inter-agent distances are included for the swarm ($S$).
+The mean distance for a swarm is calculated by~(\ref{eq:SwarmStabilityDistance2}). All the inter-agent distances are included for the swarm ($S$).
 
@@
 The model for a swarm $S$ ($\mu_d(S)$) is not sufficient to give an indication of the internal distribution of the agents.
-The standard deviation clarifies the distribution within the swarm as shown in~Equation~\ref{eq:SwarmStabilityDistance3}.
+The standard deviation clarifies the distribution within the swarm as shown in~(\ref{eq:SwarmStabilityDistance3}).
 
@@
 
-The cohesion/repulsion internal movement (\emph{cohesion/repulsion metric}) is measured by identifying the balance between the cohesion and repulsion vectors between agents as shown in Equation~\ref{eq:CohesionRepulsion}.
-This is derived from the \emph{inter-agent cohesion/repulsion vector} as shown in Equation~\ref{eq:InterAgentMovementDirectional1}.
+The cohesion/repulsion internal movement (\emph{cohesion/repulsion metric}) is measured by identifying the balance between the cohesion and repulsion vectors between agents as shown in~(\ref{eq:CohesionRepulsion}).
+This is derived from the \emph{inter-agent cohesion/repulsion vector} as shown in~(\ref{eq:InterAgentMovementDirectional1}).
 `Jitter' in the case of the agent resultant \emph{cohesion/repulsion metric} is measured as the deviation created by the inter-agent relationships.
@@
 
-The \emph{`cohesion/repulsion metric'} is therefore the scalar value of the \emph{`agent resultant magnitude'} which is altered based upon the difference between the repulsion and cohesion magnitudes~(Equation~\ref{eq:CohesionRepulsion}).
+The \emph{`cohesion/repulsion metric'} is therefore the scalar value of the \emph{`agent resultant magnitude'} which is altered based upon the difference between the repulsion and cohesion magnitudes~(\ref{eq:CohesionRepulsion}).
 
@@
 \section{Cohesion/repulsion internal movement model}\label{Section:StabilityModel}
-Section~\ref{Section:StabilityMagnitude}~discussed calculation of resultant cohesion~(Equation \ref{eq:FlyToCentre1}) and resultant repulsion~(Equation \ref{eq:Repulsion1}) vectors for each agent deriving a combined \emph{cohesion/repulsion vector} via equation~\ref{eq:InterAgentMovement2}.
+Section~\ref{Section:StabilityMagnitude}~discussed calculation of resultant cohesion~(\ref{eq:FlyToCentre1}) and resultant repulsion~(\ref{eq:Repulsion1}) vectors for each agent deriving a combined \emph{cohesion/repulsion vector} using~(\ref{eq:InterAgentMovement2}).
 
@@
 
-Now, we define for each agent $b$ a \emph{cohesion/repulsion scalar}, $P(b)$, to measure the influence of the neighbours of $b$ on $b$ as cohesive (positive) or repulsive (negative); this is Equation~\ref{eq:CohesionRepulsion}.
+Now, we define for each agent $b$ a \emph{cohesion/repulsion scalar}, $P(b)$, to measure the influence of the neighbours of $b$ on $b$ as cohesive (positive) or repulsive (negative); this is~(\ref{eq:CohesionRepulsion}).
 
@@
 
-The mean cohesion/repulsion scalar for the swarm is now given by~Equation~\ref{eq:SwarmStabilityMetricT3}.
+The mean cohesion/repulsion scalar for the swarm is now given by~(\ref{eq:SwarmStabilityMetricT3}).
 
@@
 
-The standard deviation associated with this mean is calculated as Equation~\ref{eq:SwarmStabilityQuotientT}.
+The standard deviation associated with this mean is calculated as~(\ref{eq:SwarmStabilityQuotientT}).
 %%  where $\sigma_p(S)$ is the standard deviation and $\mu_p(S)$ is the mean at the same point in time.

### 2018-03-29 paper.tex
Updated
Minors

@@
 creates.  In addition the \emph{cohesion/repulsion} metric also identifies a
-swarm as expanding or cohesive.  This additional information has practical
-benefits for the implementation of algorithms in heterogeneous field-effect swarms.
+swarm as expanding or cohesive.
 \end{abstract}
@@
 P(b) = \left\{\begin{array}{lll}
-               |v(b)|& if & |k_cv_c(b)| > |k_r v_r(b)|\\
-              -|v(b)|& otherwise
+               |v(b)|& \mathrm{if} & |k_cv_c(b)| > |k_r v_r(b)|\\
+              -|v(b)|& \mathrm{otherwise}
               \end{array}\right.
@@
 \Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=14cm]{figures/BaselineMagnitude1}
-{Baseline internal movement - magnitude.\label{coord:BaselineMagnitude1}}
+{Baseline internal movement - cohesion/repulsion.\label{coord:BaselineMagnitude1}}
 

### 2018-03-29 paper.tex
Minor

@@
 The \emph{distance metric} which uses variations in the inter-agent distances, as discussed by Navarro et al.~\cite{NIM:09}.
-The new metric uses the resultant magnitudes of the agents' \emph{inter-agent vectors} that are induced by an agent interacting with its neighbours which allows a more accurate analysis of a swarm when using non-homogeneous sensor distances.
+The new metric uses the resultant magnitudes of the agents' \emph{inter-agent vectors} that are defined by interactions of and agent with it's neighbours which allows a more accurate analysis of a swarm when using non-homogeneous sensor distances.
 The metric also identifies when a swarm is expanding or contracting.
@@
 These effects are \emph{cohesion}, to draw agents closer, and \emph{repulsion} to prevent agents colliding.
-\emph{Field effects} are the ranges around an agent that determine the effect other agents have upon its movement~(Figure~\ref{methods:FieldEffects}).
+\emph{Field effects} have ranges around an agent and determine the effect other agents have upon it's movement~(Figure~\ref{methods:FieldEffects}).
 It is usual for the cohesion field to have a radius $C_b$ which is larger than the repulsion radius $R_b$.
@@
 When the repulsion magnitude exceeds the cohesion magnitude the agent will have a tendency to move away from $b'$, i.e.
-repulsed.
+repelled.
 When an agent $b$ moves too close to an obstacle, i.e.
-within the obstacle repulsion range $O_b$, it has repulsion vector applied and the agent will tend towards moving away from the obstacle.
+within the obstacle repulsion range $O_b$, it has it's repulsion vector applied and the agent will tend towards moving away from the obstacle.
 
@@
 Equation~(\ref{eq:FlyToCentre1}) identifies the resultant cohesion effect of all the neighbours of $b$.
-$nbr(b)$ defines the set of all the agents in the swarm that are a neighbour of $b$; a neighbour is any agent that fails within the cohesion field.
+Here, $nbr(b)$ defines the set of all the agents in the swarm that are a neighbour of $b$; a neighbour is any agent that fails within the cohesion field.
 
@@
 The increment in the location of agent $b$ over time interval $t$ is shown in~(\ref{eq:AgentMovement}) where $s_b$ is the speed of agent~$b$.
+Note that \string^ is the equivalent of $\hat{v} = \frac{v}{|v|}$ the normalised vector.
 

### 2018-03-29 paper.tex
Ready for new section

@@
 
+\section{Application: Flood filling\label{metric:ApplicationFloodFilling}}
+
+Area flooding is a technique used to fill an enclosed area with agents such that they are distributed as `effectively' as possible throughout an area.
+An optimum distribution of agents is not necessarily an even distribution.
+The distribution of the agents is governed by the field effects and the interaction of agents with obstacles (\textit{interaction vectors}).
+
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FLOOD1}
+{Flood fill Start.\label{emerge:FLOOD1}}
+
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FLOOD6}
+{Flood fill End.\label{emerge:FLOOD6}}
+
+Filling an area can be applied to tasks that require an unknown environment to be analysed or surveyed.
+Consider a disaster area following a landslide or a building collapsing following an earthquake; the movement in the land and buildings will produce spaces that are unmapped.
+The unmapped areas may require investigation to locate people, resources or to create some form of sensor network to analyse the conditions within the area such as creating a heat map or to identify the location toxic gases.
+
+The concept of using a swarm to provide coverage over unknown areas is a current area of research.
+Alvissalim et al.
+(2012) discuss the application of commercially available drones to provide a communication infrastructure across an unknown disaster areas~\cite{AZHMJJM:12}.
+Scheutz and Bauer (2006) use both cohesion and repulsion as a mechanism to create coverage of an area that requires protection in an adversarial environment~\cite{SB:06}.
+
+The concept behind the field effect expansion is to increase both the repulsion and cohesion fields over a period of time.
+This increase in the field effects makes the agents increase the distance between each other expanding the swarm as a whole.
+The expansion increases until, due to boundary compression, the swarm is unable to expand further.
+In an extreme case the expansion will result in a set of field effects that create a mesh based swarm structure rather than the desired hexagonal structure that the field effects should create.
+Identifying either the change in modality or the inability to expand, which can be identified by the \textit{inter-agent vector magnitude} increase or a reduction in the inter-agent distance stability indicates area saturation has occurred.
+
+To test this hypothesis a swarm is modelled in the simulator.
+The model consists of an obstacle-based enclosed space and a swarm consisting of 60 agents.
+The experiments parameters for the simulation are shown in~table~\ref{tab:FillParameters1}
+
+\begin{table}[H]
+\begin{center}
+\begin{tabular}{| p{1.8cm} | p{1.5cm} | p{4.0cm} |}
+\hline
+\bf Parameter & \bf Value  & \bf Description \\ \hline
+$k_c$         & 5          & weight adjuster for cohesion bias\\ \hline
+$k_r$         & 60         & weight adjuster for repulsion bias\\ \hline
+Sample rate   & 100ms      & proximity sensor rate\\ \hline
+Speed         & 20 units/s & agent speed\\ \hline
+\end{tabular}\caption{Swarm parameters} \label{tab:FillParameters1}
+\end{center}
+\end{table}
+
+The field effects are incremented in turn, neighbour range followed by repulsion range.
+After each repulsion field change the swarm is allowed to redistribute itself and stabilise.
+Table~\ref{tab:emerge:BaselineConcaveReduction} shows the field effect settings that are used for the simulation.
+The field effect are selected so as to ensure the swarm parameters have the potential to create a hexagonal swarm.
+
+\begin{table}[H]
+\begin{center}
+\begin{tabular}{| p{1.8cm} | p{0.3cm} | p{0.3cm} | p{0.3cm} | p{0.3cm} | p{0.3cm} | p{0.3cm} | p{0.3cm} | p{0.7cm} |}
+\hline
+\bf Weight \bf component & \bf 1 & \bf 2 & \bf 3 & \bf 4 & \bf 5 & \bf 6 & \bf 7 & \\ \hline
+Repulsion Boundary & 50 & -  & 60 & -  & 70 & -  & 80 & units\\  \hline
+Neighbour Distance & 60 & 70 & -  & 80 & -  & 90 & -  & units\\  \hline
+\end{tabular}\caption{Field effect expansion sequence} \label{tab:FillSequence}
+\end{center}
+\end{table}
+
+Figure~\ref{emerge:FIELDFILL-MAG} shows the resultant magnitude for the swarm for the entire simulation.
+Between 100 seconds and 105 seconds there is a significant change in the magnitude where the value becomes negative.
+This indicates there is a compression effect on the swarm and it is unable to expand the inter-agent distances.
+
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=16cm]{figures/FIELDFILL-MAG}
+{Cohesion/Repulsion metric 0-120 seconds.\label{emerge:FIELDFILL-MAG}}
+
+% \begin{figure}[H]
+% \begin{center}
+% \includegraphics[width=12cm]{figures/FIELDFILL-MAG}
+% \end{center}
+% \caption{Magnitude metric 0-120 seconds\label{emerge:FIELDFILL-MAG}}
+% \end{figure}
+
+Figure~\ref{emerge:FIELDFILL-DIST}~shows the distance metric for the simulation.
+The initial deployment is shown at 0 seconds followed by incremental changes in the distances and variance of the distances as each increment is made in the field effects.
+The simulation terminates after 12 seconds.
+
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=16cm]{figures/FIELDFILL-DIST}
+{Distance metric 0-120 seconds.\label{emerge:FIELDFILL-DIST}}
+
+% \begin{figure}[H]
+% \begin{center}
+% \includegraphics[width=12cm]{figures/FIELDFILL-DIST}
+% \end{center}
+% \caption{Distance metric 0-120 seconds\label{emerge:FIELDFILL-DIST}}
+% \end{figure}
+
 \section{Conclusion\label{metric:MagnitudeDistanceComparison}}

### 2018-03-29 paper.tex
Minor Updates

@@
 The model consists of an obstacle-based enclosed space and a swarm consisting of 60 agents.
-The experiments parameters for the simulation are shown in~table~\ref{tab:FillParameters1}
+The experiments parameters for the simulation are shown in~Table~\ref{tab:FillParameters1}
 
@@
 After each repulsion field change the swarm is allowed to redistribute itself and stabilise.
-Table~\ref{tab:emerge:BaselineConcaveReduction} shows the field effect settings that are used for the simulation.
+Table~\ref{tab:FillSequence} shows the field effect settings that are used for the simulation.
 The field effect are selected so as to ensure the swarm parameters have the potential to create a hexagonal swarm.

### 2018-04-09 paper.tex
Graph Updates
New Section Added

@@
 
-\section{Application: Flood filling\label{metric:ApplicationFloodFilling}}
+\section{Application: Flood filling}\label{metric:ApplicationFloodFilling}
 
@@
 The distribution of the agents is governed by the field effects and the interaction of agents with obstacles (\textit{interaction vectors}).
-
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FLOOD1}
-{Flood fill Start.\label{emerge:FLOOD1}}
-
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FLOOD6}
-{Flood fill End.\label{emerge:FLOOD6}}
 
@@
 In an extreme case the expansion will result in a set of field effects that create a mesh based swarm structure rather than the desired hexagonal structure that the field effects should create.
-Identifying either the change in modality or the inability to expand, which can be identified by the \textit{inter-agent vector magnitude} increase or a reduction in the inter-agent distance stability indicates area saturation has occurred.
+Identifying the \textit{cohesion/repulsion metric} falling below zero indicates area saturation has occurred.
 
@@
 
-The field effects are incremented in turn, neighbour range followed by repulsion range.
+The field effects are incremented in turn, cohesion range followed by repulsion range.
 After each repulsion field change the swarm is allowed to redistribute itself and stabilise.
 Table~\ref{tab:FillSequence} shows the field effect settings that are used for the simulation.
-The field effect are selected so as to ensure the swarm parameters have the potential to create a hexagonal swarm.
+The field effects are selected so as to ensure the swarm parameters have the potential to create a hexagonal swarm.
+Figure \ref{emerge:FLOOD1} shows the initial swarm deployment and Figure \ref{emerge:FLOOD6} shows the final swarm distribution for the simulation.
+
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FLOOD1}
+{Flood fill Start.\label{emerge:FLOOD1}}
+
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FLOOD6}
+{Flood fill End.\label{emerge:FLOOD6}}
 
@@
 
-Figure~\ref{emerge:FIELDFILL-MAG} shows the resultant magnitude for the swarm for the entire simulation.
+Figure~\ref{emerge:FIELDFILL-MAG} shows the cohesion/repulsion metric for the swarm for the entire simulation.
 Between 100 seconds and 105 seconds there is a significant change in the magnitude where the value becomes negative.
-This indicates there is a compression effect on the swarm and it is unable to expand the inter-agent distances.
+This indicates there is a compression effect on the swarm which indictaes the area is filled.
 
@@
 The initial deployment is shown at 0 seconds followed by incremental changes in the distances and variance of the distances as each increment is made in the field effects.
-The simulation terminates after 12 seconds.
+The simulation terminates after 120 seconds.
 
@@
 
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FieldEffects2}
+{Heterogeneous field effects.\label{additional:fieldsWork}}
+
 The \emph{cohesion/repulsion} metric provides an indication of a swarm's potential movement.
@@
 In a homogeneous field effect swarm the two metrics together provide a deeper evaluation of a swarm's structure and `potential'.
-Consider the following: the repulsion field is increased but the internal distances do not change; as a result the \emph{cohesion/repulsion metric} rises: This indicates `something' is confining the swarm's distribution.
+Consider the following: the repulsion field is increased but the internal distances do not change; as a result the \emph{cohesion/repulsion metric} rises: This indicates `something' is confining the swarm's distribution as discussed in Section \ref{metric:ApplicationFloodFilling}.
 This analysis could be used in identifying effective swarm distribution for the coverage of a sensor array as discussed by Ramaithitima et al.~\cite{RWBK:15}
-
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FieldEffects2}
-{Complex swarm interactions.\label{additional:fieldsWork}}
 
@@
 
-A possible adaptation to this area of research is to use the swarm perimeter detection to identify a subset of agents then reduce the field effects (repulsion and cohesion) of those agents to increase the agent density at the perimeter.
+A possible adaptation to this area of research is to use the swarm perimeter detection to identify a subset of agents then reduce the field effects (repulsion and cohesion) of those agents to increase the agent density at the perimeter, this is similar to having hetrogenious swarm structures as shown in Figure~\ref{additional:fieldsWork}.
 Increasing the density would allow increased coverage of any sensor arrays that the swarm was using at the swarm edge.
+
 The issue with introducing heterogeneous field effects is that the existing distance-based metric would not allow the swarm's structure to be analysed correctly.
@@
 This is due to the metric being based upon inter-agent effects rather than physical distribution.
+
+\newpage
 

### 2018-04-17 paper.tex
Minor
Minor Typos

@@
 
-This paper will refer to the scalarised resultant magnitude as the \emph{`cohesion/repulsion metric'} of the relationship.
+This paper will refer to the scalared resultant magnitude as the \emph{`cohesion/repulsion metric'} of the relationship.
 The `state' of a swarm is therefore the effect the environmental constraints and algorithms have upon the agent resultant magnitude.
@@
 Between 100 seconds and 105 seconds there is a significant change in the magnitude where the value becomes negative.
-This indicates there is a compression effect on the swarm which indictaes the area is filled.
+This indicates there is a compression effect on the swarm which indicates the area is filled.
 
@@
 
-A possible adaptation to this area of research is to use the swarm perimeter detection to identify a subset of agents then reduce the field effects (repulsion and cohesion) of those agents to increase the agent density at the perimeter, this is similar to having hetrogenious swarm structures as shown in Figure~\ref{additional:fieldsWork}.
+A possible adaptation to this area of research is to use the swarm perimeter detection to identify a subset of agents then reduce the field effects (repulsion and cohesion) of those agents to increase the agent density at the perimeter, this is similar to having heterogeneous swarm structures as shown in Figure~\ref{additional:fieldsWork}.
 Increasing the density would allow increased coverage of any sensor arrays that the swarm was using at the swarm edge.
@@
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Images/eliot}}]{Neil Eliot}
-has been a Senior Lecturer at Northumbria University since 1998.
-Neil was awarded his Ph.D.
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Images/eliot}}]{Neil Eliot} has been a Senior Lecturer at Northumbria University since 1998.
+Prior to this Neil worked in the chemical industry and the NHS.
+He was awarded his Ph.D.
 from Northumbria in 2017, his Pg.C.
@@
 in 1989 all in Computing.
-His research areas include swarm theory and cybersecurity.
-Prior to this Neil worked in the chemical industry and the NHS.
-His research interests are in the area of network cybersecurity, smart home technologies and swarm theory.
+His research interests are in the area of swarm theory and cybersecurity.
 \end{IEEEbiography}

### 2018-04-18 paper.tex
Minors

@@
 
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FLOOD0}
+{Flood fill Start.\label{emerge:FLOOD0}}
+
 \Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FLOOD1}
-{Flood fill Start.\label{emerge:FLOOD1}}
+{Data Capture Start.\label{emerge:FLOOD1}}
 

### 2018-04-18 paper.tex
Minor abstract change and new flood graphics

--- 2018-04-17 paper.tex
@@
 creates.  In addition the \emph{cohesion/repulsion} metric also identifies a
-swarm as expanding or cohesive.
+swarm as expanding or cohesive which can be used in the analysis of a swarm 
+filling a space to identify an exit condition.
 \end{abstract}
@@
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FieldEffects}
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=7.5cm]{figures/FieldEffects}
 {Agent field effects.\label{methods:FieldEffects}}
@@
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/Stability1}
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/Stability1}
 {Internal movement cohesion (no repulsion).\label{methods:Stability1}}
@@
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/Stability2}
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/Stability2}
 {Internal movement repulsion.\label{methods:Stability2}}
@@
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/Stability3}
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/Stability3}
 {Internal movement cohesion and repulsion.\label{methods:Stability3}}
@@
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/Stability4}
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/Stability4}
 {Internal movement equilibrium.\label{methods:Stability4}}
@@
 The field effects are selected so as to ensure the swarm parameters have the potential to create a hexagonal swarm.
-Figure \ref{emerge:FLOOD1} shows the initial swarm deployment and Figure \ref{emerge:FLOOD6} shows the final swarm distribution for the simulation.
+Figure \ref{emerge:FLOOD0} shows the initial swarm deployment.
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FLOOD1}
-{Flood fill Start.\label{emerge:FLOOD1}}
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=8cm]{figures/FLOOD0}
+{Flood fill Start.\label{emerge:FLOOD0}}
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FLOOD6}
+Figure ~\ref{emerge:FLOOD1} shows the point a which data logging starts.
+The swarm at this stage has expanded into the space but is not filling it.
+This presents itself as a positive cohesion/repulsion metric value which indicates the swarm is cohesive and the agents are still ``pulling" together.
+Figure \ref{emerge:FLOOD6} shows the final swarm distribution for the simulation, at this point the cohesion/repulsion metric has fell below zero indicating the swarm is trying to expand.
+
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=8cm]{figures/FLOOD1}
+{Data Capture Start.\label{emerge:FLOOD1}}
+
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=8cm]{figures/FLOOD6}
 {Flood fill End.\label{emerge:FLOOD6}}
@@
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=16cm]{figures/FIELDFILL-MAG}
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=13cm]{figures/FIELDFILL-MAG}
 {Cohesion/Repulsion metric 0-120 seconds.\label{emerge:FIELDFILL-MAG}}
@@
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=16cm]{figures/FIELDFILL-DIST}
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=13cm]{figures/FIELDFILL-DIST}
 {Distance metric 0-120 seconds.\label{emerge:FIELDFILL-DIST}}

### 2018-04-19 paper.tex
Modified abstract

--- 2018-04-17 paper.tex
@@
   the extent to which they cause unnecessary movement, i.e. movement which
-occurs as a side affect of the algorithm rather than the the intended
-function. This paper examines the widely used distance metric as a mechanism
-to measure the internal movement of agents and introduces a new
-\emph{cohesion/repulsion metric}.  Both metrics are based upon the resultant
-internal movement of a swarm identified by analysing the changes in the
-inter-agent interactions. The two metrics differ in their approach to
-identifying the changes. The distance metric uses variations in the
-inter-agent distances and the new metric uses the resultant magnitudes from
-the agents' \emph{inter-agent vector magnitudes}  which we claim offers a
+  occurs as a side effect of the algorithm rather than from the fulfilment of
+  the intended function of the swarm. This paper examines the widely used
+  distance metric as a mechanism to measure the internal movement of agents,
+  introduces an entirely new \emph{cohesion/repulsion metric}, and compares and
+  contrasts the two.  Both metrics are based upon the resultant internal
+  movement of a swarm, identified by analysing changes in inter-agent
+  interactions. The metrics differ in their approach to identifying the
+  changes. The distance metric uses variations in inter-agent distances,
+  whereas the new metric is based on the magnitudes of \emph{inter-agent
+  cohesion and repulsion vectors}. The cohesion/repulsion metric offers a
   number of advantages over the distance metric.  Both metrics allow a
-comparison of the effects of different swarming algorithms on a swarm's
-structure to identify the `level' of change in a swarm's internal structure
-which is an indication of the amount of unnecessary movement an algorithm
-creates.  In addition the \emph{cohesion/repulsion} metric also identifies a
-swarm as expanding or cohesive.
+  comparison of the effects of different swarming algorithms on the `amount of
+  change' in a swarm's internal structure, giving an indication of the amount
+  of unnecessary movement that an algorithm generates. In addition, the
+  \emph{cohesion/repulsion} metric also allows the identification of the
+  essential characteristic of a swarm as `expanding', `stable', or
+  `contracting'. This opens up the possibility of using the cohesion/repulsion
+  metric in ways that are not possible with the distance metric in practical
+  swarming applications, such as area filling and reconnaissance.
 \end{abstract}
@@
 %% The magnitude based metric identifies the resultant internal magnitude (vector magnitude) through the cohesive and repulsive magnitudes that exist between each agent pair in a swarm. 
-An agent's resultant vector magnitude is used as a scalar values to give an overall `size' to the agent relationship.
+An agent's resultant vector magnitude is used as a scalar value to give an overall `size' to the agent relationship.
 The resultant scalar value is then modified based upon the `size' of the two relationship magnitudes.
 
-This paper will refer to the scalared resultant magnitude as the \emph{`cohesion/repulsion metric'} of the relationship.
+This paper will refer to the scalar resultant magnitude as the \emph{`cohesion/repulsion metric'} of the relationship.
 The `state' of a swarm is therefore the effect the environmental constraints and algorithms have upon the agent resultant magnitude.

### 2018-04-19 paper.tex

--- 2018-04-18 paper.tex
@@
   the extent to which they cause unnecessary movement, i.e. movement which
-occurs as a side affect of the algorithm rather than the the intended
-function. This paper examines the widely used distance metric as a mechanism
-to measure the internal movement of agents and introduces a new
-\emph{cohesion/repulsion metric}.  Both metrics are based upon the resultant
-internal movement of a swarm identified by analysing the changes in the
-inter-agent interactions. The two metrics differ in their approach to
-identifying the changes. The distance metric uses variations in the
-inter-agent distances and the new metric uses the resultant magnitudes from
-the agents' \emph{inter-agent vector magnitudes}  which we claim offers a
+  occurs as a side effect of the algorithm rather than from the fulfilment of
+  the intended function of the swarm. This paper examines the widely used
+  distance metric as a mechanism to measure the internal movement of agents,
+  introduces an entirely new \emph{cohesion/repulsion metric}, and compares and
+  contrasts the two.  Both metrics are based upon the resultant internal
+  movement of a swarm, identified by analysing changes in inter-agent
+  interactions. The metrics differ in their approach to identifying the
+  changes. The distance metric uses variations in inter-agent distances,
+  whereas the new metric is based on the magnitudes of \emph{inter-agent
+  cohesion and repulsion vectors}. The cohesion/repulsion metric offers a
   number of advantages over the distance metric.  Both metrics allow a
-comparison of the effects of different swarming algorithms on a swarm's
-structure to identify the `level' of change in a swarm's internal structure
-which is an indication of the amount of unnecessary movement an algorithm
-creates.  In addition the \emph{cohesion/repulsion} metric also identifies a
-swarm as expanding or cohesive which can be used in the analysis of a swarm 
-filling a space to identify an exit condition.
+  comparison of the effects of different swarming algorithms on the `amount of
+  change' in a swarm's internal structure, giving an indication of the amount
+  of unnecessary movement that an algorithm generates. In addition, the
+  \emph{cohesion/repulsion} metric also allows the identification of the
+  essential characteristic of a swarm as `expanding', `stable', or
+  `contracting'. This opens up the possibility of using the cohesion/repulsion
+  metric in ways that are not possible with the distance metric in practical
+  swarming applications, such as area filling and reconnaissance.
 \end{abstract}
@@
 
-\newpage
+\clearpage
 

### 2018-04-19 paper.tex
Done

@@
 These effects are \emph{cohesion}, to draw agents closer, and \emph{repulsion} to prevent agents colliding.
-\emph{Field effects} have ranges around an agent and determine the effect other agents have upon it's movement~(Figure~\ref{methods:FieldEffects}).
+\emph{Field effects} have ranges around an agent and determine the effect other agents have upon it's movement~(Fig.~\ref{methods:FieldEffects}).
 It is usual for the cohesion field to have a radius $C_b$ which is larger than the repulsion radius $R_b$.
@@
 \section{Inter-agent vector magnitude effect on internal movement}\label{Section:StabilityMagnitude}
-Figure \ref{methods:Stability5} shows the cohesion and repulsion vector contributions to $v_c(b)$ and $v_r(b)$ due to neighbour $b'$, as given in (\ref{eq:FlyToCentre1}) and (\ref{eq:Repulsion1}).
+Fig.~\ref{methods:Stability5} shows the cohesion and repulsion vector contributions to $v_c(b)$ and $v_r(b)$ due to neighbour $b'$, as given in (\ref{eq:FlyToCentre1}) and (\ref{eq:Repulsion1}).
 Notice that the vectors are along the line of separation $bb'$.
@@
 \section{Swarm movement analysis\label{metric:MagnitudeDynamics2}}
-The repulsion and cohesion vectors are generated for an agent through the interaction of their field effects~(Figure~\ref{methods:FieldEffects}).
-There are a limited number of interactions that can occur; These are illustrated in Figures \ref{methods:Stability2}, \ref{methods:Stability3}, \ref{methods:Stability4} and \ref{methods:StabilityNullVector}.
+The repulsion and cohesion vectors are generated for an agent through the interaction of their field effects~(Fig.~\ref{methods:FieldEffects}).
+There are a limited number of interactions that can occur; These are illustrated in Fig.~\ref{methods:Stability2}, \ref{methods:Stability3}, \ref{methods:Stability4} and \ref{methods:StabilityNullVector}.
 
@@
 
-Figure~\ref{methods:Stability1} shows two agents within each other's cohesion fields but sufficiently distant to be outside of the repulsion fields.
+Fig.~\ref{methods:Stability1} shows two agents within each other's cohesion fields but sufficiently distant to be outside of the repulsion fields.
 In this case $|k_cv_c| > 0$ and $|k_rv_r| = 0$: the result is the agent's resultant magnitude $v(b)$ causes the agent to move towards is neighbour $b'$.
@@
 
-Figure~\ref{methods:Stability2} shows two agents close together with repulsion dominating cohesion: $|k_cv_c|~<~|k_rv_r|$. The resultant vector directs the agents away from each other.
+Fig.~\ref{methods:Stability2} shows two agents close together with repulsion dominating cohesion: $|k_cv_c|~<~|k_rv_r|$. The resultant vector directs the agents away from each other.
 
@@
 
-Figure~\ref{methods:Stability3} shows two agents close together but with cohesion vector magnitudes greater than the repulsion magnitudes $|k_cv_c|~>~|k_rv_r|$.
+Fig.~\ref{methods:Stability3} shows two agents close together but with cohesion vector magnitudes greater than the repulsion magnitudes $|k_cv_c|~>~|k_rv_r|$.
 The resultant vector draws the agents together.
@@
 
-Figure~\ref{methods:Stability4} shows two agents close together with $|k_cv_c|~=~|k_rv_r|$ the resultant vector is a \emph{null vector} and the agents have no overall influence upon each other due to the magnitude of the resultant vector being zero.
+Fig.~\ref{methods:Stability4} shows two agents close together with $|k_cv_c|~=~|k_rv_r|$ the resultant vector is a \emph{null vector} and the agents have no overall influence upon each other due to the magnitude of the resultant vector being zero.
 Table~\ref{tab:SampleEquilibrium} is an extract of data from the simulator.
@@
 
-If we consider the equilibrium state (Figure~\ref{methods:Stability4}) the resultant $v(b)=(0,0)$.
+If we consider the equilibrium state (Fig.~\ref{methods:Stability4}) the resultant $v(b)=(0,0)$.
 A null vector cannot be normalised to produce a directional vector ($\hat{v} = \frac{v}{|v|}$ if $v\neq0$; $0$ if $v=0$) The effect is that the agent will remain stationary.
-If all agent pairs are in this condition the swarm will stop moving~(Figure~\ref{methods:StabilityNullVector}).
+If all agent pairs are in this condition the swarm will stop moving~(Fig.~\ref{methods:StabilityNullVector}).
 
@@
 
-If a swarm is goal-based, the addition of a \emph{directional vector} will prevent all agents simultaneously producing null vectors as shown in~(Figure~\ref{concave:VoidPerimeter2}, Equation~(\ref{eq:InterAgentMovementDirectional1})).
+If a swarm is goal-based, the addition of a \emph{directional vector} will prevent all agents simultaneously producing null vectors as shown in~(Fig.~\ref{concave:VoidPerimeter2}, Equation~(\ref{eq:InterAgentMovementDirectional1})).
 Here, $k_dv_{d}(b)$ is the calculated directional vector to be applied to an agent.
@@
 Due to the dynamic nature of a swarm, maintaining optimum internal movement as in~Figure \ref{methods:StabilityNullVector} is highly unlikely.
-The agent pairs may fluctuate between the 4 states~(Figures~\ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}); These transitions are \emph{jitter}.
+The agent pairs may fluctuate between the 4 states~(Fig.~\ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}); These transitions are \emph{jitter}.
 The degree to which this variation occurs can be measured using either the change in distance between the agents, or the change in the magnitude $P(b)$~(\ref{eq:CohesionRepulsion}) between the agents.
@@
 \section{Cohesion/repulsion metric\label{section:MagnitudeDynamics}}
-An agent's \emph{resultant magnitude} is identified by Gazi and Passino~\cite{GP:11} and Barnes et al~\cite{BFV:07} as a `resultant characteristic' of a swarm.
+% An agent's \emph{resultant magnitude} is identified by Gazi and Passino~\cite{GP:11} and Barnes et al~\cite{BFV:07} as a `resultant characteristic' of a swarm.
 The sum of the cohesion and repulsion vectors between an agent $b$ and a neighbour $b'$ produce a resultant vector.
 
-The cohesion/repulsion internal movement (\emph{cohesion/repulsion metric}) is measured by identifying the balance between the cohesion and repulsion vectors between agents as shown in~(\ref{eq:CohesionRepulsion}).
+% The cohesion/repulsion internal movement (\emph{cohesion/repulsion metric}) is measured by identifying the balance between the cohesion and repulsion vectors between agents as shown in~(\ref{eq:CohesionRepulsion}).
 This is derived from the \emph{inter-agent cohesion/repulsion vector} as shown in~(\ref{eq:InterAgentMovementDirectional1}).
@@
 The identification of this deviation produces the clarifying component of the agent \emph{cohesion/repulsion metric}.
-%% The magnitude based metric identifies the resultant internal magnitude (vector magnitude) through the cohesive and repulsive magnitudes that exist between each agent pair in a swarm. 
-An agent's resultant vector magnitude is used as a scalar value to give an overall `size' to the agent relationship.
+% %% The magnitude based metric identifies the resultant internal magnitude (vector magnitude) through the cohesive and repulsive magnitudes that exist between each agent pair in a swarm. 
+% An agent's resultant vector magnitude is used as a scalar value to give an overall `size' to the agent relationship.
 The resultant scalar value is then modified based upon the `size' of the two relationship magnitudes.
 
-This paper will refer to the scalar resultant magnitude as the \emph{`cohesion/repulsion metric'} of the relationship.
+% This paper will refer to the scalar resultant magnitude as the \emph{`cohesion/repulsion metric'} of the relationship.
 The `state' of a swarm is therefore the effect the environmental constraints and algorithms have upon the agent resultant magnitude.
@@
 
-The resultant \emph{cohesion/repulsion metric}, which is a scalar, is generated by comparing the length of the cohesion and repulsion vector magnitudes.
+% The resultant \emph{cohesion/repulsion metric}, which is a scalar, is generated by comparing the length of the cohesion and repulsion vector magnitudes.
 If the cohesion magnitude is larger than the repulsion vector magnitude the resultant length is considered to be positive.
@@
 
-The \emph{`cohesion/repulsion metric'} is therefore the scalar value of the \emph{`agent resultant magnitude'} which is altered based upon the difference between the repulsion and cohesion magnitudes~(\ref{eq:CohesionRepulsion}).
+% The \emph{`cohesion/repulsion metric'} is therefore the scalar value of the \emph{`agent resultant magnitude'} which is altered based upon the difference between the repulsion and cohesion magnitudes~(\ref{eq:CohesionRepulsion}).
+This paper defines the cohesion/repulsion metric scalar for an agent as the length of $v(b)$  as defined (\ref{eq:InterAgentMovement2}): that is, the weighted effects of cohesion and repulsion by neighbours of the agent.
+This number carries a negative sign if repulsion dominatesand a positive sign if cohesion dominates and a positive sign if cohesion donates.
+This is summarised by~(\ref{eq:CohesionRepulsion}).
 
 If an agent's {cohesion/repulsion metric} is negative the swarm's bias is to expand as the repulsion is greater than cohesion.
-This is seen in the disorganised stage of a swarm~(Figure~\ref{coord:BaselineMagnitude1}).
+This is seen in the disorganised stage of a swarm~(Fig.~\ref{coord:BaselineMagnitude1}).
 The disorganised stage is the initial random deployment of the agents.
@@
 
-\section{Application: Flood filling}\label{metric:ApplicationFloodFilling}
+\section{Application: Area filling}\label{metric:ApplicationFloodFilling}
 
-Area flooding is a technique used to fill an enclosed area with agents such that they are distributed as `effectively' as possible throughout an area.
-An optimum distribution of agents is not necessarily an even distribution.
-The distribution of the agents is governed by the field effects and the interaction of agents with obstacles (\textit{interaction vectors}).
+Area filling is a technique used to fill an enclosed area with agents such that they are distributed as `effectively' as possible throughout an area.
 
 Filling an area can be applied to tasks that require an unknown environment to be analysed or surveyed.
-Consider a disaster area following a landslide or a building collapsing following an earthquake; the movement in the land and buildings will produce spaces that are unmapped.
-The unmapped areas may require investigation to locate people, resources or to create some form of sensor network to analyse the conditions within the area such as creating a heat map or to identify the location toxic gases.
+Consider a disaster area following a landslide or a building collapse following an earthquake; the movement in the land and buildings will produce areas that are unmapped.
+The unmapped areas may require investigation to locate people, or resources, or to create some form of sensor network to analyse the conditions within the area, such as creating a heat map or identifying the location of toxic gases.
 
@@
 Alvissalim et al.
-(2012) discuss the application of commercially available drones to provide a communication infrastructure across an unknown disaster areas~\cite{AZHMJJM:12}.
+(2012) discuss the application of commercially available drones to provide a communication infrastructure across an unmapped disaster area~\cite{AZHMJJM:12}.
 Scheutz and Bauer (2006) use both cohesion and repulsion as a mechanism to create coverage of an area that requires protection in an adversarial environment~\cite{SB:06}.
 
-The concept behind the field effect expansion is to increase both the repulsion and cohesion fields over a period of time.
-This increase in the field effects makes the agents increase the distance between each other expanding the swarm as a whole.
+The concept behind the area filling is to increase both the repulsion and cohesion fields over a period of time.
+This increase in the field effects makes the agents increase the distance between each other, expanding the swarm as a whole.
 The expansion increases until, due to boundary compression, the swarm is unable to expand further.
@@
 
-Figure ~\ref{emerge:FLOOD1} shows the point a which data logging starts.
+Fig.~\ref{emerge:FLOOD1} shows the point a which data logging starts.
 The swarm at this stage has expanded into the space but is not filling it.
 This presents itself as a positive cohesion/repulsion metric value which indicates the swarm is cohesive and the agents are still ``pulling" together.
-Figure \ref{emerge:FLOOD6} shows the final swarm distribution for the simulation, at this point the cohesion/repulsion metric has fell below zero indicating the swarm is trying to expand.
+Fig.~\ref{emerge:FLOOD6} shows the final swarm distribution for the simulation, at this point the cohesion/repulsion metric has fell below zero indicating the swarm is trying to expand.
 
@@
 
-Figure~\ref{emerge:FIELDFILL-MAG} shows the cohesion/repulsion metric for the swarm for the entire simulation.
+Fig.~\ref{emerge:FIELDFILL-MAG} shows the cohesion/repulsion metric for the swarm for the entire simulation.
 Between 100 seconds and 105 seconds there is a significant change in the magnitude where the value becomes negative.
@@
 
-Figure~\ref{emerge:FIELDFILL-DIST}~shows the distance metric for the simulation.
+Fig.~\ref{emerge:FIELDFILL-DIST}~shows the distance metric for the simulation.
 The initial deployment is shown at 0 seconds followed by incremental changes in the distances and variance of the distances as each increment is made in the field effects.
@@
 This is independent of the physical distribution of the agents within the swarm.
-The lack of dependence on the physical distribution allows the metric to be used in heterogeneous field effect swarms~(Figure~\ref{additional:fieldsWork}) where the physical distribution of agents may vary due to the varied field effect ranges.
+The lack of dependence on the physical distribution allows the metric to be used in heterogeneous field effect swarms~(Fig.~\ref{additional:fieldsWork}) where the physical distribution of agents may vary due to the varied field effect ranges.
 
@@
 
-A possible adaptation to this area of research is to use the swarm perimeter detection to identify a subset of agents then reduce the field effects (repulsion and cohesion) of those agents to increase the agent density at the perimeter, this is similar to having heterogeneous swarm structures as shown in Figure~\ref{additional:fieldsWork}.
+A possible adaptation to this area of research is to use the swarm perimeter detection to identify a subset of agents then reduce the field effects (repulsion and cohesion) of those agents to increase the agent density at the perimeter, this is similar to having heterogeneous swarm structures as shown in Fig.~\ref{additional:fieldsWork}.
 Increasing the density would allow increased coverage of any sensor arrays that the swarm was using at the swarm edge.

### 2018-06-06 paper.tex
Todays stuff

@@
 
-\title{An Improved Metric for the Analysis of Swarms}
+\title{A New Metric for the Analysis of Swarms}
 \author{\uppercase{Neil Eliot}\authorrefmark{1},
@@
 \begin{abstract}
-  An important criterion in the assessment of swarm coordination algorithms is
-  the extent to which they cause unnecessary movement, i.e. movement which
-  occurs as a side effect of the algorithm rather than from the fulfilment of
-  the intended function of the swarm. This paper examines the widely used
-  distance metric as a mechanism to measure the internal movement of agents,
-  introduces an entirely new \emph{cohesion/repulsion metric}, and compares and
-  contrasts the two.  Both metrics are based upon the resultant internal
-  movement of a swarm, identified by analysing changes in inter-agent
-  interactions. The metrics differ in their approach to identifying the
-  changes. The distance metric uses variations in inter-agent distances,
-  whereas the new metric is based on the magnitudes of \emph{inter-agent
-  cohesion and repulsion vectors}. The cohesion/repulsion metric offers a
-  number of advantages over the distance metric.  Both metrics allow a
-  comparison of the effects of different swarming algorithms on the `amount of
-  change' in a swarm's internal structure, giving an indication of the amount
-  of unnecessary movement that an algorithm generates. In addition, the
-  \emph{cohesion/repulsion} metric also allows the identification of the
+  % An important criterion in the assessment of swarm coordination algorithms is
+  % the extent to which they cause unnecessary movement, i.e. movement which
+  % occurs as a side effect of the algorithm rather than from the fulfilment of
+  % the intended function of the swarm. 
+  A widely used metric for the assesment of swarm coordination algortithms is a metric based upon the distance between agents.
+  This paper examines this metric and introduces an entirely new metric based upon the magnitude of cohesion and repulsion between agents, comparing and
+  contrasting the two.  
+  The \emph{cohesion/repulsion} metric offers a
+  number of advantages over the distance metric. The
+  \emph{cohesion/repulsion} metric allows the identification of the
   essential characteristic of a swarm as `expanding', `stable', or
-  `contracting'. This opens up the possibility of using the cohesion/repulsion
-  metric in ways that are not possible with the distance metric in practical
-  swarming applications, such as area filling and reconnaissance.
+  `contracting'. This is not possible with the distance metric. Practical
+  swarming applications where this might apply include area filling and reconnaissance.
 \end{abstract}
@@
 
-\section{Standard deviation in distance metric}\label{Section:VarianceInDistance}
-The mechanism above provides an overall indication of the distribution of the agents.
-The model for a swarm $S$ ($\mu_d(S)$) is not sufficient to give an indication of the internal distribution of the agents.
+%\section{Standard deviation in distance metric}\label{Section:VarianceInDistance}
+The mean distance provides a indication of the large scale structure of the swarm.
+However it is not sufficient to give an indication of the internal distribution of the agents.
 The standard deviation clarifies the distribution within the swarm as shown in~(\ref{eq:SwarmStabilityDistance3}).
@@
 
-Fig.~\ref{emerge:FIELDFILL-DIST}~shows the distance metric for the simulation.
+Fig.~\ref{emerge:FIELDFILL-DIST}~shows the distance metric for the same simulation.
 The initial deployment is shown at 0 seconds followed by incremental changes in the distances and variance of the distances as each increment is made in the field effects.
 The simulation terminates after 120 seconds.
+There is no feature of this graph which indicates that the area is filled.
+The cohesion/repulsion metric clearly tells us more.
 
@@
 \section{Future work\label{metric:Future}}
+The metric can be applied to the analysis of the effects of coordination algorithms upon a swarm's structure.
+An application to floodfilling has just been discussed.
+The metric can also be applied to identifying the effects a `healing' algorithm may have upon a swarm's structure.
+By this we mean an algorithm which removes `voids' in a swarm.
+
 The detection of a swarm's perimeter \cite{MD:09, MJ:08, ZAPS:07, JG:13} identifies a subset of a swarm which can be used for coordination.
@@
 The issue with introducing heterogeneous field effects is that the existing distance-based metric would not allow the swarm's structure to be analysed correctly.
-The cohesion/repulsion metric discussed in this paper would still work within this environment as a tool to identify a swarm's structure.
+We believe that cohesion/repulsion metric discussed in this paper will still identify the swarm's structure when the swarm is hetrogenious.
 This is due to the metric being based upon inter-agent effects rather than physical distribution.

### 2018-06-07 paper.tex
Stuff
Updates

@@
 \markboth
-{Eliot \headeretal: An Improved Metric for the Analysis of Swarms}
-{Eliot \headeretal: An Improved Metric for the Analysis of Swarms}
+{Eliot \headeretal: A New Metric for the Analysis of Swarms}
+{Eliot \headeretal: A New Metric for the Analysis of Swarms}
 
@@
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=13cm]{figures/FIELDFILL-MAG}
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=14cm]{figures/FIELDFILL-MAG}
 {Cohesion/Repulsion metric 0-120 seconds.\label{emerge:FIELDFILL-MAG}}
@@
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=13cm]{figures/FIELDFILL-DIST}
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=14cm]{figures/FIELDFILL-DIST}
 {Distance metric 0-120 seconds.\label{emerge:FIELDFILL-DIST}}

### 2018-06-07 paper.tex
Updates

@@
 The \emph{distance metric} which uses variations in the inter-agent distances, as discussed by Navarro et al.~\cite{NIM:09}.
-The new metric uses the resultant magnitudes of the agents' \emph{inter-agent vectors} that are defined by interactions of and agent with it's neighbours which allows a more accurate analysis of a swarm when using non-homogeneous sensor distances.
+The new metric uses the resultant magnitudes of the agents' \emph{inter-agent vectors} that are defined by interactions of an agent with it's neighbours which allows a more accurate analysis of a swarm when using non-homogeneous sensor distances.
 The metric also identifies when a swarm is expanding or contracting.

### 2018-06-13 paper.tex

@@
   % the intended function of the swarm. 
-  A widely used metric for the assesment of swarm coordination algortithms is a metric based upon the distance between agents.
-  This paper examines this metric and introduces an entirely new metric based upon the magnitude of cohesion and repulsion between agents, comparing and
-  contrasting the two.  
-  The \emph{cohesion/repulsion} metric offers a
-  number of advantages over the distance metric. The
-  \emph{cohesion/repulsion} metric allows the identification of the
-  essential characteristic of a swarm as `expanding', `stable', or
-  `contracting'. This is not possible with the distance metric. Practical
-  swarming applications where this might apply include area filling and reconnaissance.
+  A widely used metric for the analysis of swarm coordination algorithms is
+  based upon the distance between agents. This paper examines this metric and
+  introduces an entirely new metric based upon the magnitude of the cohesion
+  and repulsion vectors between agents. In comparing and contrasting the two
+  metrics, we find that the \emph{cohesion/repulsion} metric offers a number of
+  advantages over the distance metric. In particular, the
+  \emph{cohesion/repulsion} metric allows the identification of the essential
+  characteristic of a swarm as `expanding', `stable', or `contracting'. This is
+  not possible with the distance metric.  Practical swarming applications where
+  the new metric is useful include area filling and reconnaissance.
 \end{abstract}
@@
 \section{Future work\label{metric:Future}}
-The metric can be applied to the analysis of the effects of coordination algorithms upon a swarm's structure.
-An application to floodfilling has just been discussed.
-The metric can also be applied to identifying the effects a `healing' algorithm may have upon a swarm's structure.
-By this we mean an algorithm which removes `voids' in a swarm.
 
-The detection of a swarm's perimeter \cite{MD:09, MJ:08, ZAPS:07, JG:13} identifies a subset of a swarm which can be used for coordination.
-The use of the perimeter detection allows a swarm to be used in many different ways~\cite{ZFG:13, AKK:08, APZDAMC:09, AZDPS:11}.
+The new metric can be applied to the analysis of the effects of coordination
+algorithms upon a swarm's structure. An application to flood filling has just
+been discussed. The metric can also be applied to identifying the effects of 
+`healing' algorithms, e.g. to remove `voids' from a swarm. We plan
+to investigate this, and other, practical applications of the metric.
 
-A possible adaptation to this area of research is to use the swarm perimeter detection to identify a subset of agents then reduce the field effects (repulsion and cohesion) of those agents to increase the agent density at the perimeter, this is similar to having heterogeneous swarm structures as shown in Fig.~\ref{additional:fieldsWork}.
-Increasing the density would allow increased coverage of any sensor arrays that the swarm was using at the swarm edge.
-
-The issue with introducing heterogeneous field effects is that the existing distance-based metric would not allow the swarm's structure to be analysed correctly.
-We believe that cohesion/repulsion metric discussed in this paper will still identify the swarm's structure when the swarm is hetrogenious.
-This is due to the metric being based upon inter-agent effects rather than physical distribution.
+The detection of a swarm's perimeter~\cite{MD:09, MJ:08, ZAPS:07, JG:13}
+identifies a subset of a swarm which can be used for coordination. 
+Perimeter detection is widely used in swarm coordination~\cite{ZFG:13, AKK:08,
+APZDAMC:09, AZDPS:11}. A possible adaptation of this area of research is to
+investigate the reduction of the field effects (cohesion and repulsion) of
+agents at the perimeter, so increasing agent density in this region. Increasing
+the density would allow increased coverage by the swarm at its perimeter. This
+could improve the granularity of analysis, e.g. when using sensor arrays at
+thr frontier of the swarm.
+However, this approach leads to swarms of agents having heterogeneous field
+effects, as shown in Fig.~\ref{additional:fieldsWork}. Clearly, the
+distance-based metric is not useful in the analysis of such swarms. We believe
+that the cohesion/repulsion metric introduced in this paper can identify a
+swarm's structure even when the swarm is heterogeneous, since the metric is
+based directly upon inter-agent field effects rather than the physical
+distribution of agents. This will also be the subject of forthcoming work.
 

### 2018-06-13 paper.tex
Section Update

@@
 
-\section{Calculating distance internal movement}
+% \section{Calculating distance internal movement}
 Equation~(\ref{eq:SwarmStabilityDistance1}) computes the mean distance of an agent to its neighbours $nbr(b)$. 
@@
 
-\section{Cohesion/repulsion internal movement model}\label{Section:StabilityModel}
+% \section{Cohesion/repulsion internal movement model}\label{Section:StabilityModel}
 Section~\ref{Section:StabilityMagnitude}~discussed calculation of resultant cohesion~(\ref{eq:FlyToCentre1}) and resultant repulsion~(\ref{eq:Repulsion1}) vectors for each agent deriving a combined \emph{cohesion/repulsion vector} using~(\ref{eq:InterAgentMovement2}).

### 2018-06-13 paper.tex
Future

@@
 The distance-based metric measures swarm structure in terms of position of agents.
-
-The distance-based metric provides an analysis of the distribution of the agents at a point in time and allows the agitation of the swarm to be assessed.
-It does not take into account the possible distribution of agents that the field effects \emph{could} produce.
+It provides an analysis at a point in time but does not allow the identification of the future behaviour of a swarm as `expanding', `stable' or `contracting'.
 
@@
 
-The \emph{cohesion/repulsion} metric provides an indication of a swarm's potential movement.
-This is independent of the physical distribution of the agents within the swarm.
-The lack of dependence on the physical distribution allows the metric to be used in heterogeneous field effect swarms~(Fig.~\ref{additional:fieldsWork}) where the physical distribution of agents may vary due to the varied field effect ranges.
+The \emph{cohesion/repulsion} metric provides an indication of a swarm's potential future movement. This is independent of the physical distribution of the agents within the swarm.
 
-In a homogeneous field effect swarm the two metrics together provide a deeper evaluation of a swarm's structure and `potential'.
-Consider the following: the repulsion field is increased but the internal distances do not change; as a result the \emph{cohesion/repulsion metric} rises: This indicates `something' is confining the swarm's distribution as discussed in Section \ref{metric:ApplicationFloodFilling}.
+The two metrics together provide a deeper evaluation of a swarm's structure and `potential' thean either of the metrics alone.
+Consider the following: the repulsion field is increased but the internal distances do not change; as a result the \emph{cohesion/repulsion metric} rises.
+This indicates `something' is confining the swarm's distribution as discussed in Section \ref{metric:ApplicationFloodFilling}.
 This analysis could be used in identifying effective swarm distribution for the coverage of a sensor array as discussed by Ramaithitima et al.~\cite{RWBK:15}
@@
 \section{Future work\label{metric:Future}}
-
 The new metric can be applied to the analysis of the effects of coordination
-algorithms upon a swarm's structure. An application to flood filling has just
-been discussed. The metric can also be applied to identifying the effects of 
+algorithms upon a swarm's structure. An application to flood filling has
+been discussed~(Section~\ref{metric:ApplicationFloodFilling}). The metric can also be applied to identifying the effects of 
 `healing' algorithms, e.g. to remove `voids' from a swarm. We plan

### 2018-06-13 paper.tex
Future

@@
 
-The two metrics together provide a deeper evaluation of a swarm's structure and `potential' thean either of the metrics alone.
-Consider the following: the repulsion field is increased but the internal distances do not change; as a result the \emph{cohesion/repulsion metric} rises.
+The two metrics together provide a deeper evaluation of a swarm's structure and `potential' than either of the metrics alone.
+Consider the following: the repulsion field is increased but the internal distances do not change; as a result the \emph{cohesion/repulsion metric} falls below zero.
 This indicates `something' is confining the swarm's distribution as discussed in Section \ref{metric:ApplicationFloodFilling}.

### 2018-07-10 paper.tex
Figure?

@@
 \section{Inter-agent vector magnitude effect on internal movement}\label{Section:StabilityMagnitude}
-Fig.~\ref{methods:Stability5} shows the cohesion and repulsion vector contributions to $v_c(b)$ and $v_r(b)$ due to neighbour $b'$, as given in (\ref{eq:FlyToCentre1}) and (\ref{eq:Repulsion1}).
+Figure~\ref{methods:Stability5} shows the cohesion and repulsion vector contributions to $v_c(b)$ and $v_r(b)$ due to neighbour $b'$, as given in (\ref{eq:FlyToCentre1}) and (\ref{eq:Repulsion1}).
 Notice that the vectors are along the line of separation $bb'$.
@@
 
-Fig.~\ref{methods:Stability1} shows two agents within each other's cohesion fields but sufficiently distant to be outside of the repulsion fields.
+Figure~\ref{methods:Stability1} shows two agents within each other's cohesion fields but sufficiently distant to be outside of the repulsion fields.
 In this case $|k_cv_c| > 0$ and $|k_rv_r| = 0$: the result is the agent's resultant magnitude $v(b)$ causes the agent to move towards is neighbour $b'$.
@@
 
-Fig.~\ref{methods:Stability2} shows two agents close together with repulsion dominating cohesion: $|k_cv_c|~<~|k_rv_r|$. The resultant vector directs the agents away from each other.
+Figure~\ref{methods:Stability2} shows two agents close together with repulsion dominating cohesion: $|k_cv_c|~<~|k_rv_r|$. The resultant vector directs the agents away from each other.
 
@@
 
-Fig.~\ref{methods:Stability3} shows two agents close together but with cohesion vector magnitudes greater than the repulsion magnitudes $|k_cv_c|~>~|k_rv_r|$.
+Figure~\ref{methods:Stability3} shows two agents close together but with cohesion vector magnitudes greater than the repulsion magnitudes $|k_cv_c|~>~|k_rv_r|$.
 The resultant vector draws the agents together.
@@
 
-Fig.~\ref{methods:Stability4} shows two agents close together with $|k_cv_c|~=~|k_rv_r|$ the resultant vector is a \emph{null vector} and the agents have no overall influence upon each other due to the magnitude of the resultant vector being zero.
+Figure~\ref{methods:Stability4} shows two agents close together with $|k_cv_c|~=~|k_rv_r|$ the resultant vector is a \emph{null vector} and the agents have no overall influence upon each other due to the magnitude of the resultant vector being zero.
 Table~\ref{tab:SampleEquilibrium} is an extract of data from the simulator.
@@
 
-Fig.~\ref{emerge:FLOOD1} shows the point a which data logging starts.
+Figure~\ref{emerge:FLOOD1} shows the point a which data logging starts.
 The swarm at this stage has expanded into the space but is not filling it.
@@
 
-Fig.~\ref{emerge:FIELDFILL-MAG} shows the cohesion/repulsion metric for the swarm for the entire simulation.
+Figure~\ref{emerge:FIELDFILL-MAG} shows the cohesion/repulsion metric for the swarm for the entire simulation.
 Between 100 seconds and 105 seconds there is a significant change in the magnitude where the value becomes negative.
@@
 
-Fig.~\ref{emerge:FIELDFILL-DIST}~shows the distance metric for the same simulation.
+Figure~\ref{emerge:FIELDFILL-DIST}~shows the distance metric for the same simulation.
 The initial deployment is shown at 0 seconds followed by incremental changes in the distances and variance of the distances as each increment is made in the field effects.

### 2018-07-31 paper.tex
Updates - Related Work

@@
 \newcommand{\subsubsubsection}[1]{\medskip\par\emph{#1:}}
+\newcommand{\card}[1]{\vert{#1}\vert}
+
 \hyphenation{op-tical net-works semi-conduc-tor}
@@
   based upon the distance between agents. This paper examines this metric and
-  introduces an entirely new metric based upon the magnitude of the cohesion
+  introduces an entirely new metric based upon the field effects that createof the cohesion
   and repulsion vectors between agents. In comparing and contrasting the two
@@
 Measuring this unnecessary movement allows the effectiveness of a swarming algorithm to be evaluated.
-If the internal movement can be reduced then stability of the swarm is improved and more of the swarms energy is focused on the goal.
+If the internal movement can be reduced then the stability of the swarm is improved and more of the swarms energy is focused on the goal.
 Reducing the internal movement also provides a more stable platform for the deployment of sensor arrays.
-Both metrics allow this evaluation to be carried out when all the sensor ranges are the same however the distance metric is not able to identify the internal movement accurately when the sensor ranges of each agent varies producing varying distances even though the interactions at the magnitude level indicates the agents are in fact stable and optimum.
+Both the metrics presented in this paper allow this evaluation to be carried out when all the sensor ranges are the same however the distance metric is not able to identify the internal movement accurately when the sensor ranges of each agent vary producing irregular inter-agent distances even though the interactions at the vector level indicate that the agents are in fact stable and optimum.
 This paper discusses the two metrics and how they differ in their approach to identify these changes.
-The \emph{distance metric} which uses variations in the inter-agent distances, as discussed by Navarro et al.~\cite{NIM:09}.
-The new metric uses the resultant magnitudes of the agents' \emph{inter-agent vectors} that are defined by interactions of an agent with it's neighbours which allows a more accurate analysis of a swarm when using non-homogeneous sensor distances.
+The \emph{distance metric} which uses variations in the inter-agent distances is used and discussed by many research teams including Navarro et al.~\cite{NIM:09}, Gazi and Passino~\cite{GP:02,GP:04}, Barnes et al.~\cite{BAF:06} and Schneider and Wildermuth~\cite{SW:03}.
+The new metric uses the resultant magnitudes of the agents' \emph{inter-agent vectors} that are defined by interactions of an agent with its neighbours.
 The metric also identifies when a swarm is expanding or contracting.
 
-This paper presents a swarm model and then applies the two metrics to this model to present a comparison of the two metrics and uses the new metric to identify a swarm's `state'.
+Many coordination algorithms exist to control the movement of agents within a swarm.
+The primary goal of these algorithms is to maintain a stable swarm structure.
+The main focus of this paper is the use any coordination algorithm that are based on \textit{field effects} to produce a metric that allows the swarm to be analysed.
+To allow this paper to present the findings a basic swarm model is introduced and then the two metrics are applied to this model to present a comparison.
+This includes the identification of a swarm's `state'.
+
+\section{Related Work}\label{Related}
+
+Many metrics exist for the analysis of swarms.
+However they tend to fall into two main categories based upon the area of interest when examining a swarm's characteristics.
+These two techniques are \textit{Centroidal} and \textit{Inter-agent distance} based.
+The common denominator of both these techniques is that they \textit{measure} a swarm's agent distribution and deviations from a predicted distribution.
+
+\textit{Centroidal} analysis: The centroid of a swarm is its \textit{centre of mass} as shown in Equation~\ref{eq:Centroid1}.
+This is often used as a reference point to determine the path of a swarm or to measure how agents are arranged within a swarm's structure.
+Navarro et al.
+use the centroid to produce a \textit{Mean Distance from the Boundary to the Center of the Group} metric.
+This metric takes the distance of each agent from the centroid and calculates the mean and the deviation to produce a measure~\cite{NIM:09}.
+Gazi and Passino also use this technique when analysing the stability of a swarm based on agent distribution~\cite{GP:11}.
+Gazi and Passino also use the centroid when analysing the stability of a swarm which is designed to maintain its position by analysing the changes in the centroids position over time~\cite{GP:04}.
+
+The centroid of the swarm is determined by taking the coordinate position of each agent and calculating the mean of the $x$ and $y$ positions as shown in Equation~\ref{eq:Centroid1}.
+
+\begin{equation}\label{eq:Centroid1}
+C_s = \frac{1}{\card{S}}\mathlarger{\sum_{b' \in S}} b'
+\end{equation}
+
+\textit{Inter-Agent Distance} analysis: A basic mechanism for analysing a swarm it to analyse the separation of all of the agents with respect to each agents visible neighbours.
+This technique is the most popular approach to analysing a swarms structure~\cite{BAF:06, BAFVM:06, BM:09, APZDAMC:09, GP:02, GP:04, GP:04a, GP:05, GP:11, MYP:09}.
 

### 2018-07-31 paper.tex
More Updates

@@
 
-\textit{Centroidal} analysis: The centroid of a swarm is its \textit{centre of mass} as shown in Equation~\ref{eq:Centroid1}.
+\textit{Centroidal} analysis: The centroid of a swarm is its \textit{centre of mass} as shown in Figure~\ref{com} as a red cross.
 This is often used as a reference point to determine the path of a swarm or to measure how agents are arranged within a swarm's structure.
 Navarro et al.
-use the centroid to produce a \textit{Mean Distance from the Boundary to the Center of the Group} metric.
-This metric takes the distance of each agent from the centroid and calculates the mean and the deviation to produce a measure~\cite{NIM:09}.
+use the centroid to produce a \textit{Mean Distance from the Boundary to the Center of the Group} metric~\cite{NIM:09}.
+This metric takes the distance of each agent from the centroid and calculates the mean and the deviation to produce a measure.
 Gazi and Passino also use this technique when analysing the stability of a swarm based on agent distribution~\cite{GP:11}.
@@
 
-The centroid of the swarm is determined by taking the coordinate position of each agent and calculating the mean of the $x$ and $y$ positions as shown in Equation~\ref{eq:Centroid1}.
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4.5cm]{figures/com}
+{Swarm Centre of Mass.\label{com}}
+
+The centroid of the swarm ($C_s$) is determined by taking the coordinate position of each agent and calculating the mean positions as shown in Equation~\ref{eq:Centroid1} where $b_{coord}$ is each agents coordinates and $S$ is the set of agents that make up the swarm.
 
 \begin{equation}\label{eq:Centroid1}
-C_s = \frac{1}{\card{S}}\mathlarger{\sum_{b' \in S}} b'
+C_s = \frac{1}{\card{S}}\mathlarger{\sum_{b \in S}} b_{coord}
 \end{equation}

### 2018-08-04 paper.tex
Improved Related Work

@@
 This includes the identification of a swarm's `state'.
+In the remainder of this paper Section~\ref{Section:Related} discusses related work, Section~\ref{Section:SwarmModelling} introduces a basic swarm model for evaluation.
+Section~\ref{Section:StabilityMagnitude} discusses the mathematical model that produces the swarming effect.
+Section~\ref{Section::Movement1} discusses how the model is implemented over time to produce movement.
+Section~\ref{Section:StabilityNullVector} discusses the effect of the model on stability.
+Section~\ref{Section:Jitter} discusses the effect the mathematical model can have upon the structure of a swarm in terms of potentially wasted energy.
+Section~\ref{Section:DistanceDynamics} discusses a distance based metric approach to swarm analysis.
+Section~\ref{Section:MagnitudeDynamics} introduces the new \textit{cohesion/repulsion-based} metric.
+Section~\ref{Section:ApplicationFloodFilling} compares the two metrics using a scenario of a swarm being used to fill an area.
+Section~\ref{Section:MagnitudeDistanceComparison} discusses the results of the experiments and Section~\ref{Section:FutureWork} discuss additional work that needs to be carried out to exploit the additional information that the new metric provides.
 
-\section{Related Work}\label{Related}
+\section{Related Work}\label{Section:Related}
 
 Many metrics exist for the analysis of swarms.
-However they tend to fall into two main categories based upon the area of interest when examining a swarm's characteristics.
-These two techniques are \textit{Centroidal} and \textit{Inter-agent distance} based.
-The common denominator of both these techniques is that they \textit{measure} a swarm's agent distribution and deviations from a predicted distribution.
+These metrics can be categorised as \textit{Centroidal}, \textit{Inter-agent distance} and \textit{magnitude} based.
+The common denominator of the \textit{centroidal} and \text{inter-agent distance} techniques is that they \textit{measure} the agent distribution of a swarm based on distance.
+\textit{Magnitude} based analysis is based upon inter-agent interaction.
 
-\textit{Centroidal} analysis: The centroid of a swarm is its \textit{centre of mass} as shown in Figure~\ref{com} as a red cross.
-This is often used as a reference point to determine the path of a swarm or to measure how agents are arranged within a swarm's structure.
+\textit{Centroidal} analysis involves identifying the centroid of a swarm, also referred to as the \textit{centre of mass}, as shown in Figure~\ref{com} as a red cross, and calculating distances from that position to each agent.
+The \textit{centroid} is also used as a reference point to determine the path of a swarm.
 Navarro et al.
 use the centroid to produce a \textit{Mean Distance from the Boundary to the Center of the Group} metric~\cite{NIM:09}.
-This metric takes the distance of each agent from the centroid and calculates the mean and the deviation to produce a measure.
-Gazi and Passino also use this technique when analysing the stability of a swarm based on agent distribution~\cite{GP:11}.
+This metric takes the distance of each agent from the centroid and calculates the mean and deviation to produce their measure.
+Gazi and Passino use this technique when analysing the stability of a swarm based on agent distribution~\cite{GP:11}.
 Gazi and Passino also use the centroid when analysing the stability of a swarm which is designed to maintain its position by analysing the changes in the centroids position over time~\cite{GP:04}.
+Barnes et.
+al.
+use the centroid to analyse the effects of using bivariate normal functions~\cite{BAF:06} and in their work on using potential fields~\cite{BAFVM:06}.
+Gazi and Passino use a centroid based analysis in~\cite{GP:02}~to identify directional changes is agent movements by creating a ``map'' of the potential fields of a swarm in its entirety and overlaying that to the ``centre'' of the swarm.
 
@@
 
-\textit{Inter-Agent Distance} analysis: A basic mechanism for analysing a swarm it to analyse the separation of all of the agents with respect to each agents visible neighbours.
-This technique is the most popular approach to analysing a swarms structure~\cite{BAF:06, BAFVM:06, BM:09, APZDAMC:09, GP:02, GP:04, GP:04a, GP:05, GP:11, MYP:09}.
+\textit{Inter-Agent Distance} analysis is a basic mechanism for analysing a swarm it to analyse the separation of all of the agents with respect to each agents visible neighbours.
+This is different to using the centroid approach as the measure is based on agent interactions rather than a resultant structure.
+This technique is the most popular approach to analysing a swarms structure.
+Gazi and Passino use a distance based analysis in \cite{GP:04, GP:11} to identify optimum distribution of agents based upon different cohesion and repulsion applications.
+Wu et.
+al.
+discuss using several metrics to evaluate emergent behaviours in swarm based on a Viscsek model.
+These include calculating the average number of neighbours for each agent, grouping agents based on numbers of neighbours (to identify clusters).
+They also use the average distance between agents~\cite{WCW:11}.
+
+\textit{Magnitude} based analysis uses the potential fields to calculate the ``degree'' of interaction between agent's.
+Bennet and McInnes use the resultant relationship between agents to identify equilibrium states when creating structures such as rings and controlling the direction of agents to created specific transformations~\cite{BM:09}.
+Wu et al.
+introduce the concept of a \textit{rejection potential} in an attempt to generate a kinetic energy score for a swarm~\cite{WCW:11}.
 
@@
 %% This section discusses the theory behind swarm movement and the effects that can be measured using both the metrics described above.
-\section{Swarm modelling}\label{SwarmModelling}
+\section{Swarm modelling}\label{Section:SwarmModelling}
 Currently, much swarm research uses field effects as the method of modelling inter-agent interactions~\cite{BAF:06, BAFVM:06, BM:09, APZDAMC:09, GP:02, GP:04, GP:04a, GP:05, GP:11, MYP:09}.
@@
 
-\section{Modelling movement}\label{sec:Movement1}
+\section{Modelling movement}\label{Section::Movement1}
 Each agent within a swarm calculates its \textit{movement-direction vector} based on its \textit{interaction} and \textit{destination} vectors.
@@
 
-\section{Internal movement and the null vector}\label{section:StabilityNullVector}
+\section{Internal movement and the null vector}\label{Section:StabilityNullVector}
 When the two vectors (cohesion and repulsion) have magnitudes that are equal and opposite they produce a null vector, indicating that two agents are optimally spaced for a given set of conditions.
@@
 
-\section{Residual internal movement (Jitter)}\label{metric:Jitter}
+\section{Residual internal movement (Jitter)}\label{Section:Jitter}
 Due to the dynamic nature of a swarm, maintaining optimum internal movement as in~Figure \ref{methods:StabilityNullVector} is highly unlikely.
@@
 
-\section{Distance metric\label{section:DistanceDynamics}}
+\section{Distance metric}\label{Section:DistanceDynamics}
 The distance metric considers a swarm in terms of how the agents are physically distributed: i.e.
@@
 
-\section{Cohesion/repulsion metric\label{section:MagnitudeDynamics}}
+\section{Cohesion/repulsion metric}\label{Section:MagnitudeDynamics}
 % An agent's \emph{resultant magnitude} is identified by Gazi and Passino~\cite{GP:11} and Barnes et al~\cite{BFV:07} as a `resultant characteristic' of a swarm.
@@
 
-\section{Application: Area filling}\label{metric:ApplicationFloodFilling}
+\section{Application: Area filling}\label{Section:ApplicationFloodFilling}
 
@@
 
-\section{Conclusion\label{metric:MagnitudeDistanceComparison}}
+\section{Conclusion}\label{Section:MagnitudeDistanceComparison}
 %% The two metrics (distance and \emph{cohesion/repulsion}) both measure the structure of a swarm in terms of the distribution of the agents and deviations from a mean.
@@
 
-\section{Future work\label{metric:Future}}
+\section{Future work}\label{Section:FutureWork}
 The new metric can be applied to the analysis of the effects of coordination

### 2018-08-05 paper.tex
MOre Updates

@@
 
-\title{A New Metric for the Analysis of Swarms}
+\title{A New Metric for the Analysis of Swarms that are modelled using potential fields}
 \author{\uppercase{Neil Eliot}\authorrefmark{1},
@@
 \markboth
-{Eliot \headeretal: A New Metric for the Analysis of Swarms}
-{Eliot \headeretal: A New Metric for the Analysis of Swarms}
+{Eliot \headeretal: A New Metric for the Analysis of Swarms that are modelled using potential fields}
+{Eliot \headeretal: A New Metric for the Analysis of Swarms that are modelled using potential fields}
 
@@
   % the intended function of the swarm. 
-  A widely used metric for the analysis of swarm coordination algorithms is
-  based upon the distance between agents. This paper examines this metric and
-  introduces an entirely new metric based upon the field effects that createof the cohesion
-  and repulsion vectors between agents. In comparing and contrasting the two
-  metrics, we find that the \emph{cohesion/repulsion} metric offers a number of
-  advantages over the distance metric. In particular, the
-  \emph{cohesion/repulsion} metric allows the identification of the essential
-  characteristic of a swarm as `expanding', `stable', or `contracting'. This is
-  not possible with the distance metric.  Practical swarming applications where
-  the new metric is useful include area filling and reconnaissance.
+  There are many metrics defined for the analysis of swarm coordination algorithms most of which are
+  based upon the distances between agents or between agents and a fixed point.
+This paper examines a distance-based metric that measures a swarms overall structure using inter-agent distances and introduces an new metric designed to work in a similar way, at the global level, but is based upon the potential field effects that create the swarm's structure.
+The algorithms used to implement the swarming feature are based upon mathematically generated cohesion and repulsion vectors between an agent and its neighbours.
+In comparing and contrasting the two metrics, we find that the \emph{cohesion/repulsion} metric offers a number of advantages over the distance metric.
+In particular, the \emph{cohesion/repulsion} metric allows the identification of the essential characteristic of a swarm as `expanding', `stable', or `contracting'.
+This is not possible with a distance metric.
+Practical swarming applications where the new metric can be applied include area filling and reconnaissance.
 \end{abstract}
@@
 The primary goal of these algorithms is to maintain a stable swarm structure.
-The main focus of this paper is the use any coordination algorithm that are based on \textit{field effects} to produce a metric that allows the swarm to be analysed.
-To allow this paper to present the findings a basic swarm model is introduced and then the two metrics are applied to this model to present a comparison.
-This includes the identification of a swarm's `state'.
-In the remainder of this paper Section~\ref{Section:Related} discusses related work, Section~\ref{Section:SwarmModelling} introduces a basic swarm model for evaluation.
-Section~\ref{Section:StabilityMagnitude} discusses the mathematical model that produces the swarming effect.
+A stable structure is one where the agents in the swarm ``act'' as a single entity.
+The main focus of this paper is the use any coordination algorithms that are based on \textit{potential field effects} to produce a metric that allows the swarm to be analysed.
+To facilitate this analysis a basic swarm model is introduced and then the two metrics are applied to this model to present a comparison.
+This analysis includes the identification of a swarm's ``state''.
+The remainder of this paper is structured as follows Section~\ref{Section:Related} discusses different metrics that are currently used to analyse swarms, Section~\ref{Section:SwarmModelling} introduces a basic swarm model to produce an environment that can be analysed and evaluated using the two metrics.
+Section~\ref{Section:StabilityMagnitude} discusses the mathematical model that produces the swarming effect and how the potential fields are applied.
 Section~\ref{Section::Movement1} discusses how the model is implemented over time to produce movement.
-Section~\ref{Section:StabilityNullVector} discusses the effect of the model on stability.
+Section~\ref{Section:StabilityNullVector} discusses the effect of the model on swarm stability and the effects models have on swarm movements.
 Section~\ref{Section:Jitter} discusses the effect the mathematical model can have upon the structure of a swarm in terms of potentially wasted energy.
 Section~\ref{Section:DistanceDynamics} discusses a distance based metric approach to swarm analysis.
-Section~\ref{Section:MagnitudeDynamics} introduces the new \textit{cohesion/repulsion-based} metric.
+Section~\ref{Section:MagnitudeDynamics} introduces the new \textit{cohesion/repulsion-based} metric and how it is calculated and applied.
 Section~\ref{Section:ApplicationFloodFilling} compares the two metrics using a scenario of a swarm being used to fill an area.
-Section~\ref{Section:MagnitudeDistanceComparison} discusses the results of the experiments and Section~\ref{Section:FutureWork} discuss additional work that needs to be carried out to exploit the additional information that the new metric provides.
+Section~\ref{Section:MagnitudeDistanceComparison} discusses the results of the experiments and Section~\ref{Section:FutureWork} discusses additional work that needs to be carried out to exploit the additional information that the new metric provides.
 
@@
 Many metrics exist for the analysis of swarms.
-These metrics can be categorised as \textit{Centroidal}, \textit{Inter-agent distance} and \textit{magnitude} based.
-The common denominator of the \textit{centroidal} and \text{inter-agent distance} techniques is that they \textit{measure} the agent distribution of a swarm based on distance.
-\textit{Magnitude} based analysis is based upon inter-agent interaction.
+These metrics can be categorised as \textit{Centroidal}, \textit{Agent/Inter-agent distance} and \textit{magnitude} based.
+The common denominator of \textit{centroidal} and \text{Agent/inter-agent distance} metrics is that they \textit{measure} the agent distribution within a swarm based upon distances.
+\textit{Magnitude} based metrics analysis a swarm based upon inter-agent interaction.
 
@@
 They also use the average distance between agents~\cite{WCW:11}.
+Schneider and Wildermuth~\cite{SW:03} used a slightly different approach to measuring performance.
+They identify the difference between a simulated multi-robot environment and a physical implementation and identifying the difference in the paths of the robots based on distance variance.
 
@@
 These effects are \emph{cohesion}, to draw agents closer, and \emph{repulsion} to prevent agents colliding.
-\emph{Field effects} have ranges around an agent and determine the effect other agents have upon it's movement~(Fig.~\ref{methods:FieldEffects}).
+\emph{Potential fields} are ranges around an agent that determine the effect other agents have upon it's movement~(Fig.~\ref{methods:FieldEffects}).
 It is usual for the cohesion field to have a radius $C_b$ which is larger than the repulsion radius $R_b$.
-When an agent ($b'$) moves into the \emph{neighbour field} of an agent ($b$) then $b'$ is said to be a neighbour of $b$ and is subject to cohesion.
-When an agent $b'$ moves into the repulsion field of $b$ then $b$ also subjected to repulsion.
+When an agent ($b'$) moves into the \emph{cohesion field} of an agent ($b$) then $b'$ is said to be a neighbour of $b$ and is subject to cohesion.
+When an agent $b'$ moves into the \textit{repulsion field} of $b$ then $b$ is subjected to repulsion.
 When the repulsion magnitude exceeds the cohesion magnitude the agent will have a tendency to move away from $b'$, i.e.
@@
 When an agent $b$ moves too close to an obstacle, i.e.
-within the obstacle repulsion range $O_b$, it has it's repulsion vector applied and the agent will tend towards moving away from the obstacle.
+within the obstacle repulsion range $O_b$, it has it's repulsion vector applied and the agent will move away from the obstacle.
 
@@
 \section{Inter-agent vector magnitude effect on internal movement}\label{Section:StabilityMagnitude}
-Figure~\ref{methods:Stability5} shows the cohesion and repulsion vector contributions to $v_c(b)$ and $v_r(b)$ due to neighbour $b'$, as given in (\ref{eq:FlyToCentre1}) and (\ref{eq:Repulsion1}).
+Figure~\ref{methods:Stability5} shows the cohesion and repulsion vector contributions to $v_c(b)$ and $v_r(b)$ due to neighbour $b'$, as given in Equation~\ref{eq:FlyToCentre1} and \ref{eq:Repulsion1}.
 Notice that the vectors are along the line of separation $bb'$.
@@
 
-Equation~(\ref{eq:FlyToCentre1}) identifies the resultant cohesion effect of all the neighbours of $b$.
+Equation~\ref{eq:FlyToCentre1} identifies the resultant cohesion effect of all the neighbours of $b$.
 Here, $nbr(b)$ defines the set of all the agents in the swarm that are a neighbour of $b$; a neighbour is any agent that fails within the cohesion field.
@@
 
-Equation~(\ref{eq:Repulsion1}) identifies the resultant repulsion effect of all the neighbours of $b$ where $rep(.)$ returns a set of all agents that are within the repulsion field.
+Equation~\ref{eq:Repulsion1} identifies the resultant repulsion effect of all the neighbours of $b$ where $rep(.)$ returns a set of all agents that are within the repulsion field.
 

### 2018-08-05 paper.tex
More Updates

@@
 
-If a swarm is goal-based, the addition of a \emph{directional vector} will prevent all agents simultaneously producing null vectors as shown in~(Fig.~\ref{concave:VoidPerimeter2}, Equation~(\ref{eq:InterAgentMovementDirectional1})).
+If a swarm is goal-based, the addition of a \emph{directional vector} will prevent all agents simultaneously producing null vectors as shown in~Fig.~\ref{concave:VoidPerimeter2} and Equation~(\ref{eq:InterAgentMovementDirectional1}).
 Here, $k_dv_{d}(b)$ is the calculated directional vector to be applied to an agent.
@@
 
-The mean distance for a swarm is calculated by~(\ref{eq:SwarmStabilityDistance2}). All the inter-agent distances are included for the swarm ($S$).
+The mean distance for a swarm is calculated by~Equation~\ref{eq:SwarmStabilityDistance2}. All the inter-agent distances are included for the swarm ($S$).
 
@@
 However it is not sufficient to give an indication of the internal distribution of the agents.
-The standard deviation clarifies the distribution within the swarm as shown in~(\ref{eq:SwarmStabilityDistance3}).
+The standard deviation clarifies the distribution within the swarm as shown in~Equation~\ref{eq:SwarmStabilityDistance3}.
 
@@
 % The \emph{`cohesion/repulsion metric'} is therefore the scalar value of the \emph{`agent resultant magnitude'} which is altered based upon the difference between the repulsion and cohesion magnitudes~(\ref{eq:CohesionRepulsion}).
-This paper defines the cohesion/repulsion metric scalar for an agent as the length of $v(b)$  as defined (\ref{eq:InterAgentMovement2}): that is, the weighted effects of cohesion and repulsion by neighbours of the agent.
-This number carries a negative sign if repulsion dominatesand a positive sign if cohesion dominates and a positive sign if cohesion donates.
+This paper defines the cohesion/repulsion metric scalar for an agent as the length of $v(b)$  as defined in Equation~\ref{eq:InterAgentMovement2}: that is, the weighted effects of cohesion and repulsion by neighbours of the agent.
+This number carries a negative sign if repulsion dominates and a positive sign if cohesion dominates and a positive sign if cohesion donates.
 This is summarised by~(\ref{eq:CohesionRepulsion}).
@@
 Alvissalim et al.
-(2012) discuss the application of commercially available drones to provide a communication infrastructure across an unmapped disaster area~\cite{AZHMJJM:12}.
+discuss the application of commercially available drones to provide a communication infrastructure across an unmapped disaster area~\cite{AZHMJJM:12}.
 Scheutz and Bauer (2006) use both cohesion and repulsion as a mechanism to create coverage of an area that requires protection in an adversarial environment~\cite{SB:06}.

### 2018-08-05 paper.tex
More Updates

@@
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/StabilityNullVector}
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4cm]{figures/StabilityNullVector}
 {Equilibrium with null vectors.\label{methods:StabilityNullVector}}
@@
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5.5cm]{figures/StabilityNullVector3}
-{(t).\label{concave:VoidPerimeter1}}
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4cm]{figures/StabilityNullVector3}
+{Equilibrium with directional vector (t).\label{concave:VoidPerimeter1}}
 
@@
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/StabilityNullVector2}
-{(t + 1).\label{concave:VoidPerimeter2}}
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4cm]{figures/StabilityNullVector2}
+{Equilibrium with directional vector (t + 1).\label{concave:VoidPerimeter2}}
 
@@
 % \section{Calculating distance internal movement}
-Equation~(\ref{eq:SwarmStabilityDistance1}) computes the mean distance of an agent to its neighbours $nbr(b)$. 
+Equation~(\ref{eq:SwarmStabilityDistance1}) calculates the mean distance of an agent to its neighbours $nbr(b)$. 
 %% The relative position vector generated for an agent $b$ to its neighbour $b'$, $bb'$, is shown in~(\ref{eq:FlyToCentre1}).
@@
 Consider the following: the repulsion field is increased but the internal distances do not change; as a result the \emph{cohesion/repulsion metric} falls below zero.
-This indicates `something' is confining the swarm's distribution as discussed in Section \ref{metric:ApplicationFloodFilling}.
+This indicates `something' is confining the swarm's distribution as discussed in Section \ref{Section:ApplicationFloodFilling}.
 This analysis could be used in identifying effective swarm distribution for the coverage of a sensor array as discussed by Ramaithitima et al.~\cite{RWBK:15}
@@
 algorithms upon a swarm's structure. An application to flood filling has
-been discussed~(Section~\ref{metric:ApplicationFloodFilling}). The metric can also be applied to identifying the effects of 
+been discussed~(Section~\ref{Section:ApplicationFloodFilling}). The metric can also be applied to identifying the effects of 
 `healing' algorithms, e.g. to remove `voids' from a swarm. We plan

### 2018-08-05 paper.tex
More Updates

@@
 
-\title{A New Metric for the Analysis of Swarms that are modelled using potential fields}
+\title{A New Metric for the Analysis of Swarms using Potential Fields}
 \author{\uppercase{Neil Eliot}\authorrefmark{1},
@@
 \markboth
-{Eliot \headeretal: A New Metric for the Analysis of Swarms that are modelled using potential fields}
-{Eliot \headeretal: A New Metric for the Analysis of Swarms that are modelled using potential fields}
+{Eliot \headeretal: A New Metric for the Analysis of Swarms using Potential Fields}
+{Eliot \headeretal: A New Metric for the Analysis of Swarms using Potential Fields}
 

### 2018-08-06 paper.tex
More Updates

@@
 \newtheorem{exampleb}{Example}
-\newcommand{\subsubsubsection}[1]{\medskip\par\emph{#1:}}
+\newcommand{\subsubsubsection}[1]{\medskip\par\textit{#1:}}
 \newcommand{\card}[1]{\vert{#1}\vert}
@@
   % the intended function of the swarm. 
-  There are many metrics defined for the analysis of swarm coordination algorithms most of which are
-  based upon the distances between agents or between agents and a fixed point.
-This paper examines a distance-based metric that measures a swarms overall structure using inter-agent distances and introduces an new metric designed to work in a similar way, at the global level, but is based upon the potential field effects that create the swarm's structure.
+  There are many metrics defined for the analysis of swarm coordination algorithms.
+These metrics are usually based upon the distances between agents, between agents and a fixed point or the resultant vector that \textit{potential field} effects produce.
+This paper examines a distance-based metric that measures a swarms overall structure using inter-agent distances and introduces an new metric that identifies a swarm's state based upon the resultant magnitude of the vectors produced by the inter-agent interactions that create the agent distribution with a swarm's structure.
 The algorithms used to implement the swarming feature are based upon mathematically generated cohesion and repulsion vectors between an agent and its neighbours.
-In comparing and contrasting the two metrics, we find that the \emph{cohesion/repulsion} metric offers a number of advantages over the distance metric.
-In particular, the \emph{cohesion/repulsion} metric allows the identification of the essential characteristic of a swarm as `expanding', `stable', or `contracting'.
-This is not possible with a distance metric.
+In comparing and contrasting the two metrics, we find that the \textit{cohesion/repulsion} metric offers a number of advantages over the distance metric.
+In particular, the \textit{cohesion/repulsion} metric allows the identification of the essential characteristic of a swarm as `expanding', `stable', or `contracting'.
+These state cannot be identified using a distance based metric.
 Practical swarming applications where the new metric can be applied include area filling and reconnaissance.
@@
 \begin{keywords}
-Swarming, Swarm Dynamics, Mobile Sensor Networks, Algorithms.
+Swarming, Swarm Dynamics, Swarm Metric, Algorithms.
 \end{keywords}
@@
 %\section{Introduction\label{methods:SwarmStability}}
-\emph{Swarming} in the animal kingdom of ants, bees, fish and birds for instance has long been studied by scientists.
+\textit{Swarming} in the animal kingdom of ants, bees, fish and birds for instance has long been studied by scientists.
 From these studies mathematical models and algorithms have evolved.
-The models and algorithms have in turn captured the interest of computing scientists who are interested in applying them to large groups of autonomous mobile \emph{agents} (`robots').
+The models and algorithms have in turn captured the interest of computing scientists who are interested in applying them to large groups of autonomous mobile \textit{agents} (`robots').
 The cooperative coordination of these agents can take many forms such as following a set path~\cite{HCS:09}, existing in a static space~\cite{EP:10, GP:02, GP:04} or foraging as a colony~\cite{HER:11, GK:07}.
@@
 These simple rules cause the agents to appear to work cooperatively.
-These rules can also generate movement that does not contribute to the goal of the swarm this movement is a direct result of the agents attempting to move towards an optimum position within the structure.
+These rules can also generate movement that does not contribute to the goal of the swarm, this movement is a direct result of the agents attempting to move towards an optimum position within the structure.
 This internal movement can be identified by analysing the changes in the inter-agent interactions.
@@
 This paper discusses the two metrics and how they differ in their approach to identify these changes.
-The \emph{distance metric} which uses variations in the inter-agent distances is used and discussed by many research teams including Navarro et al.~\cite{NIM:09}, Gazi and Passino~\cite{GP:02,GP:04}, Barnes et al.~\cite{BAF:06} and Schneider and Wildermuth~\cite{SW:03}.
-The new metric uses the resultant magnitudes of the agents' \emph{inter-agent vectors} that are defined by interactions of an agent with its neighbours.
-The metric also identifies when a swarm is expanding or contracting.
+\textit{Distance-based} metrics uses variations in the inter-agent spaces as discussed by many research teams including Navarro et al.~\cite{NIM:09}, Gazi and Passino~\cite{GP:02,GP:04}, Barnes et al.~\cite{BAF:06} and Schneider and Wildermuth~\cite{SW:03}.
+The new metric uses the resultant magnitudes of the agents' \textit{inter-agent vectors} that are defined by interactions of an agent with its neighbours this allows the ``degree'' of a relationship to be evaluated globally and therefore allows a swarm to be identified as expanding or contracting.
 
@@
 A stable structure is one where the agents in the swarm ``act'' as a single entity.
-The main focus of this paper is the use any coordination algorithms that are based on \textit{potential field effects} to produce a metric that allows the swarm to be analysed.
-To facilitate this analysis a basic swarm model is introduced and then the two metrics are applied to this model to present a comparison.
-This analysis includes the identification of a swarm's ``state''.
+The main focus of this paper is the use of a metric that is based upon any coordination algorithms that use \textit{potential fields} to produce a swarming effect.
+To facilitate this analysis a basic swarm model is introduced and then the two metrics are applied to this model to present a comparison of the information provided by a \textit{distance-based} and the new \textit{cohesion/repulsion-based} metric.
+The analysis includes the identification of a swarm's ``state''.
+
 The remainder of this paper is structured as follows Section~\ref{Section:Related} discusses different metrics that are currently used to analyse swarms, Section~\ref{Section:SwarmModelling} introduces a basic swarm model to produce an environment that can be analysed and evaluated using the two metrics.
@@
 Many metrics exist for the analysis of swarms.
-These metrics can be categorised as \textit{Centroidal}, \textit{Agent/Inter-agent distance} and \textit{magnitude} based.
-The common denominator of \textit{centroidal} and \text{Agent/inter-agent distance} metrics is that they \textit{measure} the agent distribution within a swarm based upon distances.
-\textit{Magnitude} based metrics analysis a swarm based upon inter-agent interaction.
+These metrics can be categorised as \textit{Centroidal-based}, \textit{Agent/Inter-agent distance-based} and \textit{vector-based}.
+The common denominator of \textit{centroidal} and \textit{Agent/inter-agent distance} metrics are that they \textit{measure} the agent distribution within a swarm based upon distances.
+\textit{Vector-based} metrics analysis a swarm based upon inter-agent interaction.
 
-\textit{Centroidal} analysis involves identifying the centroid of a swarm, also referred to as the \textit{centre of mass}, as shown in Figure~\ref{com} as a red cross, and calculating distances from that position to each agent.
-The \textit{centroid} is also used as a reference point to determine the path of a swarm.
-Navarro et al.
-use the centroid to produce a \textit{Mean Distance from the Boundary to the Center of the Group} metric~\cite{NIM:09}.
-This metric takes the distance of each agent from the centroid and calculates the mean and deviation to produce their measure.
-Gazi and Passino use this technique when analysing the stability of a swarm based on agent distribution~\cite{GP:11}.
-Gazi and Passino also use the centroid when analysing the stability of a swarm which is designed to maintain its position by analysing the changes in the centroids position over time~\cite{GP:04}.
-Barnes et.
-al.
-use the centroid to analyse the effects of using bivariate normal functions~\cite{BAF:06} and in their work on using potential fields~\cite{BAFVM:06}.
-Gazi and Passino use a centroid based analysis in~\cite{GP:02}~to identify directional changes is agent movements by creating a ``map'' of the potential fields of a swarm in its entirety and overlaying that to the ``centre'' of the swarm.
+\textit{Centroidal} analysis involves identifying the centroid of a swarm, also referred to as the \textit{centre of mass}, as shown in Figure~\ref{com} as a red cross and identified as shown in Equation~\ref{eq:Centroid1}, and calculating the distances from that position to each agent.
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4.5cm]{figures/com}
-{Swarm Centre of Mass.\label{com}}
-
-The centroid of the swarm ($C_s$) is determined by taking the coordinate position of each agent and calculating the mean positions as shown in Equation~\ref{eq:Centroid1} where $b_{coord}$ is each agents coordinates and $S$ is the set of agents that make up the swarm.
+The \textit{centroid} of the swarm ($C_s$) is determined by taking the coordinate position of each agent and calculating the mean positions as shown in Equation~\ref{eq:Centroid1} where $b_{coord}$ is each agents coordinates and $S$ is the set of agents that make up the swarm.
 
@@
 
+ Navarro et al.
+use the centroid to produce a \textit{Mean Distance from the Boundary to the Center of the Group} metric~\cite{NIM:09}.
+This metric takes the distance of each agent from the centroid and calculates the mean and deviation to produce their measure.
+Gazi and Passino use this technique when analysing the stability of a swarm based on agent distribution~\cite{GP:11}.
+The \textit{centroid} is also used as a reference point to determine the path of a swarm.
+Gazi and Passino use the centroid when analysing the stability of a swarm which is designed to maintain its position by analysing the changes in the centroids position over time~\cite{GP:04}.
+Gazi and Passino also use a centroid based analysis in~\cite{GP:02}~to identify directional changes is agent movements by creating a ``map'' of the potential fields of a swarm in its entirety and overlaying that to the ``centre'' of the swarm.
+Barnes et.
+al.
+use the centroid to analyse the effects of using bivariate normal functions in their work on using potential fields~\cite{BAF:06, BAFVM:06}.
+
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4.5cm]{figures/com}
+{Swarm Centre of Mass.\label{com}}
+
 \textit{Inter-Agent Distance} analysis is a basic mechanism for analysing a swarm it to analyse the separation of all of the agents with respect to each agents visible neighbours.
-This is different to using the centroid approach as the measure is based on agent interactions rather than a resultant structure.
+This is different to using the \textit{centroidal} approach as the measure is based on agent interactions rather than a resultant structure which effects the centroid.
 This technique is the most popular approach to analysing a swarms structure.
-Gazi and Passino use a distance based analysis in \cite{GP:04, GP:11} to identify optimum distribution of agents based upon different cohesion and repulsion applications.
+Gazi and Passino use a distance based analysis in~\cite{GP:04, GP:11} to identify optimum distribution of agents based upon different cohesion and repulsion applications.
 Wu et.
 al.
-discuss using several metrics to evaluate emergent behaviours in swarm based on a Viscsek model.
-These include calculating the average number of neighbours for each agent, grouping agents based on numbers of neighbours (to identify clusters).
-They also use the average distance between agents~\cite{WCW:11}.
+discuss using distance-based metrics to evaluate emergent behaviours in swarm based on a Viscsek model.
+These include calculating the average number of neighbours for each agent, grouping agents based on numbers of neighbours (to identify clusters) and the average distance between agents~\cite{WCW:11}.
 Schneider and Wildermuth~\cite{SW:03} used a slightly different approach to measuring performance.
@@
 
-\textit{Magnitude} based analysis uses the potential fields to calculate the ``degree'' of interaction between agent's.
+\textit{Magnitude} based analysis uses the \textit{potential fields} to calculate the ``degree'' of interaction between agents.
 Bennet and McInnes use the resultant relationship between agents to identify equilibrium states when creating structures such as rings and controlling the direction of agents to created specific transformations~\cite{BM:09}.
@@
 The models usually use two field effects to implement the swarming characteristic.
-These effects are \emph{cohesion}, to draw agents closer, and \emph{repulsion} to prevent agents colliding.
-\emph{Potential fields} are ranges around an agent that determine the effect other agents have upon it's movement~(Fig.~\ref{methods:FieldEffects}).
+These effects are \textit{cohesion}, to draw agents closer, and \textit{repulsion} to prevent agents colliding.
+\textit{Potential fields} are ranges around an agent that determine the effect other agents have upon its movement~(Fig.~\ref{methods:FieldEffects}).
 It is usual for the cohesion field to have a radius $C_b$ which is larger than the repulsion radius $R_b$.
-When an agent ($b'$) moves into the \emph{cohesion field} of an agent ($b$) then $b'$ is said to be a neighbour of $b$ and is subject to cohesion.
+When an agent ($b'$) moves into the \textit{cohesion field} of an agent ($b$) then $b'$ is said to be a neighbour of $b$ and is subject to cohesion.
 When an agent $b'$ moves into the \textit{repulsion field} of $b$ then $b$ is subjected to repulsion.
@@
 repelled.
-When an agent $b$ moves too close to an obstacle, i.e.
-within the obstacle repulsion range $O_b$, it has it's repulsion vector applied and the agent will move away from the obstacle.
+Obstacles can be introduced to the model, when an agent $b$ moves too close to an obstacle, i.e.
+within the obstacle repulsion range $O_b$, it has a \textit{repulsion vector} applied to move the agent away from the obstacle.
 
@@
 Using the cohesion and repulsion vectors generated by the relationship of $b$ to its neighbour $b'$ a resultant vector can be calculated.
-This vector creates an agent characteristic that can be used as a metric.
-Summing the vectors creates a resultant vector that affects the agent.
-Summing the vectors also provides an indication of the direction an agent will move based on the relationship.
-This is the \emph{inter-agent vector}.
+Summing the vectors creates a resultant vector that is the overall affects the agent is being subjected to and provides the direction an agent will move based on the relationship.
+This is the \textit{inter-agent vector}.
 
-Equation~(\ref{eq:InterAgentMovement1}) identifies the \emph{inter-agent vector} for agent $b$ with respect to its neighbours.
+Equation~\ref{eq:InterAgentMovement1} identifies the \textit{inter-agent vector} for agent $b$ with respect to its neighbours.
 
@@
 
-If a \emph{weighted} model is used then~(\ref{eq:InterAgentMovement1}) becomes~(\ref{eq:InterAgentMovement2}).
+If a \textit{weighted} model is used then~Equation~\ref{eq:InterAgentMovement1}) becomes~\ref{eq:InterAgentMovement2} where $k_c$ and $k_r$ are weightings to change the effect of each vector as shown in Table~\ref{tab:MetricPhysics1}.
 
@@
 
-where $k_c$ and $k_r$ are weightings to change the effect of each vector as shown in Table~\ref{tab:MetricPhysics1}.
-
-%% From here throughout sections \ref{metric:MagnitudeDynamics2} and \ref{metric:StabilityNullVector} imagine agent $b$ has just a single neighbour $b'$ and consider the effect of $b'$ on $v(b)$, the \emph{inter-agent vector} of $b$.
-%% The relationship between the agents is that: \emph{the vector for $b$ with respect to $b^{'}$, and the vector for $b^{'}$ with respect to $b$ are on the same line}~(\ref{methods:Stability5}).
+%% From here throughout sections \ref{metric:MagnitudeDynamics2} and \ref{metric:StabilityNullVector} imagine agent $b$ has just a single neighbour $b'$ and consider the effect of $b'$ on $v(b)$, the \textit{inter-agent vector} of $b$.
+%% The relationship between the agents is that: \textit{the vector for $b$ with respect to $b^{'}$, and the vector for $b^{'}$ with respect to $b$ are on the same line}~(\ref{methods:Stability5}).
 %% If we consider an agent ($b$) as having only one neighbour ($b'$) then using the formulae defined in chapter~\ref{chapter:methods} the vectors $v_c(b)$~(\ref{eq:FlyToCentre1}) and $v_r(b)$~(\ref{eq:Repulsion1}) are the magnitudes of the vectors that are generated by the field effects between the two agents.
@@
 
+\section{Internal movement and the null vector}\label{Section:StabilityNullVector}
+When the two vectors (cohesion and repulsion) have magnitudes that are equal and opposite they produce a null vector, indicating that two agents are optimally spaced for a given set of conditions.
+Although the agents are at an optimum position when resultants are zero it does not mean the swarm is optimally distributed.
+If a swarm is in a confined space it is possible for an optimum position to be created where the vector magnitude is affected by a compression effect.
+This phenomenon is used in the identification of the emergent behaviour of \textit{area flooding}.
+
+If we consider the equilibrium state (Fig.~\ref{methods:Stability4}) the resultant $v(b)=(0,0)$.
+A null vector cannot be normalised to produce a directional vector ($\hat{v} = \frac{v}{|v|}$ if $v\neq0$; $0$ if $v=0$) The effect is that the agent will remain stationary.
+If all agent pairs are in this condition the swarm will stop moving~(Fig.~\ref{methods:StabilityNullVector}).
+
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4cm]{figures/StabilityNullVector}
+{Equilibrium with null vectors.\label{methods:StabilityNullVector}}
+
+%% \begin{figure}[H]
+%% \begin{center}
+%% \includegraphics[width=5cm]{figures/StabilityNullVector}
+%% \end{center}
+%% \caption{Equilibrium with null vectors} \label{methods:StabilityNullVector}
+%% \end{figure}
+
+Due to the movement of agents being based only on neighbour interactions this situation is very rare.
+The residual motion that persists in a swarm is the background `noise' or `jitter' that an algorithm creates.
+
+If a swarm is goal-based, the addition of a \textit{directional vector} will prevent all agents simultaneously producing null vectors as shown in~Fig.~\ref{concave:VoidPerimeter2} and Equation~\ref{eq:InterAgentMovementDirectional1}.
+Here, $k_dv_{d}(b)$ is the calculated directional vector to be applied to an agent.
+
+\begin{equation}\label{eq:InterAgentMovementDirectional1}
+v(b) = k_cv_{c}(b) + k_rv_{r}(b) + k_dv_{d}(b)
+\end{equation}
+
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4cm]{figures/StabilityNullVector3}
+{Equilibrium with directional vector (t).\label{concave:VoidPerimeter1}}
+
+%% \begin{figure}[H]
+%% \begin{center}
+%% 	 \includegraphics[width=5.5cm]{figures/StabilityNullVector3}
+%% \end{center}
+%% \label{concave:VoidPerimeter1}
+%% \caption{(t)}
+%% \end{figure}
+
+\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4cm]{figures/StabilityNullVector2}
+{Equilibrium with directional vector (t + 1).\label{concave:VoidPerimeter2}}
+
+%% \begin{figure}[H]
+%% \begin{center}
+%% 	 \includegraphics[width=6cm]{figures/StabilityNullVector2}
+%% \end{center}
+%% \caption{(t + 1)}
+%% \label{concave:VoidPerimeter2}
+%% \end{figure}
+
+\section{Residual internal movement (Jitter)}\label{Section:Jitter}
+Due to the dynamic nature of a swarm, maintaining optimum internal movement as in~Figure \ref{methods:StabilityNullVector} is highly unlikely.
+The agent pairs may fluctuate between the 4 states~(Fig.~\ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}); These transitions are \textit{jitter}.
+The degree to which this variation occurs can be measured using either the change in distance between the agents, or the change in the magnitude $P(b)$~(Equation~\ref{eq:CohesionRepulsion}) between the agents.
+Jitter arises as motion to maintain the structure of a swarm.
+A swarm coordination algorithm that reduces jitter is generally desirable.
+
 \section{Modelling movement}\label{Section::Movement1}
 Each agent within a swarm calculates its \textit{movement-direction vector} based on its \textit{interaction} and \textit{destination} vectors.
-The \textit{movement vector}~($b_{pos}$) is calculated using the unit \textit{movement-direction vector} of (\ref{eq:InterAgentMovement2}) multiplied by the time elapsed~($t$) in the system and the speed characteristic of the agent~($s_b$).
+The \textit{movement vector}~($b_{pos}$) is calculated using the unit \textit{movement-direction vector} (Equation~\ref{eq:InterAgentMovementDirectional1}) multiplied by the time elapsed~($t$) in the system and the speed characteristic of the agent~($s_b$).
 
@@
 
-The increment in the location of agent $b$ over time interval $t$ is shown in~(\ref{eq:AgentMovement}) where $s_b$ is the speed of agent~$b$.
+The increment in the location of agent $b$ over time interval $t$ is shown in Equation~\ref{eq:AgentMovement} where $s_b$ is the speed of agent~$b$.
 Note that \string^ is the equivalent of $\hat{v} = \frac{v}{|v|}$ the normalised vector.
@@
 The repulsion and cohesion vectors are generated for an agent through the interaction of their field effects~(Fig.~\ref{methods:FieldEffects}).
-There are a limited number of interactions that can occur; These are illustrated in Fig.~\ref{methods:Stability2}, \ref{methods:Stability3}, \ref{methods:Stability4} and \ref{methods:StabilityNullVector}.
+There are a limited number of interactions that can occur; These are illustrated in Fig.~\ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}.
 
@@
 
-Figure~\ref{methods:Stability4} shows two agents close together with $|k_cv_c|~=~|k_rv_r|$ the resultant vector is a \emph{null vector} and the agents have no overall influence upon each other due to the magnitude of the resultant vector being zero.
+Figure~\ref{methods:Stability4} shows two agents close together with $|k_cv_c|~=~|k_rv_r|$ the resultant vector is a \textit{null vector} and the agents have no overall influence upon each other due to the magnitude of the resultant vector being zero.
 Table~\ref{tab:SampleEquilibrium} is an extract of data from the simulator.
@@
 
-\section{Internal movement and the null vector}\label{Section:StabilityNullVector}
-When the two vectors (cohesion and repulsion) have magnitudes that are equal and opposite they produce a null vector, indicating that two agents are optimally spaced for a given set of conditions.
-Although the agents are at an optimum position when resultants are zero it does not mean the swarm is optimally distributed.
-If a swarm is in a confined space it is possible for an optimum position to be created where the vector magnitude is affected by a compression effect.
-This phenomenon is used in the identification of the emergent behaviour of \emph{area flooding}.
-
-If we consider the equilibrium state (Fig.~\ref{methods:Stability4}) the resultant $v(b)=(0,0)$.
-A null vector cannot be normalised to produce a directional vector ($\hat{v} = \frac{v}{|v|}$ if $v\neq0$; $0$ if $v=0$) The effect is that the agent will remain stationary.
-If all agent pairs are in this condition the swarm will stop moving~(Fig.~\ref{methods:StabilityNullVector}).
-
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4cm]{figures/StabilityNullVector}
-{Equilibrium with null vectors.\label{methods:StabilityNullVector}}
-
-%% \begin{figure}[H]
-%% \begin{center}
-%% \includegraphics[width=5cm]{figures/StabilityNullVector}
-%% \end{center}
-%% \caption{Equilibrium with null vectors} \label{methods:StabilityNullVector}
-%% \end{figure}
-
-Due to the movement of agents being based only on neighbour interactions this situation is very rare.
-The residual motion that persists in a swarm is the background `noise' or `jitter' that an algorithm creates.
-
-If a swarm is goal-based, the addition of a \emph{directional vector} will prevent all agents simultaneously producing null vectors as shown in~Fig.~\ref{concave:VoidPerimeter2} and Equation~(\ref{eq:InterAgentMovementDirectional1}).
-Here, $k_dv_{d}(b)$ is the calculated directional vector to be applied to an agent.
-
-\begin{equation}\label{eq:InterAgentMovementDirectional1}
-v(b) = k_cv_{c}(b) + k_rv_{r}(b) + k_dv_{d}(b)
-\end{equation}
-
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4cm]{figures/StabilityNullVector3}
-{Equilibrium with directional vector (t).\label{concave:VoidPerimeter1}}
-
-%% \begin{figure}[H]
-%% \begin{center}
-%% 	 \includegraphics[width=5.5cm]{figures/StabilityNullVector3}
-%% \end{center}
-%% \label{concave:VoidPerimeter1}
-%% \caption{(t)}
-%% \end{figure}
-
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4cm]{figures/StabilityNullVector2}
-{Equilibrium with directional vector (t + 1).\label{concave:VoidPerimeter2}}
-
-%% \begin{figure}[H]
-%% \begin{center}
-%% 	 \includegraphics[width=6cm]{figures/StabilityNullVector2}
-%% \end{center}
-%% \caption{(t + 1)}
-%% \label{concave:VoidPerimeter2}
-%% \end{figure}
-
-\section{Residual internal movement (Jitter)}\label{Section:Jitter}
-Due to the dynamic nature of a swarm, maintaining optimum internal movement as in~Figure \ref{methods:StabilityNullVector} is highly unlikely.
-The agent pairs may fluctuate between the 4 states~(Fig.~\ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}); These transitions are \emph{jitter}.
-The degree to which this variation occurs can be measured using either the change in distance between the agents, or the change in the magnitude $P(b)$~(\ref{eq:CohesionRepulsion}) between the agents.
-Jitter arises as motion to maintain the structure of a swarm.
-A swarm coordination algorithm that reduces jitter is generally desirable.
-
 \section{Distance metric}\label{Section:DistanceDynamics}
@@
 
-Navarro and Fernando describe a \emph{mean distance error} metric that is based on the variations in distances between inter-agent spaces~\cite{NIM:09}.
+Navarro and Fernando describe a \textit{mean distance error} metric that is based on the variations in distances between inter-agent spaces~\cite{NIM:09}.
 This is the same as the standard deviation of the distance based internal movement metric.
@@
 %% The distance based internal movement is measured by identifying the mean length of the vectors between an agent and its neighbours.
-As with the \emph{agent resultant magnitude} a coordination algorithm produces `jitter' which is the variations from the mean.
-In the case of the distance based metric the jitter is identified by the changes in the distances rather than the changes in vector magnitude (\emph{agent resultant magnitude}).
+As with the \textit{agent resultant magnitude} a coordination algorithm produces `jitter' which is the variations from the mean.
+In the case of the distance based metric the jitter is identified by the changes in the distances rather than the changes in vector magnitude (\textit{agent resultant magnitude}).
 The distance metric is the mean and the standard deviation `jitter' of the inter-agent distances.
@@
 % \section{Calculating distance internal movement}
-Equation~(\ref{eq:SwarmStabilityDistance1}) calculates the mean distance of an agent to its neighbours $nbr(b)$. 
+Equation~\ref{eq:SwarmStabilityDistance1} calculates the mean distance of an agent to its neighbours $nbr(b)$. 
 %% The relative position vector generated for an agent $b$ to its neighbour $b'$, $bb'$, is shown in~(\ref{eq:FlyToCentre1}).
@@
 \section{Cohesion/repulsion metric}\label{Section:MagnitudeDynamics}
-% An agent's \emph{resultant magnitude} is identified by Gazi and Passino~\cite{GP:11} and Barnes et al~\cite{BFV:07} as a `resultant characteristic' of a swarm.
+% An agent's \textit{resultant magnitude} is identified by Gazi and Passino~\cite{GP:11} and Barnes et al~\cite{BFV:07} as a `resultant characteristic' of a swarm.
 The sum of the cohesion and repulsion vectors between an agent $b$ and a neighbour $b'$ produce a resultant vector.
 
-% The cohesion/repulsion internal movement (\emph{cohesion/repulsion metric}) is measured by identifying the balance between the cohesion and repulsion vectors between agents as shown in~(\ref{eq:CohesionRepulsion}).
-This is derived from the \emph{inter-agent cohesion/repulsion vector} as shown in~(\ref{eq:InterAgentMovementDirectional1}).
-`Jitter' in the case of the agent resultant \emph{cohesion/repulsion metric} is measured as the deviation created by the inter-agent relationships.
-The identification of this deviation produces the clarifying component of the agent \emph{cohesion/repulsion metric}.
+% The cohesion/repulsion internal movement (\textit{cohesion/repulsion metric}) is measured by identifying the balance between the cohesion and repulsion vectors between agents as shown in~(\ref{eq:CohesionRepulsion}).
+This is derived from the \textit{inter-agent cohesion/repulsion vector} as shown in~(\ref{eq:InterAgentMovementDirectional1}).
+`Jitter' in the case of the agent resultant \textit{cohesion/repulsion metric} is measured as the deviation created by the inter-agent relationships.
+The identification of this deviation produces the clarifying component of the agent \textit{cohesion/repulsion metric}.
 % %% The magnitude based metric identifies the resultant internal magnitude (vector magnitude) through the cohesive and repulsive magnitudes that exist between each agent pair in a swarm. 
@@
 
-% This paper will refer to the scalar resultant magnitude as the \emph{`cohesion/repulsion metric'} of the relationship.
+% This paper will refer to the scalar resultant magnitude as the \textit{`cohesion/repulsion metric'} of the relationship.
 The `state' of a swarm is therefore the effect the environmental constraints and algorithms have upon the agent resultant magnitude.
-The agent's \emph{`cohesion/repulsion metric'} can be considered part of the `quality' measure for a swarm's performance.
+The agent's \textit{`cohesion/repulsion metric'} can be considered part of the `quality' measure for a swarm's performance.
 
-% The resultant \emph{cohesion/repulsion metric}, which is a scalar, is generated by comparing the length of the cohesion and repulsion vector magnitudes.
+% The resultant \textit{cohesion/repulsion metric}, which is a scalar, is generated by comparing the length of the cohesion and repulsion vector magnitudes.
 If the cohesion magnitude is larger than the repulsion vector magnitude the resultant length is considered to be positive.
@@
 
-% The \emph{`cohesion/repulsion metric'} is therefore the scalar value of the \emph{`agent resultant magnitude'} which is altered based upon the difference between the repulsion and cohesion magnitudes~(\ref{eq:CohesionRepulsion}).
+% The \textit{`cohesion/repulsion metric'} is therefore the scalar value of the \textit{`agent resultant magnitude'} which is altered based upon the difference between the repulsion and cohesion magnitudes~(\ref{eq:CohesionRepulsion}).
 This paper defines the cohesion/repulsion metric scalar for an agent as the length of $v(b)$  as defined in Equation~\ref{eq:InterAgentMovement2}: that is, the weighted effects of cohesion and repulsion by neighbours of the agent.
 This number carries a negative sign if repulsion dominates and a positive sign if cohesion dominates and a positive sign if cohesion donates.
-This is summarised by~(\ref{eq:CohesionRepulsion}).
+This is summarised in Equation~\ref{eq:CohesionRepulsion}.
 
@@
 The disorganised stage is the initial random deployment of the agents.
-If the agent \emph{agent resultant magnitude} is positive then the swarm is exhibiting a tendency to contract and this indicates the swarm is a cohesive entity.
+If the agent \textit{agent resultant magnitude} is positive then the swarm is exhibiting a tendency to contract and this indicates the swarm is a cohesive entity.
 This could also be described as the swarm being `sticky' as the agents bias is to `pull' towards each other.
 
-The \emph{cohesion/repulsion metric} on its own does not give a complete measure of a swarm's internal state.
-There needs to be a qualifying component to the metric that identifies the degree of deviation in the resultant magnitude, this is the \emph{jitter}.
+The \textit{cohesion/repulsion metric} on its own does not give a complete measure of a swarm's internal state.
+There needs to be a qualifying component to the metric that identifies the degree of deviation in the resultant magnitude, this is the \textit{jitter}.
 The smaller the deviation the more uniform the structure of the swarm.
@@
 
-The \emph{cohesion/repulsion metric} provides a view of the swarm's state through the balance between the repulsive and the cohesive vectors that are being applied to each agent.
+The \textit{cohesion/repulsion metric} provides a view of the swarm's state through the balance between the repulsive and the cohesive vectors that are being applied to each agent.
 The deviation component identifies the degree to which the swarm has stabilised.
-The ideal status for inter-agent interactions would be for the agents to have a (\emph{cohesion/repulsion metric}) of zero or above.
+The ideal status for inter-agent interactions would be for the agents to have a (\textit{cohesion/repulsion metric}) of zero or above.
 This would indicate that the agents are distributed such that they are at their distribution limit (outer most range of the cohesion field) or at a level that causes the agents to `pull' together.
@@
 % \section{Cohesion/repulsion internal movement model}\label{Section:StabilityModel}
-Section~\ref{Section:StabilityMagnitude}~discussed calculation of resultant cohesion~(\ref{eq:FlyToCentre1}) and resultant repulsion~(\ref{eq:Repulsion1}) vectors for each agent deriving a combined \emph{cohesion/repulsion vector} using~(\ref{eq:InterAgentMovement2}).
+Section~\ref{Section:StabilityMagnitude}~discussed calculation of resultant cohesion~(Equation~\ref{eq:FlyToCentre1}) and resultant repulsion~(Equation~\ref{eq:Repulsion1}) vectors for each agent deriving a combined \textit{cohesion/repulsion vector} using as shown in Equation~\ref{eq:InterAgentMovement2}.
 
 %% This value represents the overall `potential' of an agent.
-This vector when normalised produces a component of the \emph{movement-destination vector} for a swarm.
+This vector when normalised produces a component of the \textit{movement-destination vector} for a swarm.
 If the total cohesion/repulsion vector is zero (null vector) then the agent will not move.
-$M(b)$ is the total \emph{cohesion/repulsion vector} for agent $b$ defined by:
+$M(b)$ is the total \textit{cohesion/repulsion vector} for agent $b$ defined by:
 %% 
@@
 
-Now, we define for each agent $b$ a \emph{cohesion/repulsion scalar}, $P(b)$, to measure the influence of the neighbours of $b$ on $b$ as cohesive (positive) or repulsive (negative); this is~(\ref{eq:CohesionRepulsion}).
+Now, we define for each agent $b$ a \textit{cohesion/repulsion scalar}, $P(b)$, to measure the influence of the neighbours of $b$ as cohesive (positive) or repulsive (negative); shown in Equation~\ref{eq:CohesionRepulsion}.
 
@@
 
-Although it is possible for $v(b)$ to be null there could still be variation in the constituent components. The variation calculation (standard deviation) is shown below.
+Although it is possible for $v(b)$ to be null there could still be variation in the constituent components.
+The variation calculation (standard deviation) as shown in Equation~\ref{eq:SwarmStabilityQuotientT}.
 
@@
 %% \section{Standard deviation in agent cohesion/repulsion metric}\label{Section:VarianceInPotential}
-%% To improve the metric clarification in terms of the deviation from the \emph{cohesion/repulsion metric} norm is calculated.
+%% To improve the metric clarification in terms of the deviation from the \textit{cohesion/repulsion metric} norm is calculated.
 The standard deviation of the entire swarm  is shown in Equation \ref{eq:SwarmStabilityMetricT}.
 
-The standard deviation associated with this mean is calculated as~(\ref{eq:SwarmStabilityQuotientT}).
+The standard deviation associated with this mean is calculated as Equation~\ref{eq:SwarmStabilityQuotientT}.
 %%  where $\sigma_p(S)$ is the standard deviation and $\mu_p(S)$ is the mean at the same point in time.
@@
 
-The metric for the internal movement is this pair of numbers, the mean and standard deviation of the swarm's internal \emph{cohesion/repulsion}.
+The metric for the internal movement is this pair of numbers, the mean and standard deviation of the swarm's internal \textit{cohesion/repulsion}.
 The pair $\mu_p(S)$, $\sigma_p(S)$ may be written informally as:
@@
 \section{Conclusion}\label{Section:MagnitudeDistanceComparison}
-%% The two metrics (distance and \emph{cohesion/repulsion}) both measure the structure of a swarm in terms of the distribution of the agents and deviations from a mean.
-The \emph{cohesion/repulsion} metric also allows the state of the swarm to be identified.
+%% The two metrics (distance and \textit{cohesion/repulsion}) both measure the structure of a swarm in terms of the distribution of the agents and deviations from a mean.
+The \textit{cohesion/repulsion} metric also allows the state of the swarm to be identified.
 %% 
-%% The distance metric is based upon the physical \emph{distribution} of the agents and the cohesion/repulsion metric is based upon the logical \emph{interaction} of the agents.
+%% The distance metric is based upon the physical \textit{distribution} of the agents and the cohesion/repulsion metric is based upon the logical \textit{interaction} of the agents.
 The distance-based metric measures swarm structure in terms of position of agents.
@@
 
-The \emph{cohesion/repulsion} metric provides an indication of a swarm's potential future movement. This is independent of the physical distribution of the agents within the swarm.
+The \textit{cohesion/repulsion} metric provides an indication of a swarm's potential future movement. This is independent of the physical distribution of the agents within the swarm.
 
 The two metrics together provide a deeper evaluation of a swarm's structure and `potential' than either of the metrics alone.
-Consider the following: the repulsion field is increased but the internal distances do not change; as a result the \emph{cohesion/repulsion metric} falls below zero.
+Consider the following: the repulsion field is increased but the internal distances do not change; as a result the \textit{cohesion/repulsion metric} falls below zero.
 This indicates `something' is confining the swarm's distribution as discussed in Section \ref{Section:ApplicationFloodFilling}.

### 2018-08-06 paper.tex
Updates

@@
   There are many metrics defined for the analysis of swarm coordination algorithms.
-These metrics are usually based upon the distances between agents, between agents and a fixed point or the resultant vector that \textit{potential field} effects produce.
+These metrics are usually based upon the distances between agents, the distance between agents and a fixed point or the resultant vector that \textit{potential field} effects produce.
 This paper examines a distance-based metric that measures a swarms overall structure using inter-agent distances and introduces an new metric that identifies a swarm's state based upon the resultant magnitude of the vectors produced by the inter-agent interactions that create the agent distribution with a swarm's structure.

### 2018-08-06 paper.tex
More Updates

@@
 introduce the concept of a \textit{rejection potential} in an attempt to generate a kinetic energy score for a swarm~\cite{WCW:11}.
+
+This paper builds upon the work relating to the \textit{Magnitude-based} metric analysis as discussed by Bennet and McInnes and Wu et al.
+by expanding the analysis to the entire swarm and converting the magnitudes to a scalar value so that the metric produces an arbitrary value that can be used to compare swarm algorithms.
 

### 2018-08-07 paper.tex
Almost completed updates. Structure to update.

@@
 These metrics are usually based upon the distances between agents, the distance between agents and a fixed point or the resultant vector that \textit{potential field} effects produce.
-This paper examines a distance-based metric that measures a swarms overall structure using inter-agent distances and introduces an new metric that identifies a swarm's state based upon the resultant magnitude of the vectors produced by the inter-agent interactions that create the agent distribution with a swarm's structure.
+This paper examines a distance-based metric that measures a swarms overall structure using inter-agent distances and introduces an new metric that identifies a swarm's state based upon the resultant magnitude of the vectors produced by the inter-agent interactions that create the agent distribution within a swarm's structure.
 The algorithms used to implement the swarming feature are based upon mathematically generated cohesion and repulsion vectors between an agent and its neighbours.
@@
 A stable structure is one where the agents in the swarm ``act'' as a single entity.
-The main focus of this paper is the use of a metric that is based upon any coordination algorithms that use \textit{potential fields} to produce a swarming effect.
+The main focus of this paper is the use of a metric that can be applied to any coordination algorithms that use \textit{potential fields} to produce a swarming effect.
 To facilitate this analysis a basic swarm model is introduced and then the two metrics are applied to this model to present a comparison of the information provided by a \textit{distance-based} and the new \textit{cohesion/repulsion-based} metric.
 The analysis includes the identification of a swarm's ``state''.
+The metric is then applied to a flood filling scenario to highlight a possible use of the new metric.
 
@@
 
-\section{Residual internal movement (Jitter)}\label{Section:Jitter}
-Due to the dynamic nature of a swarm, maintaining optimum internal movement as in~Figure \ref{methods:StabilityNullVector} is highly unlikely.
+% \section{Residual internal movement (Jitter)}\label{Section:Jitter}
+% Due to the dynamic nature of a swarm, maintaining optimum internal movement as in~Figure \ref{methods:StabilityNullVector} is highly unlikely.
 The agent pairs may fluctuate between the 4 states~(Fig.~\ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}); These transitions are \textit{jitter}.
@@
 Figure~\ref{methods:Stability1} shows two agents within each other's cohesion fields but sufficiently distant to be outside of the repulsion fields.
-In this case $|k_cv_c| > 0$ and $|k_rv_r| = 0$: the result is the agent's resultant magnitude $v(b)$ causes the agent to move towards is neighbour $b'$.
+In this case $|k_cv_c| > 0$ and $|k_rv_r| = 0$: the result is the agent's resultant magnitude $v(b)$ causes the agent to move towards its neighbour $b'$.
 Likewise the neighbour's resultant magnitude $v(b')$ will cause it to move towards $b$.
 Table~\ref{tab:SampleReplusion0} shows the repulsion magnitude with a value of 0.
-The only influence on the agent pairs are cohesive vectors.
+The only influence on the agent pair are cohesive vectors.
 
@@
 
-\section{Distance metric}\label{Section:DistanceDynamics}
-The distance metric considers a swarm in terms of how the agents are physically distributed: i.e.
+\section{Inter-Agent Distance metric}\label{Section:DistanceDynamics}
+The inter-agent distance metric considers a swarm in terms of how the agents are physically distributed: i.e.
 only the inter-agent distances and their standard deviation are considered.
-The standard deviation indicates the extent to which the swarm is out of balance and will move to re-balance itself.
+The standard deviation indicates the extent to which the swarm is out of balance and therefore the degree of movement required to re-balance the swarm structure.
 If the standard deviation is zero then all the agents are evenly spaced.
@@
 
-The distance-based metric for the internal distribution of the agents is the pair consisting of $\mu_d(S)$, $\sigma_d(S)$. This can be written informally as:
+The distance-based metric for the internal distribution of the agents is therefore $\mu_d(S)$ and $\sigma_d(S)$. This can be written informally as:
 
@@
 This paper defines the cohesion/repulsion metric scalar for an agent as the length of $v(b)$  as defined in Equation~\ref{eq:InterAgentMovement2}: that is, the weighted effects of cohesion and repulsion by neighbours of the agent.
-This number carries a negative sign if repulsion dominates and a positive sign if cohesion dominates and a positive sign if cohesion donates.
+This number carries a negative sign if repulsion dominates and a positive sign if cohesion dominates.
 This is summarised in Equation~\ref{eq:CohesionRepulsion}.
@@
 could improve the granularity of analysis, e.g. when using sensor arrays at
-thr frontier of the swarm.
+the frontier of the swarm.
 However, this approach leads to swarms of agents having heterogeneous field

### 2018-08-07 paper.tex
Updated
Minor

@@
 \textit{Swarming} in the animal kingdom of ants, bees, fish and birds for instance has long been studied by scientists.
-From these studies mathematical models and algorithms have evolved.
+From these studies mathematical models and algorithms have evolved such as those produced by Reynolds~\cite{REY:87}.
 The models and algorithms have in turn captured the interest of computing scientists who are interested in applying them to large groups of autonomous mobile \textit{agents} (`robots').
@@
 Section~\ref{Section:StabilityNullVector} discusses the effect of the model on swarm stability and the effects models have on swarm movements.
-Section~\ref{Section:Jitter} discusses the effect the mathematical model can have upon the structure of a swarm in terms of potentially wasted energy.
+% Section~\ref{Section:Jitter} discusses the effect the mathematical model can have upon the structure of a swarm in terms of potentially wasted energy. 
 Section~\ref{Section:DistanceDynamics} discusses a distance based metric approach to swarm analysis.

### 2018-08-20 paper.tex
Minor revisions

@@
 These metrics are usually based upon the distances between agents, the distance between agents and a fixed point or the resultant vector that \textit{potential field} effects produce.
-This paper examines a distance-based metric that measures a swarms overall structure using inter-agent distances and introduces an new metric that identifies a swarm's state based upon the resultant magnitude of the vectors produced by the inter-agent interactions that create the agent distribution within a swarm's structure.
-The algorithms used to implement the swarming feature are based upon mathematically generated cohesion and repulsion vectors between an agent and its neighbours.
+This paper examines a distance-based metric that measures a swarm's overall structure using inter-agent distances and introduces a new metric that identifies a swarm's state based upon the resultant magnitude of the vectors produced by the inter-agent interactions that create the agent distribution within a swarm's structure.
+The algorithms used to implement the swarming feature are based upon cohesion and repulsion vectors between an agent and its neighbours.
 In comparing and contrasting the two metrics, we find that the \textit{cohesion/repulsion} metric offers a number of advantages over the distance metric.
 In particular, the \textit{cohesion/repulsion} metric allows the identification of the essential characteristic of a swarm as `expanding', `stable', or `contracting'.
-These state cannot be identified using a distance based metric.
-Practical swarming applications where the new metric can be applied include area filling and reconnaissance.
+These state cannot be identified using a distance-based metric.
+Practical swarming applications where the new metric can be applied advantageously include area filling and reconnaissance.
 \end{abstract}
@@
 \begin{keywords}
-Swarming, Swarm Dynamics, Swarm Metric, Algorithms.
+Swarming, Swarm Dynamics, Swarm Metric.
 \end{keywords}
@@
 \textit{Distance-based} metrics uses variations in the inter-agent spaces as discussed by many research teams including Navarro et al.~\cite{NIM:09}, Gazi and Passino~\cite{GP:02,GP:04}, Barnes et al.~\cite{BAF:06} and Schneider and Wildermuth~\cite{SW:03}.
-The new metric uses the resultant magnitudes of the agents' \textit{inter-agent vectors} that are defined by interactions of an agent with its neighbours this allows the ``degree'' of a relationship to be evaluated globally and therefore allows a swarm to be identified as expanding or contracting.
+The new metric uses the resultant magnitudes of the agents' \textit{inter-agent vectors} that are defined by interactions of an agent with its neighbours this allows the ``degree'' of a relationship to be evaluated globally and therefore allows a swarm to be identified as ``expanding'', ``stable'' or ``contracting''.
 
 Many coordination algorithms exist to control the movement of agents within a swarm.
-The primary goal of these algorithms is to maintain a stable swarm structure.
+The primary goal of these algorithms are to maintain a stable swarm structure.
 A stable structure is one where the agents in the swarm ``act'' as a single entity.
 The main focus of this paper is the use of a metric that can be applied to any coordination algorithms that use \textit{potential fields} to produce a swarming effect.
-To facilitate this analysis a basic swarm model is introduced and then the two metrics are applied to this model to present a comparison of the information provided by a \textit{distance-based} and the new \textit{cohesion/repulsion-based} metric.
+Most swarming algorithms define a set of equations that calculate the repulsion and cohesion effects.
+These two equations are the basis of the new metric.
+
+To facilitate the analysis of the new metric a basic swarm model is introduced and then two metrics are applied to this model to present a comparison of the information provided by a \textit{distance-based} and the new \textit{cohesion/repulsion-based} metric.
 The analysis includes the identification of a swarm's ``state''.
@@
 % Section~\ref{Section:Jitter} discusses the effect the mathematical model can have upon the structure of a swarm in terms of potentially wasted energy. 
-Section~\ref{Section:DistanceDynamics} discusses a distance based metric approach to swarm analysis.
+Section~\ref{Section:DistanceDynamics} discusses a distance-based metric approach to swarm analysis.
 Section~\ref{Section:MagnitudeDynamics} introduces the new \textit{cohesion/repulsion-based} metric and how it is calculated and applied.
@@
 
-\textit{Centroidal} analysis involves identifying the centroid of a swarm, also referred to as the \textit{centre of mass}, as shown in Figure~\ref{com} as a red cross and identified as shown in Equation~\ref{eq:Centroid1}, and calculating the distances from that position to each agent.
-
-The \textit{centroid} of the swarm ($C_s$) is determined by taking the coordinate position of each agent and calculating the mean positions as shown in Equation~\ref{eq:Centroid1} where $b_{coord}$ is each agents coordinates and $S$ is the set of agents that make up the swarm.
+\textit{Centroidal} analysis involves identifying the centroid of a swarm, also referred to as the \textit{centre of mass}, as shown in Figure~\ref{com} as a red cross and defined in Equation~\ref{eq:Centroid1}.
 
 \begin{equation}\label{eq:Centroid1}
-C_s = \frac{1}{\card{S}}\mathlarger{\sum_{b \in S}} b_{coord}
+  \bar{r} = \frac{1}{\card{S}}\mathlarger{\sum_{b \in S}} r_b
 \end{equation}
 
+Here, ($\bar{r}$) denotes the position vector of the swarm's \textit{centroid}; $r_b$ is the position of each agent($b$) and $S$ is the set of agents that make up the swarm.
+
 Navarro et al.
-use the centroid to produce a \textit{Mean Distance from the Boundary to the Center of the Group} metric~\cite{NIM:09}.
+use the centroid to produce a "Mean Distance from the Boundary to the Center of the Group" metric~\cite{NIM:09}.
 This metric takes the distance of each agent from the centroid and calculates the mean and deviation to produce their measure.
@@
 The \textit{centroid} is also used as a reference point to determine the path of a swarm.
-Gazi and Passino use the centroid when analysing the stability of a swarm which is designed to maintain its position by analysing the changes in the centroids position over time~\cite{GP:04}.
-Gazi and Passino also use a centroid based analysis in~\cite{GP:02}~to identify directional changes is agent movements by creating a ``map'' of the potential fields of a swarm in its entirety and overlaying that to the ``centre'' of the swarm.
+Gazi and Passino use the centroid when analysing the stability of a swarm which is designed to maintain its position by analysing the changes in the centroid's position over time~\cite{GP:04}.
+Gazi and Passino also use a centroid-based analysis in~\cite{GP:02}~to identify directional changes in agent movement by creating a ``map'' of the potential fields of a swarm in its entirety and overlaying that to the ``centre'' of the swarm.
 Barnes et.
@@
 
-\textit{Inter-Agent Distance} analysis is a basic mechanism for analysing a swarm it to analyse the separation of all of the agents with respect to each agents visible neighbours.
+\textit{Inter-Agent Distance} analysis is a basic mechanism for analysing a swarm in terms of separation of all of the agents with respect to each agent's visible neighbours.
 This is different to using the \textit{centroidal} approach as the measure is based on agent interactions rather than a resultant structure which effects the centroid.
 This technique is the most popular approach to analysing a swarms structure.
-Gazi and Passino use a distance based analysis in~\cite{GP:04, GP:11} to identify optimum distribution of agents based upon different cohesion and repulsion applications.
+Gazi and Passino use a distance-based analysis in~\cite{GP:04, GP:11} to identify optimum distribution of agents based upon different cohesion and repulsion applications.
 Wu et.
 al.
-discuss using distance-based metrics to evaluate emergent behaviours in swarm based on a Viscsek model.
+discuss using distance-based metrics to evaluate emergent behaviours in swarm based on a Viscsek model~\cite{WCW:11}.
 These include calculating the average number of neighbours for each agent, grouping agents based on numbers of neighbours (to identify clusters) and the average distance between agents~\cite{WCW:11}.
-Schneider and Wildermuth~\cite{SW:03} used a slightly different approach to measuring performance.
-They identify the difference between a simulated multi-robot environment and a physical implementation and identifying the difference in the paths of the robots based on distance variance.
+Schneider and Wildermuth~\cite{SW:03} use a slightly different approach to measuring performance.
+They identify the difference in the paths of robots in a simulated versus a physical implementation and calculate a distance variance between them.
 
-\textit{Magnitude} based analysis uses the \textit{potential fields} to calculate the ``degree'' of interaction between agents.
+\textit{Magnitude-based} analysis uses the \textit{potential fields} to calculate the ``degree'' of interaction between agents.
 Bennet and McInnes use the resultant relationship between agents to identify equilibrium states when creating structures such as rings and controlling the direction of agents to created specific transformations~\cite{BM:09}.
 Wu et al.
-introduce the concept of a \textit{rejection potential} in an attempt to generate a kinetic energy score for a swarm~\cite{WCW:11}.
+introduce the concept of a \textit{rejection potential} in an attempt to generate a "kinetic energy" score for a swarm~\cite{WCW:11}.
 
 This paper builds upon the work relating to the \textit{Magnitude-based} metric analysis as discussed by Bennet and McInnes and Wu et al.
-by expanding the analysis to the entire swarm and converting the magnitudes to a scalar value so that the metric produces an arbitrary value that can be used to compare swarm algorithms.
+by expanding the analysis to the entire swarm, deriving a scalar value can be used in the analysis of swarm behaviour.
 
@@
 The expansion increases until, due to boundary compression, the swarm is unable to expand further.
-In an extreme case the expansion will result in a set of field effects that create a mesh based swarm structure rather than the desired hexagonal structure that the field effects should create.
+In an extreme case the expansion will result in a set of field effects that create a mesh-based swarm structure rather than the desired hexagonal structure that the field effects should create.
 Identifying the \textit{cohesion/repulsion metric} falling below zero indicates area saturation has occurred.

### 2018-08-20 paper.tex
Minor Revisions

@@
 only the inter-agent distances and their standard deviation are considered.
-The standard deviation indicates the extent to which the swarm is out of balance and therefore the degree of movement required to re-balance the swarm structure.
+The standard deviation indicates the extent to which the swarm is out of balance and therefore the degree of movement required to rebalance the swarm structure.
 If the standard deviation is zero then all the agents are evenly spaced.
@@
 Navarro and Fernando describe a \textit{mean distance error} metric that is based on the variations in distances between inter-agent spaces~\cite{NIM:09}.
-This is the same as the standard deviation of the distance based internal movement metric.
+This is the same as the standard deviation of the distance-based internal movement metric.
 

### 2018-08-21 paper.tex
Minor

@@
 Many metrics exist for the analysis of swarms.
-These metrics can be categorised as \textit{Centroidal-based}, \textit{Agent/Inter-agent distance-based} and \textit{vector-based}.
+These metrics can be categorised as \textit{Centroidal-based}, \textit{Agent/Inter-agent distance-based} and \textit{Magnitude-based}.
 The common denominator of \textit{centroidal} and \textit{Agent/inter-agent distance} metrics are that they \textit{measure} the agent distribution within a swarm based upon distances.
-\textit{Vector-based} metrics analysis a swarm based upon inter-agent interaction.
+\textit{Magnitude-based} metrics analysis a swarm based upon inter-agent interaction.
 
@@
 %% The distance metric is based upon the physical \textit{distribution} of the agents and the cohesion/repulsion metric is based upon the logical \textit{interaction} of the agents.
-The distance-based metric measures swarm structure in terms of position of agents.
+Distance-based metrics measure a swarm's structure in terms of the position of agents.
+This is independent of the algorithm that produces the swarm's structure.
 It provides an analysis at a point in time but does not allow the identification of the future behaviour of a swarm as `expanding', `stable' or `contracting'.
@@
 
-The \textit{cohesion/repulsion} metric provides an indication of a swarm's potential future movement. This is independent of the physical distribution of the agents within the swarm.
+The \textit{cohesion/repulsion} metric is based upon the algorithms that have produced the structure and is therefore an analysis of how the algorithms are performing rather than the structure presented.
+The new metric also provides an indication of a swarm's potential future movement.
+This is independent of the physical distribution of the agents within the swarm.
+The metric also provides the same information as the distance metric in terms of agent distribution when the field effects are consistent across all agents i.e.
+monolithic field effects.
+This is shown in Fig.
+\ref{emerge:FIELDFILL-MAG} and \ref{emerge:FIELDFILL-DIST} by the profiles of the two graphs being the same.
 

### 2018-08-21 paper.tex
More Minors

@@
 This is independent of the algorithm that produces the swarm's structure.
-It provides an analysis at a point in time but does not allow the identification of the future behaviour of a swarm as `expanding', `stable' or `contracting'.
+It provides an analysis of agent distribution at a point in time but does not allow the identification of the future behaviour of a swarm as `expanding', `stable' or `contracting'.
 
@@
 
-The \textit{cohesion/repulsion} metric is based upon the algorithms that have produced the structure and is therefore an analysis of how the algorithms are performing rather than the structure presented.
+The new \textit{cohesion/repulsion} metric is based upon the algorithms that have produced the structure and is therefore an analysis of how the algorithms are performing rather than the structure presented.
 The new metric also provides an indication of a swarm's potential future movement.
@@
 \ref{emerge:FIELDFILL-MAG} and \ref{emerge:FIELDFILL-DIST} by the profiles of the two graphs being the same.
+In addition the new metric provides a mechanism for evaluating swarms that use different algorithms to create a swarming effect and allows the algorithms inter-agent interactions to be compared rather than the resultant structure the algorithm creates.
 
-The two metrics together provide a deeper evaluation of a swarm's structure and `potential' than either of the metrics alone.
-Consider the following: the repulsion field is increased but the internal distances do not change; as a result the \textit{cohesion/repulsion metric} falls below zero.
+The two metrics can be used together to provide a deeper evaluation of a swarm's structure and `potential'.
+Consider the following scenario: the repulsion field is increased but the internal distances do not change; as a result the \textit{cohesion/repulsion metric} falls below zero.
 This indicates `something' is confining the swarm's distribution as discussed in Section \ref{Section:ApplicationFloodFilling}.
@@
 APZDAMC:09, AZDPS:11}. A possible adaptation of this area of research is to
-investigate the reduction of the field effects (cohesion and repulsion) of
-agents at the perimeter, so increasing agent density in this region. Increasing
-the density would allow increased coverage by the swarm at its perimeter. This
+investigate the reduction of the field effects (cohesion and repulsion) on the perimeter
+so as to increase the agent density in this region of the swam. This change in the swarm's structure
 could improve the granularity of analysis, e.g. when using sensor arrays at

### 2018-10-16 paper.tex
Updates 16-10-2018

@@
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Images/kenda}}]{David Kendall}
-has been a Senior Lecturer in Computing at Northumbria University since 1989, except for a brief period as a Lecturer in Computer Science at Durham University (2001-02).
-Previously, he was a Research Associate at Newcastle University (1987-89), following experience as a Senior Software Engineer in industry (1983-86).
-He has an M.A.
-degree in Literae Humaniores from Oxford University, where he studied at New College.
-He received M.Sc.
+received the M.A.
+degree in literae humaniores from New College, Oxford University, and the M.Sc.
 and Ph.D.
-degrees in Computing Science from Newcastle University.
-His research interests are in the area of formal modelling and analysis of embedded systems.
+degrees in computing science from Newcastle University.
+He was a Research Associate at Newcastle University from 1987 to 1989, following experience as a Senior Software Engineer in industry from 1983 to 1986.
+He has been a Senior Lecturer in computing at Northumbria University since 1989, except for a brief period as a Lecturer in computer science at Durham University from 2001 to 2002.
+His research interests are in the areas of formal methods, embedded systems, distributed systems, security, and privacy.
 \end{IEEEbiography}

### 2018-10-17 paper.tex
Spelling and grammar revised up to end of Section II

@@
   % the intended function of the swarm. 
-  There are many metrics defined for the analysis of swarm coordination algorithms.
-These metrics are usually based upon the distances between agents, the distance between agents and a fixed point or the resultant vector that \textit{potential field} effects produce.
-This paper examines a distance-based metric that measures a swarm's overall structure using inter-agent distances and introduces a new metric that identifies a swarm's state based upon the resultant magnitude of the vectors produced by the inter-agent interactions that create the agent distribution within a swarm's structure.
-The algorithms used to implement the swarming feature are based upon cohesion and repulsion vectors between an agent and its neighbours.
-In comparing and contrasting the two metrics, we find that the \textit{cohesion/repulsion} metric offers a number of advantages over the distance metric.
-In particular, the \textit{cohesion/repulsion} metric allows the identification of the essential characteristic of a swarm as `expanding', `stable', or `contracting'.
-These state cannot be identified using a distance-based metric.
-Practical swarming applications where the new metric can be applied advantageously include area filling and reconnaissance.
+  There are many metrics defined for the analysis of swarm coordination
+  algorithms. These metrics are usually based upon the distances between
+  agents, the distance between agents and a fixed point, or the resultant vectors
+  that \textit{potential field} effects produce. This paper examines a
+  distance-based metric that measures a swarm's overall structure using
+  inter-agent distances. It introduces a new metric that identifies a swarm's
+  state based upon the resultant magnitude of the vectors produced by the
+  agent interactions that create the agent distribution within the swarm's
+  structure. The algorithms used to implement the swarming feature are based
+  upon cohesion and repulsion vectors between an agent and its neighbours. In
+  comparing and contrasting the two metrics, we find that the
+  \textit{cohesion/repulsion} metric offers a number of advantages over the
+  distance metric. In particular, the \textit{cohesion/repulsion} metric allows
+  the identification of the essential characteristic of a swarm as `expanding',
+  `stable', or `contracting'. These states cannot be identified using a
+  distance-based metric. Practical swarming applications where the new metric
+  can be applied advantageously include area-filling and reconnaissance.
 \end{abstract}
@@
 %\section{Introduction\label{methods:SwarmStability}}
-\textit{Swarming} in the animal kingdom of ants, bees, fish and birds for instance has long been studied by scientists.
-From these studies mathematical models and algorithms have evolved such as those produced by Reynolds~\cite{REY:87}.
-The models and algorithms have in turn captured the interest of computing scientists who are interested in applying them to large groups of autonomous mobile \textit{agents} (`robots').
-The cooperative coordination of these agents can take many forms such as following a set path~\cite{HCS:09}, existing in a static space~\cite{EP:10, GP:02, GP:04} or foraging as a colony~\cite{HER:11, GK:07}.
-One of the attributes of swarms that has captured the interest of scientists is that the models and algorithms used to coordinate them are generally sets of simple rules.
-These simple rules cause the agents to appear to work cooperatively.
-These rules can also generate movement that does not contribute to the goal of the swarm, this movement is a direct result of the agents attempting to move towards an optimum position within the structure.
-This internal movement can be identified by analysing the changes in the inter-agent interactions.
-Measuring this unnecessary movement allows the effectiveness of a swarming algorithm to be evaluated.
-If the internal movement can be reduced then the stability of the swarm is improved and more of the swarms energy is focused on the goal.
-Reducing the internal movement also provides a more stable platform for the deployment of sensor arrays.
-Both the metrics presented in this paper allow this evaluation to be carried out when all the sensor ranges are the same however the distance metric is not able to identify the internal movement accurately when the sensor ranges of each agent vary producing irregular inter-agent distances even though the interactions at the vector level indicate that the agents are in fact stable and optimum.
-This paper discusses the two metrics and how they differ in their approach to identify these changes.
-\textit{Distance-based} metrics uses variations in the inter-agent spaces as discussed by many research teams including Navarro et al.~\cite{NIM:09}, Gazi and Passino~\cite{GP:02,GP:04}, Barnes et al.~\cite{BAF:06} and Schneider and Wildermuth~\cite{SW:03}.
-The new metric uses the resultant magnitudes of the agents' \textit{inter-agent vectors} that are defined by interactions of an agent with its neighbours this allows the ``degree'' of a relationship to be evaluated globally and therefore allows a swarm to be identified as ``expanding'', ``stable'' or ``contracting''.
 
-Many coordination algorithms exist to control the movement of agents within a swarm.
-The primary goal of these algorithms are to maintain a stable swarm structure.
-A stable structure is one where the agents in the swarm ``act'' as a single entity.
-The main focus of this paper is the use of a metric that can be applied to any coordination algorithms that use \textit{potential fields} to produce a swarming effect.
-Most swarming algorithms define a set of equations that calculate the repulsion and cohesion effects.
-These two equations are the basis of the new metric.
+\textit{Swarming} in the animal kingdom of ants, bees, fish and birds, for
+instance, has long been studied by scientists. From these studies, mathematical
+models and algorithms have evolved, such as those produced by
+Reynolds~\cite{REY:87}. The models and algorithms have in turn captured the
+interest of computing scientists, who are interested in applying them to large
+groups of autonomous mobile \textit{agents} (`robots'). The cooperative
+coordination of these agents can take many forms, such as following a set
+path~\cite{HCS:09}, existing in a static space~\cite{EP:10, GP:02, GP:04} or
+foraging as a colony~\cite{HER:11, GK:07}. One of the attributes of swarms that
+has captured the interest of scientists is that the models and algorithms used
+to coordinate them are generally sets of simple rules. These simple rules cause
+the agents to appear to work cooperatively. Often, the rules can also generate
+movement that does not contribute to the goal of the swarm. This movement is a
+direct result of the agents attempting to move towards an optimum position
+within the structure. This internal movement can be identified by analysing the
+changes in the inter-agent interactions. Measuring this unnecessary movement
+allows the effectiveness of a swarming algorithm to be evaluated. If the
+internal movement can be reduced, then the stability of the swarm is improved
+and more of the swarm's energy is focused on the goal. Reducing the internal
+movement also provides a more stable platform for the deployment of sensor
+arrays. Both metrics, presented in this paper, allow this evaluation to be
+carried out when all the sensor ranges are the same. However, the distance metric
+is not able to identify the internal movement accurately when the sensor ranges
+of each agent vary, producing irregular inter-agent distances, even though the
+interactions at the vector level indicate that the agents are in fact stable
+and optimum. This paper discusses the two metrics and how they differ in their
+approach to identifying these changes. \textit{Distance-based} metrics use
+variations in the inter-agent spaces, as discussed by many research teams
+including Navarro et al.~\cite{NIM:09}, Gazi and Passino~\cite{GP:02,GP:04},
+Barnes et al.~\cite{BAF:06} and Schneider and Wildermuth~\cite{SW:03}. The new
+metric uses the resultant magnitudes of the agents' \textit{inter-agent
+vectors} that are defined by interactions of an agent with its neighbours. This
+allows the ``degree'' of a relationship to be evaluated globally and therefore
+allows a swarm to be identified as ``expanding'', ``stable'' or
+``contracting''.
 
-To facilitate the analysis of the new metric a basic swarm model is introduced and then two metrics are applied to this model to present a comparison of the information provided by a \textit{distance-based} and the new \textit{cohesion/repulsion-based} metric.
-The analysis includes the identification of a swarm's ``state''.
-The metric is then applied to a flood filling scenario to highlight a possible use of the new metric.
+Many coordination algorithms exist to control the movement of agents within a
+swarm. The primary goal of these algorithms is to maintain a stable swarm
+structure. A stable structure is one where the agents in the swarm ``act'' as a
+single entity. The main focus of this paper is on metrics that can be
+applied to any coordination algorithms that use \textit{potential fields} to
+produce a swarming effect. Most swarming algorithms define a set of equations
+that calculate the repulsion and cohesion effects. These equations are the
+basis of the new metric.
 
-The remainder of this paper is structured as follows Section~\ref{Section:Related} discusses different metrics that are currently used to analyse swarms, Section~\ref{Section:SwarmModelling} introduces a basic swarm model to produce an environment that can be analysed and evaluated using the two metrics.
-Section~\ref{Section:StabilityMagnitude} discusses the mathematical model that produces the swarming effect and how the potential fields are applied. 
-Section~\ref{Section::Movement1} discusses how the model is implemented over time to produce movement.
-Section~\ref{Section:StabilityNullVector} discusses the effect of the model on swarm stability and the effects models have on swarm movements.
-% Section~\ref{Section:Jitter} discusses the effect the mathematical model can have upon the structure of a swarm in terms of potentially wasted energy. 
-Section~\ref{Section:DistanceDynamics} discusses a distance-based metric approach to swarm analysis.
-Section~\ref{Section:MagnitudeDynamics} introduces the new \textit{cohesion/repulsion-based} metric and how it is calculated and applied.
-Section~\ref{Section:ApplicationFloodFilling} compares the two metrics using a scenario of a swarm being used to fill an area.
-Section~\ref{Section:MagnitudeDistanceComparison} discusses the results of the experiments and Section~\ref{Section:FutureWork} discusses additional work that needs to be carried out to exploit the additional information that the new metric provides.
+To facilitate the analysis of the new metric, a basic swarm model is introduced
+and then two metrics are applied to this model to present a comparison of the
+information provided by a \textit{distance-based} and the new
+\textit{cohesion/repulsion-based} metric. The analysis includes the
+identification of a swarm's ``state''. The metric is then applied to an 
+area-filling scenario to highlight a possible use of the new metric.
+
+The remainder of this paper is structured as follows:
+Section~\ref{Section:Related} discusses different metrics that are currently
+used to analyse swarms; Section~\ref{Section:SwarmModelling} introduces a basic
+swarm model to produce an environment that can be analysed and evaluated using
+the two metrics; Section~\ref{Section:StabilityMagnitude} discusses the
+mathematical model that produces the swarming effect and how the potential
+fields are applied; Section~\ref{Section::Movement1} discusses how the model is
+implemented over time to produce movement;
+Section~\ref{Section:StabilityNullVector} discusses the effect of the model on
+swarm stability and the effects models have on swarm movements; 
+% Section~\ref{Section:Jitter} discusses the effect the mathematical model can
+% have upon the structure of a swarm in terms of potentially wasted energy. 
+Section~\ref{Section:DistanceDynamics} discusses a distance-based metric for
+swarm analysis; Section~\ref{Section:MagnitudeDynamics} introduces the new
+\textit{cohesion/repulsion-based} metric, showing how it is calculated and
+applied; Section~\ref{Section:ApplicationFloodFilling} compares the two metrics
+using an area-filling scenario;
+Section~\ref{Section:MagnitudeDistanceComparison} discusses the results of
+experiments, and, finally, Section~\ref{Section:FutureWork} discusses
+additional work that needs to be carried out to exploit the additional
+information that the new metric provides.
 
@@
 
-Many metrics exist for the analysis of swarms.
-These metrics can be categorised as \textit{Centroidal-based}, \textit{Agent/Inter-agent distance-based} and \textit{Magnitude-based}.
-The common denominator of \textit{centroidal} and \textit{Agent/inter-agent distance} metrics are that they \textit{measure} the agent distribution within a swarm based upon distances.
-\textit{Magnitude-based} metrics analysis a swarm based upon inter-agent interaction.
+Many metrics exist for the analysis of swarms. These metrics can be categorised
+as \textit{centroidal-based}, \textit{agent/inter-agent distance-based} and
+\textit{magnitude-based}. The common denominator of \textit{centroidal} and
+\textit{agent/inter-agent distance} metrics is that they \textit{measure} the
+agent distribution within a swarm based upon distances.
+\textit{Magnitude-based} metrics analyse a swarm based upon inter-agent
+interaction.
 
-\textit{Centroidal} analysis involves identifying the centroid of a swarm, also referred to as the \textit{centre of mass}, as shown in Figure~\ref{com} as a red cross and defined in Equation~\ref{eq:Centroid1}.
+\textit{Centroidal} analysis involves identifying the centroid of a swarm, also
+referred to as the \textit{centre of mass}, as shown in Figure~\ref{com} as a
+red cross and defined in Equation~\ref{eq:Centroid1}.
 
@@
 
-Here, ($\bar{r}$) denotes the position vector of the swarm's \textit{centroid}; $r_b$ is the position of each agent($b$) and $S$ is the set of agents that make up the swarm.
+Here, $\bar{r}$ denotes the position vector of the swarm's \textit{centroid};
+$r_b$ is the position of each agent, $b$, and $S$ is the set of agents that make
+up the swarm.
 
-Navarro et al.
-use the centroid to produce a "Mean Distance from the Boundary to the Center of the Group" metric~\cite{NIM:09}.
-This metric takes the distance of each agent from the centroid and calculates the mean and deviation to produce their measure.
-Gazi and Passino use this technique when analysing the stability of a swarm based on agent distribution~\cite{GP:11}.
-The \textit{centroid} is also used as a reference point to determine the path of a swarm.
-Gazi and Passino use the centroid when analysing the stability of a swarm which is designed to maintain its position by analysing the changes in the centroid's position over time~\cite{GP:04}.
-Gazi and Passino also use a centroid-based analysis in~\cite{GP:02}~to identify directional changes in agent movement by creating a ``map'' of the potential fields of a swarm in its entirety and overlaying that to the ``centre'' of the swarm.
-Barnes et.
-al.
-use the centroid to analyse the effects of using bivariate normal functions in their work on using potential fields~\cite{BAF:06, BAFVM:06}.
+Navarro et al. use the centroid to produce a "Mean Distance from the Boundary
+to the Center of the Group" metric~\cite{NIM:09}. This metric takes the
+distance of each agent from the centroid and calculates the mean and deviation
+to produce their measure. Gazi and Passino use this technique when analysing
+the stability of a swarm based on agent distribution~\cite{GP:11}. The
+\textit{centroid} is also used as a reference point to determine the path of a
+swarm. Gazi and Passino use the centroid when analysing the stability of a
+swarm which is designed to maintain its position, by analysing the changes in
+the centroid's position over time~\cite{GP:04}. Gazi and Passino also use a
+centroid-based analysis in~\cite{GP:02}~to identify directional changes in
+agent movement by creating a ``map'' of the potential fields of a swarm in its
+entirety and overlaying that to the ``centre'' of the swarm.  Barnes et. al.
+use the centroid to analyse the effects of using bivariate normal functions in
+their work on using potential fields~\cite{BAF:06, BAFVM:06}.
 
@@
 
-\textit{Inter-Agent Distance} analysis is a basic mechanism for analysing a swarm in terms of separation of all of the agents with respect to each agent's visible neighbours.
-This is different to using the \textit{centroidal} approach as the measure is based on agent interactions rather than a resultant structure which effects the centroid.
-This technique is the most popular approach to analysing a swarms structure.
-Gazi and Passino use a distance-based analysis in~\cite{GP:04, GP:11} to identify optimum distribution of agents based upon different cohesion and repulsion applications.
-Wu et.
-al.
-discuss using distance-based metrics to evaluate emergent behaviours in swarm based on a Viscsek model~\cite{WCW:11}.
-These include calculating the average number of neighbours for each agent, grouping agents based on numbers of neighbours (to identify clusters) and the average distance between agents~\cite{WCW:11}.
-Schneider and Wildermuth~\cite{SW:03} use a slightly different approach to measuring performance.
-They identify the difference in the paths of robots in a simulated versus a physical implementation and calculate a distance variance between them.
+\textit{Inter-agent distance} analysis is a basic mechanism for analysing a
+swarm in terms of separation of all of the agents with respect to each agent's
+visible neighbours. This is different to using the \textit{centroidal} approach,
+as the measure is based on agent interactions rather than a resultant structure
+which affects the centroid. This technique is the most popular approach to
+analysing a swarm's structure. Gazi and Passino use a distance-based analysis
+in~\cite{GP:04, GP:11} to identify optimum distribution of agents based upon
+different cohesion and repulsion applications. Wu et. al. discuss using
+distance-based metrics to evaluate emergent behaviours in swarms based on a
+Viscsek model~\cite{WCW:11}. These include calculating the average number of
+neighbours for each agent, grouping agents based on numbers of neighbours (to
+identify clusters), and calculating the average distance between agents~\cite{WCW:11}.
+Schneider and Wildermuth~\cite{SW:03} use a slightly different approach to
+measuring performance. They identify the difference in the paths of robots in a
+simulated versus a physical implementation and calculate a distance variance
+between them.
 
-\textit{Magnitude-based} analysis uses the \textit{potential fields} to calculate the ``degree'' of interaction between agents.
-Bennet and McInnes use the resultant relationship between agents to identify equilibrium states when creating structures such as rings and controlling the direction of agents to created specific transformations~\cite{BM:09}.
-Wu et al.
-introduce the concept of a \textit{rejection potential} in an attempt to generate a "kinetic energy" score for a swarm~\cite{WCW:11}.
+\textit{Magnitude-based} analysis uses the \textit{potential fields} to
+calculate the ``degree'' of interaction between agents. Bennet and McInnes use
+the resultant relationship between agents to identify equilibrium states when
+creating structures such as rings and controlling the direction of agents to
+create specific transformations~\cite{BM:09}. Wu et al. introduce the concept
+of a \textit{rejection potential} in an attempt to generate a "kinetic energy"
+score for a swarm~\cite{WCW:11}.
 
-This paper builds upon the work relating to the \textit{Magnitude-based} metric analysis as discussed by Bennet and McInnes and Wu et al.
-by expanding the analysis to the entire swarm, deriving a scalar value can be used in the analysis of swarm behaviour.
+This paper builds upon the work relating to the \textit{magnitude-based} metric
+analysis, as discussed by Bennet and McInnes, and Wu et al., by expanding the
+analysis to the entire swarm and deriving a scalar value that can be used in the
+analysis of swarm behaviour.
 
@@
 \section{Swarm modelling}\label{Section:SwarmModelling}
-Currently, much swarm research uses field effects as the method of modelling inter-agent interactions~\cite{BAF:06, BAFVM:06, BM:09, APZDAMC:09, GP:02, GP:04, GP:04a, GP:05, GP:11, MYP:09}.
-The models usually use two field effects to implement the swarming characteristic.
-These effects are \textit{cohesion}, to draw agents closer, and \textit{repulsion} to prevent agents colliding.
-\textit{Potential fields} are ranges around an agent that determine the effect other agents have upon its movement~(Fig.~\ref{methods:FieldEffects}).
-It is usual for the cohesion field to have a radius $C_b$ which is larger than the repulsion radius $R_b$.
-When an agent ($b'$) moves into the \textit{cohesion field} of an agent ($b$) then $b'$ is said to be a neighbour of $b$ and is subject to cohesion.
-When an agent $b'$ moves into the \textit{repulsion field} of $b$ then $b$ is subjected to repulsion.
-When the repulsion magnitude exceeds the cohesion magnitude the agent will have a tendency to move away from $b'$, i.e.
-repelled.
-Obstacles can be introduced to the model, when an agent $b$ moves too close to an obstacle, i.e.
-within the obstacle repulsion range $O_b$, it has a \textit{repulsion vector} applied to move the agent away from the obstacle.
+Currently, much swarm research uses field effects as the method of modelling
+inter-agent interactions~\cite{BAF:06, BAFVM:06, BM:09, APZDAMC:09, GP:02,
+GP:04, GP:04a, GP:05, GP:11, MYP:09}. The models usually use two field effects
+to implement the swarming characteristic. These effects are \textit{cohesion},
+to draw agents closer, and \textit{repulsion} to prevent agents colliding.
+\textit{Potential fields} are ranges around an agent that determine the effect
+other agents have upon its movement~(Fig.~\ref{methods:FieldEffects}). It is
+usual for the cohesion field to have a radius $C_b$ which is larger than the
+repulsion radius $R_b$. When an agent $b'$ moves into the \textit{cohesion
+field} of an agent $b$, then $b'$ is said to be a neighbour of $b$, and $b$ is
+subject to cohesion. When an agent $b'$ moves into the \textit{repulsion field}
+of $b$, then $b$ is subject to repulsion. When the repulsion magnitude exceeds
+the cohesion magnitude, the agent $b$ will have a tendency to move away from
+$b'$, i.e. to be repelled. Obstacles can be introduced into the model. When an
+agent $b$ moves too close to an obstacle, i.e. within the obstacle repulsion
+range $O_b$, it has a \textit{repulsion vector} applied, giving it a tendency
+to move away from the obstacle.
 
@@
 \section{Inter-agent vector magnitude effect on internal movement}\label{Section:StabilityMagnitude}
-Figure~\ref{methods:Stability5} shows the cohesion and repulsion vector contributions to $v_c(b)$ and $v_r(b)$ due to neighbour $b'$, as given in Equation~\ref{eq:FlyToCentre1} and \ref{eq:Repulsion1}.
-Notice that the vectors are along the line of separation $bb'$.
-\\
+Figure~\ref{methods:Stability5} shows the cohesion and repulsion vector
+contributions of neighbour $b'$ to $v_c(b)$ and $v_r(b)$, as given in
+Equations~\ref{eq:FlyToCentre1} and \ref{eq:Repulsion1}. Notice that the vectors
+are along the line of separation $bb'$. \\
 
@@
 
-Equation~\ref{eq:FlyToCentre1} identifies the resultant cohesion effect of all the neighbours of $b$.
-Here, $nbr(b)$ defines the set of all the agents in the swarm that are a neighbour of $b$; a neighbour is any agent that fails within the cohesion field.
+Equation~\ref{eq:FlyToCentre1} identifies the resultant cohesion effect of all
+the neighbours of $b$. Here, $nbr(b)$ defines the set of all the agents in the
+swarm that are a neighbour of $b$; a neighbour is any agent that falls within
+the cohesion field.
 
@@
 
-Equation~\ref{eq:Repulsion1} identifies the resultant repulsion effect of all the neighbours of $b$ where $rep(.)$ returns a set of all agents that are within the repulsion field.
+Equation~\ref{eq:Repulsion1} identifies the resultant repulsion effect of all
+the neighbours of $b$, where $rep(b)$ returns the set of all agents that are
+within the repulsion field of $b$.
 
@@
 
-Using the cohesion and repulsion vectors generated by the relationship of $b$ to its neighbour $b'$ a resultant vector can be calculated.
-Summing the vectors creates a resultant vector that is the overall affects the agent is being subjected to and provides the direction an agent will move based on the relationship.
-This is the \textit{inter-agent vector}.
-
-Equation~\ref{eq:InterAgentMovement1} identifies the \textit{inter-agent vector} for agent $b$ with respect to its neighbours.
+Using the cohesion and repulsion vectors generated by the relationship of $b$
+to its neighbours, a resultant vector can be calculated that summarises the
+effects to which the agent is being subjected, and provides the direction in
+which the agent will move, based on this relationship. This is the
+\textit{inter-agent vector}.  Equation~\ref{eq:InterAgentMovement1} identifies
+the \textit{inter-agent vector} for agent $b$ with respect to its neighbours.
 
@@
 
-If a \textit{weighted} model is used then~Equation~\ref{eq:InterAgentMovement1}) becomes~\ref{eq:InterAgentMovement2} where $k_c$ and $k_r$ are weightings to change the effect of each vector as shown in Table~\ref{tab:MetricPhysics1}.
+If a \textit{weighted} model is used then~Equation~\ref{eq:InterAgentMovement1}
+is replaced by~Equation~\ref{eq:InterAgentMovement2}, where $k_c$ and $k_r$ are
+weightings to change the effect of each vector, as exemplified in
+Table~\ref{tab:MetricPhysics1}.
 
@@
 \section{Internal movement and the null vector}\label{Section:StabilityNullVector}
-When the two vectors (cohesion and repulsion) have magnitudes that are equal and opposite they produce a null vector, indicating that two agents are optimally spaced for a given set of conditions.
-Although the agents are at an optimum position when resultants are zero it does not mean the swarm is optimally distributed.
-If a swarm is in a confined space it is possible for an optimum position to be created where the vector magnitude is affected by a compression effect.
-This phenomenon is used in the identification of the emergent behaviour of \textit{area flooding}.
 
-If we consider the equilibrium state (Fig.~\ref{methods:Stability4}) the resultant $v(b)=(0,0)$.
-A null vector cannot be normalised to produce a directional vector ($\hat{v} = \frac{v}{|v|}$ if $v\neq0$; $0$ if $v=0$) The effect is that the agent will remain stationary.
-If all agent pairs are in this condition the swarm will stop moving~(Fig.~\ref{methods:StabilityNullVector}).
+When the two vectors (cohesion and repulsion) have magnitudes that are equal
+and opposite they produce a null vector, indicating that two agents are
+optimally spaced for a given set of conditions. Although the agents are at an
+optimum position when resultants are zero it does not mean the swarm is
+optimally distributed. If a swarm is in a confined space it is possible for an
+optimum position to be created where the vector magnitude is affected by a
+compression effect. This phenomenon is used in the identification of the
+emergent behaviour of \textit{area filling}.
+
+If we consider the equilibrium state (Fig.~\ref{methods:Stability4}) the
+resultant $v(b)=(0,0)$. A null vector cannot be normalised to produce a
+directional vector ($\hat{v} = \frac{v}{|v|}$ if $v\neq0$; $0$ if $v=0$) The
+effect is that the agent will remain stationary. If all agent pairs are in this
+condition the swarm will stop moving~(Fig.~\ref{methods:StabilityNullVector}).
 
@@
 
-Due to the movement of agents being based only on neighbour interactions this situation is very rare.
-The residual motion that persists in a swarm is the background `noise' or `jitter' that an algorithm creates.
+Due to the movement of agents being based only on neighbour interactions this
+situation is very rare. The residual motion that persists in a swarm is the
+background `noise' or `jitter' that an algorithm creates.
 
-If a swarm is goal-based, the addition of a \textit{directional vector} will prevent all agents simultaneously producing null vectors as shown in~Fig.~\ref{concave:VoidPerimeter2} and Equation~\ref{eq:InterAgentMovementDirectional1}.
-Here, $k_dv_{d}(b)$ is the calculated directional vector to be applied to an agent.
+If a swarm is goal-based, the addition of a \textit{directional vector} will
+prevent all agents simultaneously producing null vectors as shown
+in~Fig.~\ref{concave:VoidPerimeter2} and
+Equation~\ref{eq:InterAgentMovementDirectional1}. Here, $k_dv_{d}(b)$ is the
+calculated directional vector to be applied to an agent.
 
@@
 \section{Modelling movement}\label{Section::Movement1}
-Each agent within a swarm calculates its \textit{movement-direction vector} based on its \textit{interaction} and \textit{destination} vectors.
-The \textit{movement vector}~($b_{pos}$) is calculated using the unit \textit{movement-direction vector} (Equation~\ref{eq:InterAgentMovementDirectional1}) multiplied by the time elapsed~($t$) in the system and the speed characteristic of the agent~($s_b$).
+Each agent within a swarm calculates its \textit{movement-direction vector}
+based on its \textit{interaction} and \textit{destination} vectors. The
+\textit{movement vector}~($b_{pos}$) is calculated using the unit
+\textit{movement-direction vector}
+(Equation~\ref{eq:InterAgentMovementDirectional1}) multiplied by the time
+elapsed~($t$) in the system and the speed characteristic of the agent~($s_b$).
 
-This process is carried out for every agent in the swarm to generate the swarm's next position.
+This process is carried out for every agent in the swarm to generate the
+swarm's next position.
 
@@
 
-The increment in the location of agent $b$ over time interval $t$ is shown in Equation~\ref{eq:AgentMovement} where $s_b$ is the speed of agent~$b$.
-Note that \string^ is the equivalent of $\hat{v} = \frac{v}{|v|}$ the normalised vector.
+The increment in the location of agent $b$ over time interval $t$ is shown in
+Equation~\ref{eq:AgentMovement} where $s_b$ is the speed of agent~$b$. Note
+that \string^ is the equivalent of $\hat{v} = \frac{v}{|v|}$ the normalised
+vector.
 
 \section{Swarm movement analysis\label{metric:MagnitudeDynamics2}}
-The repulsion and cohesion vectors are generated for an agent through the interaction of their field effects~(Fig.~\ref{methods:FieldEffects}).
-There are a limited number of interactions that can occur; These are illustrated in Fig.~\ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}.
+The repulsion and cohesion vectors are generated for an agent through the
+interaction of their field effects~(Fig.~\ref{methods:FieldEffects}). There are
+a limited number of interactions that can occur; These are illustrated in
+Fig.~\ref{methods:Stability1}, \ref{methods:Stability2},
+\ref{methods:Stability3} and \ref{methods:Stability4}.
 
-Following repeated experiments in the initial deployment of a swarm it was found that using the same parameters for a swarm's environmental parameters and the cohesion and repulsion ranges that the results where always very similar.
-The results shown in this paper are therefore an example set typical of the results obtained.
-The data extracts shown in Tables \ref{tab:SampleReplusion0}, \ref{tab:SampleReplusionPositive}, \ref{tab:SampleCohesionPositive} and  \ref{tab:SampleEquilibrium} are the simulation results that are produced by the parameters listed in Table~\ref{tab:MetricPhysics1}.
-The simulation consists of 200 agents over a 20 second period.
-The cohesion of an agent pair is shown as $k_cv_c$ and the repulsion as $k_rv_r$.
+Following repeated experiments in the initial deployment of a swarm it was
+found that using the same parameters for a swarm's environmental parameters and
+the cohesion and repulsion ranges that the results where always very similar.
+The results shown in this paper are therefore an example set typical of the
+results obtained. The data extracts shown in Tables \ref{tab:SampleReplusion0},
+\ref{tab:SampleReplusionPositive}, \ref{tab:SampleCohesionPositive} and
+\ref{tab:SampleEquilibrium} are the simulation results that are produced by the
+parameters listed in Table~\ref{tab:MetricPhysics1}. The simulation consists of
+200 agents over a 20 second period. The cohesion of an agent pair is shown as
+$k_cv_c$ and the repulsion as $k_rv_r$.
 
@@
 
-Figure~\ref{methods:Stability1} shows two agents within each other's cohesion fields but sufficiently distant to be outside of the repulsion fields.
-In this case $|k_cv_c| > 0$ and $|k_rv_r| = 0$: the result is the agent's resultant magnitude $v(b)$ causes the agent to move towards its neighbour $b'$.
-Likewise the neighbour's resultant magnitude $v(b')$ will cause it to move towards $b$.
-Table~\ref{tab:SampleReplusion0} shows the repulsion magnitude with a value of 0.
-The only influence on the agent pair are cohesive vectors.
+Figure~\ref{methods:Stability1} shows two agents within each other's cohesion
+fields but sufficiently distant to be outside of the repulsion fields. In this
+case $|k_cv_c| > 0$ and $|k_rv_r| = 0$: the result is the agent's resultant
+magnitude $v(b)$ causes the agent to move towards its neighbour $b'$. Likewise
+the neighbour's resultant magnitude $v(b')$ will cause it to move towards $b$.
+Table~\ref{tab:SampleReplusion0} shows the repulsion magnitude with a value of
+0. The only influence on the agent pair are cohesive vectors.
 
@@
 
-Figure~\ref{methods:Stability2} shows two agents close together with repulsion dominating cohesion: $|k_cv_c|~<~|k_rv_r|$. The resultant vector directs the agents away from each other.
+Figure~\ref{methods:Stability2} shows two agents close together with repulsion
+dominating cohesion: $|k_cv_c|~<~|k_rv_r|$. The resultant vector directs the
+agents away from each other.
 
-Table~\ref{tab:SampleReplusionPositive} shows the repulsion magnitude with a value greater than the cohesion magnitude.
+Table~\ref{tab:SampleReplusionPositive} shows the repulsion magnitude with a
+value greater than the cohesion magnitude.
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/Stability2}
-{Internal movement repulsion.\label{methods:Stability2}}
+\Figure[t!](topskip=0pt, botskip=0pt,
+midskip=0pt)[width=6cm]{figures/Stability2} {Internal movement
+repulsion.\label{methods:Stability2}}
 
@@
 
-Figure~\ref{methods:Stability3} shows two agents close together but with cohesion vector magnitudes greater than the repulsion magnitudes $|k_cv_c|~>~|k_rv_r|$.
-The resultant vector draws the agents together.
-The magnitude of the resultant cohesion vector reduces due to the cancelling effect of the repulsion vector.
-Table~\ref{tab:SampleCohesionPositive} shows a data extract with the cohesion magnitude greater than repulsion.
+Figure~\ref{methods:Stability3} shows two agents close together but with
+cohesion vector magnitudes greater than the repulsion magnitudes
+$|k_cv_c|~>~|k_rv_r|$. The resultant vector draws the agents together. The
+magnitude of the resultant cohesion vector reduces due to the cancelling effect
+of the repulsion vector. Table~\ref{tab:SampleCohesionPositive} shows a data
+extract with the cohesion magnitude greater than repulsion.
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/Stability3}
-{Internal movement cohesion and repulsion.\label{methods:Stability3}}
+\Figure[t!](topskip=0pt, botskip=0pt,
+midskip=0pt)[width=6cm]{figures/Stability3} {Internal movement cohesion and
+repulsion.\label{methods:Stability3}}
 
@@
 
-Figure~\ref{methods:Stability4} shows two agents close together with $|k_cv_c|~=~|k_rv_r|$ the resultant vector is a \textit{null vector} and the agents have no overall influence upon each other due to the magnitude of the resultant vector being zero.
-Table~\ref{tab:SampleEquilibrium} is an extract of data from the simulator.
-The data shows near equilibrium but due to the dynamic nature of a swarm system no agents meet the condition fully.
+Figure~\ref{methods:Stability4} shows two agents close together with
+$|k_cv_c|~=~|k_rv_r|$ the resultant vector is a \textit{null vector} and the
+agents have no overall influence upon each other due to the magnitude of the
+resultant vector being zero. Table~\ref{tab:SampleEquilibrium} is an extract of
+data from the simulator. The data shows near equilibrium but due to the dynamic
+nature of a swarm system no agents meet the condition fully.
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/Stability4}
-{Internal movement equilibrium.\label{methods:Stability4}}
+\Figure[t!](topskip=0pt, botskip=0pt,
+midskip=0pt)[width=6cm]{figures/Stability4} {Internal movement
+equilibrium.\label{methods:Stability4}}
 
@@
 \section{Inter-Agent Distance metric}\label{Section:DistanceDynamics}
-The inter-agent distance metric considers a swarm in terms of how the agents are physically distributed: i.e.
-only the inter-agent distances and their standard deviation are considered.
-The standard deviation indicates the extent to which the swarm is out of balance and therefore the degree of movement required to rebalance the swarm structure.
-If the standard deviation is zero then all the agents are evenly spaced.
+The inter-agent distance metric considers a swarm in terms of how the agents
+are physically distributed: i.e. only the inter-agent distances and their
+standard deviation are considered. The standard deviation indicates the extent
+to which the swarm is out of balance and therefore the degree of movement
+required to rebalance the swarm structure. If the standard deviation is zero
+then all the agents are evenly spaced.
 
@@
 
-Navarro and Fernando describe a \textit{mean distance error} metric that is based on the variations in distances between inter-agent spaces~\cite{NIM:09}.
-This is the same as the standard deviation of the distance-based internal movement metric.
+Navarro and Fernando describe a \textit{mean distance error} metric that is
+based on the variations in distances between inter-agent spaces~\cite{NIM:09}.
+This is the same as the standard deviation of the distance-based internal
+movement metric.
 
@@
 
-The mean distance for a swarm is calculated by~Equation~\ref{eq:SwarmStabilityDistance2}. All the inter-agent distances are included for the swarm ($S$).
+The mean distance for a swarm is calculated
+by~Equation~\ref{eq:SwarmStabilityDistance2}. All the inter-agent distances are
+included for the swarm ($S$).
 
@@
 %\section{Standard deviation in distance metric}\label{Section:VarianceInDistance}
-The mean distance provides a indication of the large scale structure of the swarm.
-However it is not sufficient to give an indication of the internal distribution of the agents.
-The standard deviation clarifies the distribution within the swarm as shown in~Equation~\ref{eq:SwarmStabilityDistance3}.
+The mean distance provides a indication of the large scale structure of the
+swarm. However it is not sufficient to give an indication of the internal
+distribution of the agents. The standard deviation clarifies the distribution
+within the swarm as shown in~Equation~\ref{eq:SwarmStabilityDistance3}.
 
@@
 
-The distance-based metric for the internal distribution of the agents is therefore $\mu_d(S)$ and $\sigma_d(S)$. This can be written informally as:
+The distance-based metric for the internal distribution of the agents is
+therefore $\mu_d(S)$ and $\sigma_d(S)$. This can be written informally as:
 
@@
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=14cm]{figures/BaselineDistance1}
-{Baseline internal movement - distance.\label{coord:BaselineDistance1}}
+\Figure[t!](topskip=0pt, botskip=0pt,
+midskip=0pt)[width=14cm]{figures/BaselineDistance1} {Baseline internal movement
+- distance.\label{coord:BaselineDistance1}}
 
@@
 % The \textit{`cohesion/repulsion metric'} is therefore the scalar value of the \textit{`agent resultant magnitude'} which is altered based upon the difference between the repulsion and cohesion magnitudes~(\ref{eq:CohesionRepulsion}).
-This paper defines the cohesion/repulsion metric scalar for an agent as the length of $v(b)$  as defined in Equation~\ref{eq:InterAgentMovement2}: that is, the weighted effects of cohesion and repulsion by neighbours of the agent.
-This number carries a negative sign if repulsion dominates and a positive sign if cohesion dominates.
-This is summarised in Equation~\ref{eq:CohesionRepulsion}.
 
-If an agent's {cohesion/repulsion metric} is negative the swarm's bias is to expand as the repulsion is greater than cohesion.
-This is seen in the disorganised stage of a swarm~(Fig.~\ref{coord:BaselineMagnitude1}).
-The disorganised stage is the initial random deployment of the agents.
-If the agent \textit{agent resultant magnitude} is positive then the swarm is exhibiting a tendency to contract and this indicates the swarm is a cohesive entity.
-This could also be described as the swarm being `sticky' as the agents bias is to `pull' towards each other.
+This paper defines the cohesion/repulsion metric scalar for an agent as the
+length of $v(b)$  as defined in Equation~\ref{eq:InterAgentMovement2}: that is,
+the weighted effects of cohesion and repulsion by neighbours of the agent. This
+number carries a negative sign if repulsion dominates and a positive sign if
+cohesion dominates. This is summarised in Equation~\ref{eq:CohesionRepulsion}.
 
-The \textit{cohesion/repulsion metric} on its own does not give a complete measure of a swarm's internal state.
-There needs to be a qualifying component to the metric that identifies the degree of deviation in the resultant magnitude, this is the \textit{jitter}.
-The smaller the deviation the more uniform the structure of the swarm.
-These two components identify the degree to which a swarm has progressed towards a stable state.
+If an agent's {cohesion/repulsion metric} is negative the swarm's bias is to
+expand as the repulsion is greater than cohesion. This is seen in the
+disorganised stage of a swarm~(Fig.~\ref{coord:BaselineMagnitude1}). The
+disorganised stage is the initial random deployment of the agents. If the agent
+\textit{agent resultant magnitude} is positive then the swarm is exhibiting a
+tendency to contract and this indicates the swarm is a cohesive entity. This
+could also be described as the swarm being `sticky' as the agents bias is to
+`pull' towards each other.
 
-The \textit{cohesion/repulsion metric} provides a view of the swarm's state through the balance between the repulsive and the cohesive vectors that are being applied to each agent.
-The deviation component identifies the degree to which the swarm has stabilised.
-The ideal status for inter-agent interactions would be for the agents to have a (\textit{cohesion/repulsion metric}) of zero or above.
-This would indicate that the agents are distributed such that they are at their distribution limit (outer most range of the cohesion field) or at a level that causes the agents to `pull' together.
-The ideal degree of deviation is zero as this indicates an even distribution of agents.
-These two aspects of a swarm's features are not considered by Gazi and Passino~\cite{GP:11} or Barnes et al~\cite{BFV:07} as a means of quantifying the structure of a swarm in terms of stability.
+The \textit{cohesion/repulsion metric} on its own does not give a complete
+measure of a swarm's internal state. There needs to be a qualifying component
+to the metric that identifies the degree of deviation in the resultant
+magnitude, this is the \textit{jitter}. The smaller the deviation the more
+uniform the structure of the swarm. These two components identify the degree to
+which a swarm has progressed towards a stable state.
+
+The \textit{cohesion/repulsion metric} provides a view of the swarm's state
+through the balance between the repulsive and the cohesive vectors that are
+being applied to each agent. The deviation component identifies the degree to
+which the swarm has stabilised. The ideal status for inter-agent interactions
+would be for the agents to have a (\textit{cohesion/repulsion metric}) of zero
+or above. This would indicate that the agents are distributed such that they
+are at their distribution limit (outer most range of the cohesion field) or at
+a level that causes the agents to `pull' together. The ideal degree of
+deviation is zero as this indicates an even distribution of agents. These two
+aspects of a swarm's features are not considered by Gazi and
+Passino~\cite{GP:11} or Barnes et al~\cite{BFV:07} as a means of quantifying
+the structure of a swarm in terms of stability.
 
 % \section{Cohesion/repulsion internal movement model}\label{Section:StabilityModel}
-Section~\ref{Section:StabilityMagnitude}~discussed calculation of resultant cohesion~(Equation~\ref{eq:FlyToCentre1}) and resultant repulsion~(Equation~\ref{eq:Repulsion1}) vectors for each agent deriving a combined \textit{cohesion/repulsion vector} using as shown in Equation~\ref{eq:InterAgentMovement2}.
+
+Section~\ref{Section:StabilityMagnitude}~discussed calculation of resultant
+cohesion~(Equation~\ref{eq:FlyToCentre1}) and resultant
+repulsion~(Equation~\ref{eq:Repulsion1}) vectors for each agent deriving a
+combined \textit{cohesion/repulsion vector} using as shown in
+Equation~\ref{eq:InterAgentMovement2}.
 
@@
 
-Now, we define for each agent $b$ a \textit{cohesion/repulsion scalar}, $P(b)$, to measure the influence of the neighbours of $b$ as cohesive (positive) or repulsive (negative); shown in Equation~\ref{eq:CohesionRepulsion}.
+Now, we define for each agent $b$ a \textit{cohesion/repulsion scalar}, $P(b)$,
+to measure the influence of the neighbours of $b$ as cohesive (positive) or
+repulsive (negative); shown in Equation~\ref{eq:CohesionRepulsion}.
 
@@
 
-Although it is possible for $v(b)$ to be null there could still be variation in the constituent components.
-The variation calculation (standard deviation) as shown in Equation~\ref{eq:SwarmStabilityQuotientT}.
+Although it is possible for $v(b)$ to be null there could still be variation in
+the constituent components. The variation calculation (standard deviation) as
+shown in Equation~\ref{eq:SwarmStabilityQuotientT}.
 
@@
 
-The mean cohesion/repulsion scalar for the swarm is now given by~(\ref{eq:SwarmStabilityMetricT3}).
+The mean cohesion/repulsion scalar for the swarm is now given
+by~(\ref{eq:SwarmStabilityMetricT3}).
 
@@
 
-The standard deviation associated with this mean is calculated as Equation~\ref{eq:SwarmStabilityQuotientT}.
+The standard deviation associated with this mean is calculated as
+Equation~\ref{eq:SwarmStabilityQuotientT}.
+
 %%  where $\sigma_p(S)$ is the standard deviation and $\mu_p(S)$ is the mean at the same point in time.
@@
 
-The metric for the internal movement is this pair of numbers, the mean and standard deviation of the swarm's internal \textit{cohesion/repulsion}.
-The pair $\mu_p(S)$, $\sigma_p(S)$ may be written informally as:
+The metric for the internal movement is this pair of numbers, the mean and
+standard deviation of the swarm's internal \textit{cohesion/repulsion}. The
+pair $\mu_p(S)$, $\sigma_p(S)$ may be written informally as:
 
@@
 
-Area filling is a technique used to fill an enclosed area with agents such that they are distributed as `effectively' as possible throughout an area.
+Area filling is a technique used to fill an enclosed area with agents such that
+they are distributed as `effectively' as possible throughout an area.
 
-Filling an area can be applied to tasks that require an unknown environment to be analysed or surveyed.
-Consider a disaster area following a landslide or a building collapse following an earthquake; the movement in the land and buildings will produce areas that are unmapped.
-The unmapped areas may require investigation to locate people, or resources, or to create some form of sensor network to analyse the conditions within the area, such as creating a heat map or identifying the location of toxic gases.
+Filling an area can be applied to tasks that require an unknown environment to
+be analysed or surveyed. Consider a disaster area following a landslide or a
+building collapse following an earthquake; the movement in the land and
+buildings will produce areas that are unmapped. The unmapped areas may require
+investigation to locate people, or resources, or to create some form of sensor
+network to analyse the conditions within the area, such as creating a heat map
+or identifying the location of toxic gases.
 
-The concept of using a swarm to provide coverage over unknown areas is a current area of research.
-Alvissalim et al.
-discuss the application of commercially available drones to provide a communication infrastructure across an unmapped disaster area~\cite{AZHMJJM:12}.
-Scheutz and Bauer (2006) use both cohesion and repulsion as a mechanism to create coverage of an area that requires protection in an adversarial environment~\cite{SB:06}.
+The concept of using a swarm to provide coverage over unknown areas is a
+current area of research. Alvissalim et al. discuss the application of
+commercially available drones to provide a communication infrastructure across
+an unmapped disaster area~\cite{AZHMJJM:12}. Scheutz and Bauer (2006) use both
+cohesion and repulsion as a mechanism to create coverage of an area that
+requires protection in an adversarial environment~\cite{SB:06}.
 
-The concept behind the area filling is to increase both the repulsion and cohesion fields over a period of time.
-This increase in the field effects makes the agents increase the distance between each other, expanding the swarm as a whole.
-The expansion increases until, due to boundary compression, the swarm is unable to expand further.
-In an extreme case the expansion will result in a set of field effects that create a mesh-based swarm structure rather than the desired hexagonal structure that the field effects should create.
-Identifying the \textit{cohesion/repulsion metric} falling below zero indicates area saturation has occurred.
+The concept behind the area filling is to increase both the repulsion and
+cohesion fields over a period of time. This increase in the field effects makes
+the agents increase the distance between each other, expanding the swarm as a
+whole. The expansion increases until, due to boundary compression, the swarm is
+unable to expand further. In an extreme case the expansion will result in a set
+of field effects that create a mesh-based swarm structure rather than the
+desired hexagonal structure that the field effects should create. Identifying
+the \textit{cohesion/repulsion metric} falling below zero indicates area
+saturation has occurred.
 
-To test this hypothesis a swarm is modelled in the simulator.
-The model consists of an obstacle-based enclosed space and a swarm consisting of 60 agents.
-The experiments parameters for the simulation are shown in~Table~\ref{tab:FillParameters1}
+To test this hypothesis a swarm is modelled in the simulator. The model
+consists of an obstacle-based enclosed space and a swarm consisting of 60
+agents. The experiments parameters for the simulation are shown
+in~Table~\ref{tab:FillParameters1}
 
@@
 
-The field effects are incremented in turn, cohesion range followed by repulsion range.
-After each repulsion field change the swarm is allowed to redistribute itself and stabilise.
-Table~\ref{tab:FillSequence} shows the field effect settings that are used for the simulation.
-The field effects are selected so as to ensure the swarm parameters have the potential to create a hexagonal swarm.
+The field effects are incremented in turn, cohesion range followed by repulsion
+range. After each repulsion field change the swarm is allowed to redistribute
+itself and stabilise. Table~\ref{tab:FillSequence} shows the field effect
+settings that are used for the simulation. The field effects are selected so as
+to ensure the swarm parameters have the potential to create a hexagonal swarm.
 Figure \ref{emerge:FLOOD0} shows the initial swarm deployment.
@@
 
-Figure~\ref{emerge:FLOOD1} shows the point a which data logging starts.
-The swarm at this stage has expanded into the space but is not filling it.
-This presents itself as a positive cohesion/repulsion metric value which indicates the swarm is cohesive and the agents are still ``pulling" together.
-Fig.~\ref{emerge:FLOOD6} shows the final swarm distribution for the simulation, at this point the cohesion/repulsion metric has fell below zero indicating the swarm is trying to expand.
+Figure~\ref{emerge:FLOOD1} shows the point a which data logging starts. The
+swarm at this stage has expanded into the space but is not filling it. This
+presents itself as a positive cohesion/repulsion metric value which indicates
+the swarm is cohesive and the agents are still ``pulling" together.
+Fig.~\ref{emerge:FLOOD6} shows the final swarm distribution for the simulation,
+at this point the cohesion/repulsion metric has fell below zero indicating the
+swarm is trying to expand.
 
@@
 
-Figure~\ref{emerge:FIELDFILL-MAG} shows the cohesion/repulsion metric for the swarm for the entire simulation.
-Between 100 seconds and 105 seconds there is a significant change in the magnitude where the value becomes negative.
-This indicates there is a compression effect on the swarm which indicates the area is filled.
+Figure~\ref{emerge:FIELDFILL-MAG} shows the cohesion/repulsion metric for the
+swarm for the entire simulation. Between 100 seconds and 105 seconds there is a
+significant change in the magnitude where the value becomes negative. This
+indicates there is a compression effect on the swarm which indicates the area
+is filled.
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=14cm]{figures/FIELDFILL-MAG}
-{Cohesion/Repulsion metric 0-120 seconds.\label{emerge:FIELDFILL-MAG}}
+\Figure[t!](topskip=0pt, botskip=0pt,
+midskip=0pt)[width=14cm]{figures/FIELDFILL-MAG} {Cohesion/Repulsion metric
+0-120 seconds.\label{emerge:FIELDFILL-MAG}}
 
@@
 
-Figure~\ref{emerge:FIELDFILL-DIST}~shows the distance metric for the same simulation.
-The initial deployment is shown at 0 seconds followed by incremental changes in the distances and variance of the distances as each increment is made in the field effects.
-The simulation terminates after 120 seconds.
-There is no feature of this graph which indicates that the area is filled.
-The cohesion/repulsion metric clearly tells us more.
+Figure~\ref{emerge:FIELDFILL-DIST}~shows the distance metric for the same
+simulation. The initial deployment is shown at 0 seconds followed by
+incremental changes in the distances and variance of the distances as each
+increment is made in the field effects. The simulation terminates after 120
+seconds. There is no feature of this graph which indicates that the area is
+filled. The cohesion/repulsion metric clearly tells us more.
 
-\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=14cm]{figures/FIELDFILL-DIST}
-{Distance metric 0-120 seconds.\label{emerge:FIELDFILL-DIST}}
+\Figure[t!](topskip=0pt, botskip=0pt,
+midskip=0pt)[width=14cm]{figures/FIELDFILL-DIST} {Distance metric 0-120
+seconds.\label{emerge:FIELDFILL-DIST}}
 
@@
 %% The distance metric is based upon the physical \textit{distribution} of the agents and the cohesion/repulsion metric is based upon the logical \textit{interaction} of the agents.
-Distance-based metrics measure a swarm's structure in terms of the position of agents.
-This is independent of the algorithm that produces the swarm's structure.
-It provides an analysis of agent distribution at a point in time but does not allow the identification of the future behaviour of a swarm as `expanding', `stable' or `contracting'.
+
+Distance-based metrics measure a swarm's structure in terms of the position of
+agents. This is independent of the algorithm that produces the swarm's
+structure. It provides an analysis of agent distribution at a point in time but
+does not allow the identification of the future behaviour of a swarm as
+`expanding', `stable' or `contracting'.
 
@@
 
-The new \textit{cohesion/repulsion} metric is based upon the algorithms that have produced the structure and is therefore an analysis of how the algorithms are performing rather than the structure presented.
-The new metric also provides an indication of a swarm's potential future movement.
-This is independent of the physical distribution of the agents within the swarm.
-The metric also provides the same information as the distance metric in terms of agent distribution when the field effects are consistent across all agents i.e.
-monolithic field effects.
-This is shown in Fig.
-\ref{emerge:FIELDFILL-MAG} and \ref{emerge:FIELDFILL-DIST} by the profiles of the two graphs being the same.
-In addition the new metric provides a mechanism for evaluating swarms that use different algorithms to create a swarming effect and allows the algorithms inter-agent interactions to be compared rather than the resultant structure the algorithm creates.
+The new \textit{cohesion/repulsion} metric is based upon the algorithms that
+have produced the structure and is therefore an analysis of how the algorithms
+are performing rather than the structure presented. The new metric also
+provides an indication of a swarm's potential future movement. This is
+independent of the physical distribution of the agents within the swarm. The
+metric also provides the same information as the distance metric in terms of
+agent distribution when the field effects are consistent across all agents i.e.
+monolithic field effects. This is shown in Fig. \ref{emerge:FIELDFILL-MAG} and
+\ref{emerge:FIELDFILL-DIST} by the profiles of the two graphs being the same.
+In addition the new metric provides a mechanism for evaluating swarms that use
+different algorithms to create a swarming effect and allows the algorithms
+inter-agent interactions to be compared rather than the resultant structure the
+algorithm creates.
 
-The two metrics can be used together to provide a deeper evaluation of a swarm's structure and `potential'.
-Consider the following scenario: the repulsion field is increased but the internal distances do not change; as a result the \textit{cohesion/repulsion metric} falls below zero.
-This indicates `something' is confining the swarm's distribution as discussed in Section \ref{Section:ApplicationFloodFilling}.
-This analysis could be used in identifying effective swarm distribution for the coverage of a sensor array as discussed by Ramaithitima et al.~\cite{RWBK:15}
+The two metrics can be used together to provide a deeper evaluation of a
+swarm's structure and `potential'. Consider the following scenario: the
+repulsion field is increased but the internal distances do not change; as a
+result the \textit{cohesion/repulsion metric} falls below zero. This indicates
+`something' is confining the swarm's distribution as discussed in Section
+\ref{Section:ApplicationFloodFilling}. This analysis could be used in
+identifying effective swarm distribution for the coverage of a sensor array as
+discussed by Ramaithitima et al.~\cite{RWBK:15}
 
@@
 \section{Future work}\label{Section:FutureWork}
+
 The new metric can be applied to the analysis of the effects of coordination
-algorithms upon a swarm's structure. An application to flood filling has
-been discussed~(Section~\ref{Section:ApplicationFloodFilling}). The metric can also be applied to identifying the effects of 
-`healing' algorithms, e.g. to remove `voids' from a swarm. We plan
-to investigate this, and other, practical applications of the metric.
+algorithms upon a swarm's structure. An application to area filling has been
+discussed~(Section~\ref{Section:ApplicationFloodFilling}). The metric can also
+be applied to identifying the effects of `healing' algorithms, e.g. to remove
+`voids' from a swarm. We plan to investigate this, and other, practical
+applications of the metric.
 
 The detection of a swarm's perimeter~\cite{MD:09, MJ:08, ZAPS:07, JG:13}
-identifies a subset of a swarm which can be used for coordination. 
-Perimeter detection is widely used in swarm coordination~\cite{ZFG:13, AKK:08,
+identifies a subset of a swarm which can be used for coordination.  Perimeter
+detection is widely used in swarm coordination~\cite{ZFG:13, AKK:08,
 APZDAMC:09, AZDPS:11}. A possible adaptation of this area of research is to
-investigate the reduction of the field effects (cohesion and repulsion) on the perimeter
-so as to increase the agent density in this region of the swam. This change in the swarm's structure
-could improve the granularity of analysis, e.g. when using sensor arrays at
-the frontier of the swarm.
-However, this approach leads to swarms of agents having heterogeneous field
-effects, as shown in Fig.~\ref{additional:fieldsWork}. Clearly, the
-distance-based metric is not useful in the analysis of such swarms. We believe
-that the cohesion/repulsion metric introduced in this paper can identify a
-swarm's structure even when the swarm is heterogeneous, since the metric is
-based directly upon inter-agent field effects rather than the physical
-distribution of agents. This will also be the subject of forthcoming work.
+investigate the reduction of the field effects (cohesion and repulsion) on the
+perimeter so as to increase the agent density in this region of the swam. This
+change in the swarm's structure could improve the granularity of analysis, e.g.
+when using sensor arrays at the frontier of the swarm.  However, this approach
+leads to swarms of agents having heterogeneous field effects, as shown in
+Fig.~\ref{additional:fieldsWork}. Clearly, the distance-based metric is not
+useful in the analysis of such swarms. We believe that the cohesion/repulsion
+metric introduced in this paper can identify a swarm's structure even when the
+swarm is heterogeneous, since the metric is based directly upon inter-agent
+field effects rather than the physical distribution of agents. This will also
+be the subject of forthcoming work.
 

### 2018-10-18 paper.tex
Revisions up to end of section IV

@@
 Section~\ref{Section:Related} discusses different metrics that are currently
-used to analyse swarms; Section~\ref{Section:SwarmModelling} introduces a basic
-swarm model to produce an environment that can be analysed and evaluated using
-the two metrics; Section~\ref{Section:StabilityMagnitude} discusses the
-mathematical model that produces the swarming effect and how the potential
-fields are applied; Section~\ref{Section::Movement1} discusses how the model is
-implemented over time to produce movement;
-Section~\ref{Section:StabilityNullVector} discusses the effect of the model on
-swarm stability and the effects models have on swarm movements; 
+used to analyse swarms; Section~\ref{Section:SwarmModelling} introduces a 
+model of swarm behaviour, using potential fields, that can be analysed and evaluated using
+the two metrics; Section~\ref{Section:SwarmDynamics} shows examples of the typical
+dynamic behaviour of swarms under the potential field model;
 % Section~\ref{Section:Jitter} discusses the effect the mathematical model can
@@
 %% This section discusses the theory behind swarm movement and the effects that can be measured using both the metrics described above.
-\section{Swarm modelling}\label{Section:SwarmModelling}
+\section{Swarm modelling with potential fields}\label{Section:SwarmModelling}
 Currently, much swarm research uses field effects as the method of modelling
@@
 
-\section{Inter-agent vector magnitude effect on internal movement}\label{Section:StabilityMagnitude}
+%\section{Inter-agent vector magnitude effect on internal movement}\label{Section:StabilityMagnitude}
 Figure~\ref{methods:Stability5} shows the cohesion and repulsion vector
@@
 
-\section{Internal movement and the null vector}\label{Section:StabilityNullVector}
+%\section{Internal movement and the null vector}\label{Section:StabilityNullVector}
 
@@
 
-If we consider the equilibrium state (Fig.~\ref{methods:Stability4}) the
-resultant $v(b)=(0,0)$. A null vector cannot be normalised to produce a
+If we consider the equilibrium state, % (Fig.~\ref{methods:Stability4})
+the resultant $v(b)=(0,0)$. A null vector cannot be normalised to produce a
 directional vector ($\hat{v} = \frac{v}{|v|}$ if $v\neq0$; $0$ if $v=0$) The
@@
 
-Due to the movement of agents being based only on neighbour interactions this
+Due to the movement of agents being based only on neighbour interactions, this
 situation is very rare. The residual motion that persists in a swarm is the
@@
 
-If a swarm is goal-based, the addition of a \textit{directional vector} will
-prevent all agents simultaneously producing null vectors as shown
-in~Fig.~\ref{concave:VoidPerimeter2} and
-Equation~\ref{eq:InterAgentMovementDirectional1}. Here, $k_dv_{d}(b)$ is the
-calculated directional vector to be applied to an agent.
-
+A swarm can also have a \emph{goal}. This is modelled simply by the addition of 
+a \textit{directional vector}, as shown in~Equation~\ref{eq:InterAgentMovementDirectional1}.
 \begin{equation}\label{eq:InterAgentMovementDirectional1}
@@
 \end{equation}
+Here, $k_dv_{d}(b)$ is the
+calculated directional vector to be applied to an agent.
+
+Goal-based swarms tend to 
+prevent all agents simultaneously producing null vectors, as suggested by
+Fig.~\ref{concave:VoidPerimeter1}~and~\ref{concave:VoidPerimeter2}.
 
@@
 
-\section{Modelling movement}\label{Section::Movement1}
+\section{Swarm dynamics}\label{Section:SwarmDynamics}
+%\section{Modelling movement}\label{Section::Movement1}
 Each agent within a swarm calculates its \textit{movement-direction vector}
-based on its \textit{interaction} and \textit{destination} vectors. The
+based on its \textit{inter-agent} and \textit{directional} vectors. The
 \textit{movement vector}~($b_{pos}$) is calculated using the unit
 \textit{movement-direction vector}
-(Equation~\ref{eq:InterAgentMovementDirectional1}) multiplied by the time
-elapsed~($t$) in the system and the speed characteristic of the agent~($s_b$).
+(Equation~\ref{eq:InterAgentMovementDirectional1}), multiplied by the time
+elapsed~($t$) in the system, and the speed characteristic of the agent~($s_b$).
 
@@
 
-The increment in the location of agent $b$ over time interval $t$ is shown in
-Equation~\ref{eq:AgentMovement} where $s_b$ is the speed of agent~$b$. Note
-that \string^ is the equivalent of $\hat{v} = \frac{v}{|v|}$ the normalised
+The change in the location of agent $b$, over time interval $t$, is shown in
+Equation~\ref{eq:AgentMovement}, where $s_b$ is the speed of agent~$b$. Note
+that \string^ is the equivalent of $\hat{v} = \frac{v}{|v|}$, the normalised
 vector.
 
-\section{Swarm movement analysis\label{metric:MagnitudeDynamics2}}
 The repulsion and cohesion vectors are generated for an agent through the
-interaction of their field effects~(Fig.~\ref{methods:FieldEffects}). There are
-a limited number of interactions that can occur; These are illustrated in
+interaction of its field effects~(Fig.~\ref{methods:FieldEffects}). There are
+a limited number of interactions that can occur. These are illustrated in
 Fig.~\ref{methods:Stability1}, \ref{methods:Stability2},
@@
 
-Following repeated experiments in the initial deployment of a swarm it was
-found that using the same parameters for a swarm's environmental parameters and
-the cohesion and repulsion ranges that the results where always very similar.
-The results shown in this paper are therefore an example set typical of the
+Following repeated experiments in the initial deployment of a swarm, it was
+found that, using the same parameters for a swarm's environment and
+cohesion and repulsion ranges, always produced very similar behaviour.
+The results shown in this section are an example set typical of the
 results obtained. The data extracts shown in Tables \ref{tab:SampleReplusion0},
@@
 fields but sufficiently distant to be outside of the repulsion fields. In this
-case $|k_cv_c| > 0$ and $|k_rv_r| = 0$: the result is the agent's resultant
+case $|k_cv_c| > 0$ and $|k_rv_r| = 0$: the result is that the agent's resultant
 magnitude $v(b)$ causes the agent to move towards its neighbour $b'$. Likewise
@@
 Figure~\ref{methods:Stability4} shows two agents close together with
-$|k_cv_c|~=~|k_rv_r|$ the resultant vector is a \textit{null vector} and the
-agents have no overall influence upon each other due to the magnitude of the
-resultant vector being zero. Table~\ref{tab:SampleEquilibrium} is an extract of
-data from the simulator. The data shows near equilibrium but due to the dynamic
-nature of a swarm system no agents meet the condition fully.
+$|k_cv_c|~=~|k_rv_r|$. The resultant vector is a \textit{null vector} and the
+agents have no overall influence upon each other, since the magnitude of the
+resultant vector is zero. Table~\ref{tab:SampleEquilibrium} is an extract of
+data from the simulator. The data shows near equilibrium, but due to the dynamic
+nature of a swarm system, no agents meet the condition fully.
 
@@
 
-Section~\ref{Section:StabilityMagnitude}~discussed calculation of resultant
+Section~\ref{Section:SwarmModelling}~discussed calculation of resultant
 cohesion~(Equation~\ref{eq:FlyToCentre1}) and resultant

### 2018-10-18 paper.tex

@@
   algorithms. These metrics are usually based upon the distances between
-  agents, the distance between agents and a fixed point, or the resultant vectors
-  that \textit{potential field} effects produce. This paper examines a
+  agents, the distance between agents and a fixed point, or the resultant
+  vectors that \textit{potential field} effects produce. This paper examines a
   distance-based metric that measures a swarm's overall structure using
-  inter-agent distances. It introduces a new metric that identifies a swarm's
-  state based upon the resultant magnitude of the vectors produced by the
-  agent interactions that create the agent distribution within the swarm's
-  structure. The algorithms used to implement the swarming feature are based
-  upon cohesion and repulsion vectors between an agent and its neighbours. In
-  comparing and contrasting the two metrics, we find that the
+  inter-agent distances. More importantly, it introduces a new metric that
+  identifies a swarm's state based upon the resultant magnitude of the vectors
+  produced by the agent interactions that create the agent distribution within
+  the swarm's structure. The algorithms used to implement the swarming feature
+  are based upon cohesion and repulsion vectors between an agent and its
+  neighbours. In comparing and contrasting the two metrics, we find that the
   \textit{cohesion/repulsion} metric offers a number of advantages over the
@@
 
-A swarm can also have a \emph{goal}. This is modelled simply by the addition of 
+A swarm can also have a \textit{goal}. This is modelled simply by the addition of 
 a \textit{directional vector}, as shown in~Equation~\ref{eq:InterAgentMovementDirectional1}.
@@
 \label{eq:AgentMovement}
-b_{pos}=s_{b}t\big(v(b)\big)\string^ 
+  b_{pos}=s_{b}t\widehat{v(b)} 
 \end{equation}
@@
 Navarro and Fernando describe a \textit{mean distance error} metric that is
-based on the variations in distances between inter-agent spaces~\cite{NIM:09}.
+based on the variations in distances between agents~\cite{NIM:09}.
 This is the same as the standard deviation of the distance-based internal
-movement metric.
+movement metric described in this section.
 
@@
 % \section{Calculating distance internal movement}
-Equation~\ref{eq:SwarmStabilityDistance1} calculates the mean distance of an agent to its neighbours $nbr(b)$. 
+Equation~\ref{eq:SwarmStabilityDistance1} calculates the mean distance of an agent $b$ to its neighbours $nbr(b)$. 
 %% The relative position vector generated for an agent $b$ to its neighbour $b'$, $bb'$, is shown in~(\ref{eq:FlyToCentre1}).
@@
 
-The mean distance for a swarm is calculated
+The mean distance for a swarm is given 
 by~Equation~\ref{eq:SwarmStabilityDistance2}. All the inter-agent distances are
@@
 %\section{Standard deviation in distance metric}\label{Section:VarianceInDistance}
-The mean distance provides a indication of the large scale structure of the
+The mean distance provides an indication of the large scale structure of the
 swarm. However it is not sufficient to give an indication of the internal
@@
 \end{equation}
+
+Figure~\ref{coord:BaselineDistance1} shows a short example of the evolution 
+of the metric over time.
 
@@
 
-This paper defines the cohesion/repulsion metric scalar for an agent as the
-length of $v(b)$  as defined in Equation~\ref{eq:InterAgentMovement2}: that is,
-the weighted effects of cohesion and repulsion by neighbours of the agent. This
-number carries a negative sign if repulsion dominates and a positive sign if
-cohesion dominates. This is summarised in Equation~\ref{eq:CohesionRepulsion}.
-
-If an agent's {cohesion/repulsion metric} is negative the swarm's bias is to
-expand as the repulsion is greater than cohesion. This is seen in the
-disorganised stage of a swarm~(Fig.~\ref{coord:BaselineMagnitude1}). The
-disorganised stage is the initial random deployment of the agents. If the agent
-\textit{agent resultant magnitude} is positive then the swarm is exhibiting a
-tendency to contract and this indicates the swarm is a cohesive entity. This
-could also be described as the swarm being `sticky' as the agents bias is to
-`pull' towards each other.
-
-The \textit{cohesion/repulsion metric} on its own does not give a complete
+This paper defines the \textit{cohesion/repulsion metric} for an agent as the
+length of $v(b)$,  as defined in Equation~\ref{eq:InterAgentMovement2}: that is
+the size of the weighted effects of cohesion and repulsion caused by neighbours of the
+agent. This number carries a negative sign if repulsion dominates and a
+positive sign if cohesion dominates. This is summarised in
+Equation~\ref{eq:CohesionRepulsion}.
+
+If the overall cohesion/repulsion metric of the swarm is negative, the
+swarm's tendency is to expand, as the repulsion is greater than cohesion. This is
+seen in the disorganised stage of a
+swarm~(Fig.~\ref{coord:BaselineMagnitude1}). The disorganised stage is the
+initial random deployment of the agents. If the overall cohesion/repulsion
+metric is positive, the swarm's tendency is to contract, and
+suggests that the swarm is a cohesive entity. This could also be described as
+the swarm being `sticky', as the agents' tendency is to `pull' towards each
+other.
+
+The cohesion/repulsion metric on its own does not give a complete
 measure of a swarm's internal state. There needs to be a qualifying component
 to the metric that identifies the degree of deviation in the resultant
-magnitude, this is the \textit{jitter}. The smaller the deviation the more
+magnitudes; this is the \textit{jitter}. The smaller the deviation, the more
 uniform the structure of the swarm. These two components identify the degree to
@@
 
-The \textit{cohesion/repulsion metric} provides a view of the swarm's state
+The cohesion/repulsion metric provides a view of the swarm's state
 through the balance between the repulsive and the cohesive vectors that are
@@
 which the swarm has stabilised. The ideal status for inter-agent interactions
-would be for the agents to have a (\textit{cohesion/repulsion metric}) of zero
+would be for the agents to have a cohesion/repulsion metric of zero
 or above. This would indicate that the agents are distributed such that they
@@
 cohesion~(Equation~\ref{eq:FlyToCentre1}) and resultant
-repulsion~(Equation~\ref{eq:Repulsion1}) vectors for each agent deriving a
-combined \textit{cohesion/repulsion vector} using as shown in
+repulsion~(Equation~\ref{eq:Repulsion1}) vectors for each agent, deriving a
+combined cohesion/repulsion vector, as shown in
 Equation~\ref{eq:InterAgentMovement2}.
@@
 Although it is possible for $v(b)$ to be null there could still be variation in
-the constituent components. The variation calculation (standard deviation) as
+the constituent components. The variation calculation (standard deviation) is
 shown in Equation~\ref{eq:SwarmStabilityQuotientT}.
@@
 The mean cohesion/repulsion scalar for the swarm is now given
-by~(\ref{eq:SwarmStabilityMetricT3}).
+by~Equation~\ref{eq:SwarmStabilityMetricT3}.
 
@@
 
-Figure~\ref{coord:BaselineMagnitude1} shows the evolution of the metric in time.
+Figure~\ref{coord:BaselineMagnitude1} shows a short example of the evolution of
+the metric over time.
 
@@
 Area filling is a technique used to fill an enclosed area with agents such that
-they are distributed as `effectively' as possible throughout an area.
+they are distributed as `effectively' as possible throughout the area.
 
@@
 consists of an obstacle-based enclosed space and a swarm consisting of 60
-agents. The experiments parameters for the simulation are shown
+agents. The experimental parameters for the simulation are shown
 in~Table~\ref{tab:FillParameters1}
@@
 The field effects are incremented in turn, cohesion range followed by repulsion
-range. After each repulsion field change the swarm is allowed to redistribute
+range. After each pair of increments, the swarm is allowed to redistribute
 itself and stabilise. Table~\ref{tab:FillSequence} shows the field effect
@@
 
-Figure~\ref{emerge:FLOOD1} shows the point a which data logging starts. The
+Figure~\ref{emerge:FLOOD1} shows the point at which data logging starts. The
 swarm at this stage has expanded into the space but is not filling it. This
-presents itself as a positive cohesion/repulsion metric value which indicates
+presents itself as a positive cohesion/repulsion metric value, which indicates
 the swarm is cohesive and the agents are still ``pulling" together.
-Fig.~\ref{emerge:FLOOD6} shows the final swarm distribution for the simulation,
-at this point the cohesion/repulsion metric has fell below zero indicating the
-swarm is trying to expand.
+Fig.~\ref{emerge:FLOOD6} shows the final swarm distribution for the simulation.
+At this point the cohesion/repulsion metric has fallen below zero, indicating
+the swarm is trying to expand.
 
@@
 significant change in the magnitude where the value becomes negative. This
-indicates there is a compression effect on the swarm which indicates the area
-is filled.
+indicates that there is a compression effect on the swarm and, hence, that the
+area has been filled.
 
@@
 Figure~\ref{emerge:FIELDFILL-DIST}~shows the distance metric for the same
-simulation. The initial deployment is shown at 0 seconds followed by
-incremental changes in the distances and variance of the distances as each
+simulation. The initial deployment is shown at 0 seconds, followed by
+incremental changes in the distances, and variance of the distances, as each
 increment is made in the field effects. The simulation terminates after 120
@@
 monolithic field effects. This is shown in Fig. \ref{emerge:FIELDFILL-MAG} and
-\ref{emerge:FIELDFILL-DIST} by the profiles of the two graphs being the same.
-In addition the new metric provides a mechanism for evaluating swarms that use
-different algorithms to create a swarming effect and allows the algorithms
-inter-agent interactions to be compared rather than the resultant structure the
-algorithm creates.
+\ref{emerge:FIELDFILL-DIST}, where the profiles of the two graphs are the same.
+In addition, the new metric provides a mechanism for evaluating swarms that use
+different algorithms to create a swarming effect and allows the algorithms'
+inter-agent interactions to be compared rather than the resultant structures
+that the algorithms create.
 

### 2018-10-18 paper.tex
Minors to definitions

@@
 visible neighbours. This is different to using the \textit{centroidal} approach,
-as the measure is based on agent interactions rather than a resultant structure
-which affects the centroid. This technique is the most popular approach to
+as the measure is based on a resultant structure which affects the centroid. 
+This technique is the most popular approach to
 analysing a swarm's structure. Gazi and Passino use a distance-based analysis
